\subsection{Logical theories}\label{subsec:logical_theories}

\begin{definition}\label{def:first_order_theory}\mcite[def. 17.1]{OpenLogicFull}
  The \term{closure} of the set \( \Gamma \) of formulas in the \hyperref[def:first_order_syntax]{first-order language} \( \mscrL \) is the set
  \begin{equation*}
    \cl(\Gamma) \coloneqq \set{ \varphi \in \boldop{Form} \given \Gamma \vDash \varphi }.
  \end{equation*}

  Due to \fullref{rem:propositional_logic_as_first_order_logic}, this definition also holds for formulas in propositional logic.

  The set \( \Gamma \) is called \term{closed} if it equals its own closure. A closed set of closed formulas is also sometimes called a \term{theory} in e.g. \cite[def. 33.1]{OpenLogicFull}, but we will not put this restriction because it is not clear what are the axioms of an arbitrary closed set of formulas. Instead, we will use the term \enquote{theory} to mean the set of axioms itself.

  As discussed in \fullref{rem:deduction_with_free_variables}, we are only interested in closed formulas. If a formula is not closed, we instead consider its \hyperref[thm:implicit_universal_quantification]{universal closure}.

  If it is necessary to distinguish between derivability and entailment, we can use the syntactic closure \( \cl^\vdash \) and semantic closure \( \cl^\vDash \).

  \begin{thmenum}
    \thmitem{def:first_order_theory/axiomatized}\mcite[def. 17.1]{OpenLogicFull} We say that the set \( \Gamma \) is \term{axiomatized} by \( \Delta \) if \( \Gamma = \cl(\Delta) \).

    \medskip

    \thmitem{def:first_order_theory/complete}\mcite[def. 33.7]{OpenLogicFull} The set \( \Gamma \) of formulas is said to be \term{complete} if every for every formula in \( \varphi \), either \( \Gamma \vDash \varphi \) or \( \Gamma \vDash \neg \varphi \).

    This is not to be confused with \fullref{def:derivability_and_satisfiability/completeness}, which defines completeness of a deductive system via how it relates to semantics.
  \end{thmenum}
\end{definition}

\begin{definition}\label{def:first_order_theory_consistency}\mcite[def. 22.6]{OpenLogicFull}
  Given a deductive system, the set \( \Gamma \) of formulas is said to be \term[bg=противоречива,ru=противоречивая]{inconsistent} if \( \Gamma \vdash \bot \) and \term{consistent} otherwise.
\end{definition}

\begin{proposition}\label{thm:formulas_unsatisfiable_iff_inconsistent}
   Assuming \hyperref[def:classical_logic]{classical logic}, a theory \( \Gamma \) is \hyperref[def:propositional_semantics/satisfiability]{unsatisfiable} if and only if it is semantically \hyperref[def:first_order_theory_consistency]{inconsistent}.
\end{proposition}
\begin{proof}
  \SufficiencySubProof Assume first that \( \Gamma \) is unsatisfiable. Then for all zero models of \( \Gamma \) any formula is satisfied vacuously, in particular that any model of \( \Gamma \) satisfies \( \bot \). Thus, \( \Gamma \vDash \bot \) and, by \fullref{thm:classical_first_order_logic_is_sound_and_complete}, \( \Gamma \vdash \bot \). Therefore, \( \Gamma \) is inconsistent.

  \NecessitySubProof Let \( \Gamma \) be inconsistent and suppose that \( \mscrX = (X, I) \) is a model of \( \Gamma \). Fix any valuation \( v \) in \( \mscrX \). Since \( \Gamma \vdash \bot \), \fullref{thm:classical_first_order_logic_is_sound_and_complete} implies that \( \bot\Bracks{v} = T \).

  By the \hyperref[def:first_order_valuation/formula_valuation]{definition of formula valuation}, however, we have \( \bot\Bracks{v} = F \). The obtained contradiction shows that \( \mscrX \) cannot be a model of \( \Gamma \) and since this structure was chosen arbitrarily, we conclude that \( \Gamma \) is unsatisfiable.
\end{proof}

\begin{theorem}[First-order compactness theorem]\label{thm:first_order_compactness_theorem}\mcite[thm. 23.21]{OpenLogicFull}
  Assuming \hyperref[def:classical_logic]{classical logic}, for a theory \( \Gamma \) we have \( \Gamma \vDash \varphi \) for the closed formula \( \varphi \) if and only if there exists a finite subset of \( \Delta \subseteq \Gamma \) such that \( \Delta \vDash \varphi \).
\end{theorem}
\begin{proof}
  \SufficiencySubProof Suppose that \( \Gamma \vDash \varphi \). By \fullref{thm:classical_first_order_logic_is_sound_and_complete}, \( \Gamma \vdash \varphi \). Every \hyperref[def:proof_tree]{proof tree} is finite, and thus \( \varphi \) can be derived from only the finitely many premises \( \Delta \) of the proof. Then \( \Delta \vdash \varphi \) and, by the other direction of \fullref{thm:classical_first_order_logic_is_sound_and_complete}, we have \( \Delta \vDash \varphi \).

  \NecessitySubProof If \( \Delta \vDash \varphi \) for some finite \( \Delta \subseteq \Gamma \), then necessarily \( \Gamma \vDash \varphi \).
\end{proof}

\begin{corollary}\label{thm:first_order_compactness_theorem_satisfiability}\mcite[thm. 23.21]{OpenLogicFull}
  If every finite subset of \( \Gamma \) is satisfiable, then \( \Gamma \) itself is satisfiable.

  This is also called \enquote{the compactness theorem}.
\end{corollary}
\begin{proof}
  We can restate the theorem using its contraposition: If \( \Gamma \) is unsatisfiable, then there exists a finite subset \( \Delta \subseteq \Gamma \) that is also unsatisfiable. Due to \fullref{thm:formulas_unsatisfiable_iff_inconsistent}, this can be also restated as: If \( \Gamma \) is inconsistent, then there exists a finite subset \( \Delta \subseteq \Gamma \) that is also inconsistent.

  But the last formulation is a special case of \fullref{thm:first_order_compactness_theorem} with \( \varphi = \bot \).
\end{proof}

\begin{theorem}[Downward L\"owenheim-Skolem theorem]\label{thm:downward_lowenheim_skolem_theorem}\mcite{nLab:lowenheim_skolem_theorem}
  Let \( \Gamma \) be a \hyperref[def:first_order_theory]{first-order theory} over some arbitrary language. Suppose that \( \Gamma \) has a \hyperref[def:first_order_semantics/satisfiability]{model} of \hyperref[def:set_finiteness]{infinite} \hyperref[thm:cardinality_existence]{cardinality}.

  Then \( \Gamma \) also has a \hyperref[def:set_countability]{countable} model.
\end{theorem}

\begin{example}[Skolem's paradox]\label{ex:skolems_paradox}\mcite[ex. 23.32]{OpenLogicFull}
  From \fullref{thm:downward_lowenheim_skolem_theorem} it follows that \hyperref[def:zfc]{\logic{ZFC}}, if it is consistent, has a model at is at most countably infinite. The \hyperref[def:zfc/infinity]{axiom of infinity} states that no model of \logic{ZFC} is finite. Therefore, there exists a model of \logic{ZFC} that is countably infinite. But we can construct uncountable sets in \logic{ZFC}.

  Therefore, either:
  \begin{itemize}
    \item Uncountable sets within the metatheory are fundamentally different from those within the object theory.
    \item Uncountable sets are paradoxical and \logic{ZFC} must disallow them in order to be consistent.
    \item \logic{ZFC} is inconsistent even when restricted to countable sets.
  \end{itemize}
\end{example}

\begin{theorem}[Upward L\"owenheim-Skolem theorem]\label{thm:upward_lowenheim_skolem_theorem}\mcite{nLab:lowenheim_skolem_theorem}
  Let \( \Gamma \) be a \hyperref[def:first_order_theory]{first-order theory} over some arbitrary language. Suppose that \( \Gamma \) has a \hyperref[def:first_order_semantics/satisfiability]{model} of \hyperref[def:set_finiteness]{infinite} \hyperref[thm:cardinality_existence]{cardinality} \( \kappa \).

  Then for every \hyperref[def:set_finiteness]{infinite} \hyperref[def:cardinal]{cardinal} \( \mu \geq \card(\Gamma) \), the theory \( \Gamma \) also has a model of cardinality \( \mu \).
\end{theorem}

\begin{definition}\label{def:category_of_small_first_order_models}
  Let \( \mscrL \) be a first-order language and let \( \Gamma \) be a nonempty set of formulas. We describe the \term{categories of small models} for \( \Gamma \) as the \hyperref[def:concrete_category]{concrete category}. See \fullref{rem:positive_formulas_in_category_of_models} for why it is a reasonable restriction for \( \Gamma \) to be a set of \hyperref[def:positive_formula]{positive formulas}.

  Suppose that we are given a \hyperref[def:grothendieck_universe]{Grothendieck universe} \( \mscrU \), which is safe to assume to be the smallest suitable one as explained in \fullref{def:large_and_small_sets}.

  \begin{itemize}
    \item The \hyperref[def:category/objects]{set of objects} is the set of all \( \mscrU \)-small models of \( \Gamma \).

    \item The \hyperref[def:category/morphisms]{set of morphisms} between two models is the set of all \hyperref[def:first_order_homomorphism]{structure homomorphisms} between them.

    \item \hyperref[def:category/composition]{Composition of morphisms} is the usual \hyperref[def:multi_valued_function/composition]{function composition}.

    \item The \hyperref[def:category/identity]{identity morphism} on a model is the \hyperref[def:multi_valued_function/identity]{identity function} on its universe.
  \end{itemize}
\end{definition}

\begin{remark}\label{rem:positive_formulas_in_category_of_models}
  We usually want \( \Gamma \) to be a set of \hyperref[def:positive_formula]{positive formulas} because, in general, homomorphisms are not injective and the homomorphic image of a model may fail to be a model. \Fullref{thm:positive_formulas_preserved_under_homomorphism} shows that if all formulas in \( \Gamma \) are positive, the image of any homomorphism is again a model and thus an object in the category.
\end{remark}

\begin{example}\label{ex:def:category_of_small_first_order_models}
  This is an incomplete list of categories corresponding to \hyperref[def:first_order_theory]{first-order theories} that can be found in this document:
  \begin{itemize}
    \item The categories \hyperref[def:pointed_set/category]{\( \cat{Set}_* \)}, \hyperref[def:set_with_involution/category]{\( \cat{Inv} \)}, \hyperref[def:magma/category]{\( \cat{Mag} \)}, \hyperref[def:unital_magma/category]{\( \cat{Mag}_* \)}, \hyperref[def:unital_magma/associative]{\( \cat{Mon} \)}, \hyperref[def:group/category]{\( \cat{Grp} \)} and \hyperref[def:abelian_group]{\( \cat{Ab} \)}, whose relations are crucial for the definition and properties of groups.

    \item The categories \hyperref[def:semiring/ring]{\( \cat{Ring} \)} of rings, \hyperref[def:semiring/ring]{\( \cat{Mod}_\mscrR \)} of modules and \hyperref[def:vector_space]{\( \cat{Vect}_\BbbK \)} of vector spaces, which are both based on abelian groups.

    \item The categories \hyperref[def:category_of_topological_groups]{\( \cat{TopGrp} \)} and \hyperref[def:category_of_topological_groups]{\( \cat{TopVect}_\BbbK \)} of topological groups and vector spaces, respectively.

    \item The categories \hyperref[def:partially_ordered_set/category]{\( \cat{Pos} \)} in order theory and the related \hyperref[def:semilattice/category]{\( \cat{Lat} \)}, \hyperref[def:heyting_algebra/category]{\( \cat{Heyt} \)} and \hyperref[def:boolean_algebra/category]{\( \cat{Bool} \)} in lattice theory.
  \end{itemize}

  In contrast:
  \begin{itemize}
    \item We define the category \hyperref[def:category_of_small_topological_spaces]{\( \cat{Top} \)}  of topological spaces and all of its related categories within set theory without a corresponding first-order theory.

    \item The category \hyperref[def:category_of_small_sets]{\( \cat{Set} \)} of sets with respect to either \logic{ZFC}, \logic{ZFC+U} or na\"ive set theory is not the same as the category of \( \mscrU \)-small models of set theory. Instead, it is a category within a fixed set theory. Within the metatheory of this document, we work within a fixed model of \logic{ZFC+U} with respect to \hyperref[def:classical_logic]{classical logic}.

    \item Similarly, we do not care about models of \hyperref[def:peano_arithmetic]{Peano arithmetic} enough to study its category of \( \mscrU \)-small models. Instead, we only use a single model and denote it by \( \BbbN \).
  \end{itemize}
\end{example}

\begin{definition}\label{def:lindenbaum_tarski_algebra}\mcite{nLab:lindenbaum_tarski_algebra}
  Assume some fixed \hyperref[def:deductive_system]{deductive system} in propositional or first-order logic. Let \( \Gamma \) be a \hyperref[def:first_order_theory]{closed set} of \hyperref[def:first_order_syntax/ground_formula]{closed formulas} within the corresponding logic.

  Then \( (\Gamma, \vdash) \) is a \hyperref[def:preordered_set]{preordered set}. The \term{Lindenbaum-Tarski algebra} of the theory \( \Gamma \) is the partially ordered set obtained from \( (\Gamma, \vdash) \) using \fullref{thm:preorder_to_partial_order}.

  More concretely, the Lindenbaum-Tarski algebra of \( \Gamma \) is a quotient set of \( \Gamma \) by the relation \hyperref[def:proof_derivability]{interderivability} and endowed with the partial order
  \begin{equation}\label{eq:def:lindenbaum_tarski_algebra/order}
    [\varphi] \leq [\psi] \T{if and only if} \varphi \vdash \psi.
  \end{equation}

  Of course, we can define the algebra using entailment rather than derivability, but in the cases we consider, the two are equivalent and derivability is simpler to work with.
\end{definition}
\begin{proof}
  The correctness of \eqref{eq:def:lindenbaum_tarski_algebra/order}, i.e. the fact that the relation \( \leq \) does not depend on the choice of representatives from the quotient sets, follows from \fullref{thm:preorder_to_partial_order}.

  We must only demonstrate that \( (\Gamma, \vdash) \) is indeed a preordered set. Reflexivity of \( \vdash \) follows from \fullref{def:axiomatic_deductive_system} and transitivity follows from \fullref{thm:deductive_system_transitivity}.
\end{proof}

\begin{proposition}\label{thm:intuitionistic_lindenbaum_tarski_algebra}
  Assume that we are working in the \hyperref[def:intuitionistic_propositional_deductive_systems]{intuitionistic propositional natural deduction system}. The \hyperref[def:lindenbaum_tarski_algebra]{Lindenbaum-Tarski algebra} of a closed set of closed formulas \( \Gamma \) then is a \hyperref[def:heyting_algebra]{Heyting algebra}.

  In the \hyperref[def:classical_propositional_deductive_systems]{classical derivation system}, the algebra is instead a \hyperref[def:boolean_algebra]{Boolean algebra}.

  In the \hyperref[def:minimal_propositional_axiomatic_deductive_system]{minimal derivation system}, we have an unbounded lattice with only a top element, but no bottom. Consequently, conditionals and pseudocomplements may fail to exist.

  Explicitly:
  \begin{thmenum}
    \thmitem{thm:intuitionistic_lindenbaum_tarski_algebra/join} The \hyperref[def:semilattice/join]{join} of the equivalence classes \( [\psi_1] \) and \( [\psi_2] \) is the class \( [\psi_1 \vee \psi_2] \) of their \hyperref[def:propositional_language/connectives/disjunction]{disjunction}.

    \thmitem{thm:intuitionistic_lindenbaum_tarski_algebra/top} The class of \hyperref[def:propositional_semantics/tautology]{tautologies} \( [\top] \) is the \hyperref[def:partially_ordered_set_extremal_points/top_and_bottom]{top element}.

    \thmitem{thm:intuitionistic_lindenbaum_tarski_algebra/meet} Similarly to joins, the \hyperref[def:semilattice/meet]{meet} of \( [\psi_1] \) and \( [\psi_2] \) is the equivalence class \( [\psi_1 \wedge \psi_2] \) of their \hyperref[def:propositional_language/connectives/conjunction]{conjunction}.

    \thmitem{thm:intuitionistic_lindenbaum_tarski_algebra/bottom} The class of \hyperref[def:propositional_semantics/contradiction]{contradictions} \( [\bot] \) is the \hyperref[def:partially_ordered_set_extremal_points/top_and_bottom]{bottom element}.

    \thmitem{thm:intuitionistic_lindenbaum_tarski_algebra/conditional} The \hyperref[eq:def:heyting_algebra/conditional]{conditional} of \( [\psi_1] \) and \( [\psi_2] \) is the equivalence class \( [\psi_1 \rightarrow \psi_2] \).

    \thmitem{thm:intuitionistic_lindenbaum_tarski_algebra/complement} The \hyperref[eq:def:heyting_algebra/pseudocomplement]{pseudocomplement} \( \widetilde{[\psi]} \) of \( [\psi] \) equals \( [\neg \psi] \).

    In the classical derivation system this pseudocomplement is a complement, i.e. it satisfies \eqref{def:bounded_lattice_complement/join} and \eqref{def:bounded_lattice_complement/meet}.
  \end{thmenum}
\end{proposition}
\begin{proof}
  \SubProofOf{thm:intuitionistic_lindenbaum_tarski_algebra/join} We will show that \( [\psi_1 \vee \psi_2] \) is the supremum of \( [\psi_1] \) and \( [\psi_2] \).

  The inference rule \eqref{eq:def:minimal_propositional_natural_deduction_system/or/intro_left} implies that \( \psi_1 \vdash \psi_1 \vee \psi_2 \) and \eqref{eq:def:minimal_propositional_natural_deduction_system/or/intro_right} implies that \( \psi_2 \vdash \psi_1 \vee \psi_2 \). Thus, \( \psi_1 \vee \psi_2 \) is an upper bound for both \( \psi_1 \) and \( \psi_2 \) under the ordering \( \vdash \).

  Let \( \varphi \) be any formula in \( \Gamma \) such that \( \psi_1 \vdash \varphi \), \( \psi_2 \vdash \varphi \) and \( \varphi \vdash (\psi_1 \vee \psi_2) \). Then the following instance of \eqref{eq:def:minimal_propositional_natural_deduction_system/or/elim}
  \begin{equation*}
    \begin{prooftree}
      \hypo{ [\psi_1 \vee \psi_2] }
      \hypo{ [\psi_1]^1 }
      \ellipsis {} { \varphi }
      \hypo{ [\psi_2]^1 }
      \ellipsis {} { \varphi }
      \infer[left label=\( 1 \)]3[\ref{eq:def:minimal_propositional_natural_deduction_system/or/elim}]{ \varphi }
    \end{prooftree}
  \end{equation*}
  demonstrates that \( \psi_1, \psi_2, (\psi_1 \vee \psi_2) \vdash \varphi \). Hence, from \fullref{thm:deductive_system_transitivity} it follows that \( (\psi_1 \vee \psi_2) \vdash \varphi \).

  Obviously interderivability is not affected by the choice of representatives from \( [\psi_1] \) and \( [\psi_2] \), hence \( [\varphi] = [\psi_1 \vee \psi_2] \) and this is indeed the supremum of \( [\psi_1] \) and \( [\psi_2] \).

  \SubProofOf{thm:intuitionistic_lindenbaum_tarski_algebra/top} The inference rule \eqref{eq:def:minimal_propositional_natural_deduction_system/top/intro} shows that \( [\varphi] \leq [\top] \) for any formula \( \varphi \) and \( [\top] \) is the top element.

  \SubProofOf{thm:intuitionistic_lindenbaum_tarski_algebra/meet} Let \( \psi_1 \) and \( \psi_2 \) be any formulas in \( \Gamma \). The inference rule \eqref{eq:def:minimal_propositional_natural_deduction_system/and/elim_left} implies that \( \psi_1 \wedge \psi_2 \vdash \psi_2 \) and \eqref{eq:def:minimal_propositional_natural_deduction_system/and/elim_right} implies that \( \psi_1 \wedge \psi_2 \vdash \psi_1 \). Thus, \( \psi_1 \wedge \psi_2 \) is a lower bound for both \( \psi_1 \) and \( \psi_2 \) under the ordering \( \vdash \).

  We must show that \( \psi_1 \wedge \psi_2 \) is interderivable with the greatest lower bound. Let \( \varphi \) be a formula in \( \Gamma \) such that \( \varphi \vdash \psi_1 \), \( \varphi \vdash \psi_2 \) and \( (\psi_1 \wedge \psi_2) \vdash \varphi \). We will show that \( \varphi \vdash (\psi_1 \wedge \psi_2) \).

  The rule \eqref{eq:def:minimal_propositional_natural_deduction_system/and/intro} implies that
  \begin{equation*}
    \psi_1, \psi_2 \vdash \psi_1 \wedge \psi_2.
  \end{equation*}

  But \( \varphi \) derives both \( \psi_1 \) and \( \psi_2 \), hence due to \fullref{thm:deductive_system_transitivity},
  \begin{equation*}
    \varphi \vdash (\psi_1 \wedge \psi_2).
  \end{equation*}

  Analogously to the proof of \fullref{thm:intuitionistic_lindenbaum_tarski_algebra/join}, we conclude that the choice of representatives of the equivalence classes is irrelevant and that \( [\varphi] = [\psi_1 \wedge \psi_2] \) is the infimum of \( [\psi_1] \) and \( [\psi_2] \).

  \SubProofOf{thm:intuitionistic_lindenbaum_tarski_algebra/bottom} That \( [\bot] \) is the bottom is a restatement of \eqref{eq:thm:minimal_propositional_negation_laws/efq}.

  \SubProofOf{thm:intuitionistic_lindenbaum_tarski_algebra/conditional} Restating \eqref{eq:def:heyting_algebra/conditional}, we must prove that \( [\psi_1 \rightarrow \psi_2] \) equals
  \begin{equation*}
    ([\psi_1] \rightarrow [\psi_2]) = \underbrace{\sup\set[\Big]{ [\varphi] \given* \varphi \in \Gamma \T{and} (\varphi \wedge \psi_1) \vdash \psi_2 }}_{\Phi}.
  \end{equation*}

  An equivalent condition for \( \varphi \) to be in \( \Phi \) is, due to \fullref{thm:conjunction_of_premises},
  \begin{equation*}
    \varphi, \psi_1 \vdash \psi_2.
  \end{equation*}

  Then by the deduction theorem,
  \begin{equation*}
    \varphi \vdash (\psi_1 \rightarrow \psi_2).
  \end{equation*}

  Thus, \( [\psi_1 \rightarrow \psi_2] \) is an upper bound of \( \Psi \).

  It remains to show that \( (\psi_1 \rightarrow \psi_2) \in \Psi \). Both \( \psi_1 \) and \( (\psi_1 \rightarrow \psi_2) \) follow from the formula \( \parens[\Big]{ (\psi_1 \rightarrow \psi_2) \wedge \psi_1 } \) and by applying \eqref{eq:def:def:axiomatic_deductive_system/mp}, we obtain
  \begin{equation*}
    \psi_1, (\psi_1 \rightarrow \psi_2) \vdash \psi_2.
  \end{equation*}

  From \fullref{thm:conjunction_of_premises},
  \begin{equation*}
    \parens[\Big]{ (\psi_1 \rightarrow \psi_2) \wedge \psi_1 } \vdash \psi_2.
  \end{equation*}

  Therefore, \( [\psi_1 \rightarrow \psi_2] \in \Phi \) and it is indeed the supremum of \( \Psi \).

  \SubProofOf{thm:intuitionistic_lindenbaum_tarski_algebra/complement} The pseudocomplement \( \widetilde{[\psi]} \) is, by definition,
  \begin{equation*}
    \widetilde{[\psi]}
    =
    [\psi] \rightarrow [\bot].
  \end{equation*}

  From what we have already proved, we can conclude that \( \widetilde{[\psi]} = [\psi \rightarrow \bot] \). From \fullref{def:minimal_propositional_axiomatic_deductive_system/negation} it follows that the formulas \( \psi \rightarrow \bot \) and \( \neg \psi \) are interderivable, thus \( \widetilde{[\psi]} = [\neg \psi] \).

  If we are working in classical logic where \eqref{eq:thm:minimal_propositional_negation_laws/lem} holds, then
  \begin{equation*}
    \sup\set{ [\psi], \widetilde{[\psi]} }
    \reloset {\ref{thm:intuitionistic_lindenbaum_tarski_algebra/join}} =
    [\psi \vee \neg \psi]
    \reloset {\eqref{eq:thm:minimal_propositional_negation_laws/lem}} =
    [\top],
  \end{equation*}
  which proves \eqref{def:bounded_lattice_complement/join}.

  The dual law \eqref{def:bounded_lattice_complement/meet}
  \begin{equation*}
    \inf\set{ [\psi], \widetilde{[\psi]} }
    \reloset {\ref{thm:intuitionistic_lindenbaum_tarski_algebra/meet}} =
    [\psi \wedge \neg \psi]
    \reloset {\eqref{eq:thm:minimal_propositional_negation_laws/lnc}} =
    [\neg \top]
    =
    \widetilde{[\top]}
    \reloset {\eqref{eq:def:heyting_algebra/pseudocomplement}} =
    [\bot].
  \end{equation*}
\end{proof}

\begin{remark}\label{rem:thm:intuitionistic_lindenbaum_tarski_algebra/syntactic_proof}
  Notice that the proof of \fullref{thm:intuitionistic_lindenbaum_tarski_algebra} relies entirely on the derivation system. On the other hand, we define \hyperref[def:propositional_heyting_algebra_semantics]{Heyting semantics} for intuitionistic formulas.

  Thus, we have shown that Heyting algebras arise naturally from the intuitionistic natural deduction system and that their role as a semantic framework is completely justified.
\end{remark}
