\subsection{Preorders}\label{subsec:preorders}

\begin{definition}\label{def:preordered_set}\mcite{nLab:preorder}
  A \hyperref[def:binary_relation]{binary relation} \( \leq \) on a set \( \mscrP \) is called a \term{preorder} if it is \hyperref[def:binary_relation/reflexive]{reflexive} and \hyperref[def:binary_relation/transitive]{transitive}.

  A \term{preordered set} is a set \( \mscrP \) equipped with a preorder \( \leq \).

  \begin{thmenum}[series=def:preordered_set]
    \thmitem{def:preordered_set/theory} Consider a \hyperref[def:first_order_language]{first-order language} \( \mscrL \) with a single \hyperref[rem:order_infix_notation]{infix} binary predicate symbol \( \leq \). The \hyperref[def:first_order_semantics/theory]{theory} in \( \mscrL \) axiomatized by \hyperref[def:binary_relation/reflexive]{reflexivity} and \hyperref[def:binary_relation/transitive]{transitivity} for \( \leq \) is called the \term{theory of preordered sets}.

    \thmitem{def:preordered_set/homomorphism} A \hyperref[def:first_order_homomorphism]{homomorphism} between the preordered sets \( (\mscrP, \leq) \) and \( (\mscrP, \leq) \) is, explicitly, a function \( \varphi: \mscrP \to \mscrQ \) such that
    \begin{equation}\label{eq:def:preordered_set/homomorphism}
      x \leq y \T{implies} \varphi(x) \leq \varphi(y)
    \end{equation}

    If
    \begin{equation*}
      x \neq y \T{implies} \varphi(x) \neq \varphi(y),
    \end{equation*}
    we call the function \( \varphi \) \term{strictly monotone}.

    In particular, if \( \mscrP \) is the set of \hyperref[rem:peano_arithmetic_zero/nonnegative]{nonnegative integers}, then we speak of \term{monotone sequences}
    \begin{equation*}
      \{ x_k \}_{k=1}^\infty,
    \end{equation*}
    where \( x_{k-1} \leq_Q x_k \) for all \( k = 1, 2, 3, \ldots \).

    \thmitem{def:preordered_set/substructure} Any subset \( A \) of a preordered set \( (X, \leq) \) becomes a preordered set with the induced preorder \( \leq_A \) defined as the restriction of \( \leq \) to only elements of \( A \).

    \thmitem{def:preordered_set/category} We give no special name for the \hyperref[def:category_of_first_order_models]{category of models} for the theory of preordered sets.

    \thmitem{def:preordered_set/dual} The partially ordered set \( (\mscrP, \geq) \) where \( \geq \) is the \hyperref[def:binary_relation/converse]{converse relation}, is the called the \term{dual preordered set}. See \fullref{def:thin_category} for a discussion of the duality.
  \end{thmenum}

  We introduce the following terminology for elements of \( (\mscrP, \leq) \):
  \begin{thmenum}[resume=def:preordered_set]
    \thmitem{def:preordered_set/comparability} The elements \( x, y \in \mscrP \) are called \term{comparable} if either \( x \leq y \) or \( y \leq x \).

    \thmitem{def:preordered_set/upper_and_lower_bounds}\mcite[170]{Enderton1977Sets} The member \( x \in \mscrP \) is an \term{upper bound} for \( A \subseteq \mscrP \) if \( y \leq x \) for any \( y \in A \). Dually, \( x \) is a \term{lower bound} of \( A \) if it is an upper bound in the dual preordered set \( (\mscrP, \geq) \).

    \thmitem{def:preordered_set/bounded_set} The set \( A \subseteq \mscrP \) is called \term{bounded from above} (resp. \term{bounded from below}) if it has an upper bound (resp. lower bound). If a set is bounded from both directions, we simply say that it is \term{bounded}.

    \thmitem{def:preordered_set/maximal_and_minimal_element}\mcite[170]{Enderton1977Sets} The member \( x \in \mscrP \) is a \term{maximal element} of \( A \subseteq \mscrP \) (resp. \term{minimal element} of \( A \) in \( (\mscrP, \geq) \)) if \( x \leq y \) implies \( x = y \) for any \( y \in A \).

    \thmitem{def:preordered_set/maximum_and_minimum}\mcite[171]{Enderton1977Sets} The member \( x \in \mscrP \) is a \term{largest element} or \term{maximum} \( \max A \) of \( A \subseteq \mscrP \) (resp. \term{smallest element} or \term{minimum} \( \min A \) of \( A \) in \( (\mscrP, \geq) \)) if \( y \leq x \) for any \( y \in A \). While there may exist distinct maximal elements in \( A \), there exists at most one maximum.

    \thmitem{def:preordered_set/supremum_and_infimum}\mcite[170]{Enderton1977Sets} The member \( x \in \mscrP \) is a \term{supremum} (resp. \term{infimum}) for \( A \subseteq \mscrP \) if \( x \) is the least upper bound (resp. greatest lower bound) of \( A \), i.e. the smallest element of the set consisting of all upper bounds of \( \mscrP \). Suprema and infima may exist when neither maximal nor largest elements exist.
  \end{thmenum}
\end{definition}

\begin{definition}\label{def:thin_category}\mcite{nLab:thin_category}
  A \hyperref[def:category]{category} \( \cat{P} \) is called a \term{thin category} if, for every two objects \( A, B \in \boldop{P} \), whenever \( f, g \in \boldop{P}(A, B) \), we have \( f = g \).

  If \( \cat{P} \) is locally small, this is equivalent to saying that any set of morphisms \( \boldop{P}(A, B) \) is at most a singleton.
\end{definition}

\begin{proposition}\label{thm:preorder_category_correspondence}
  To every \hyperref[def:preordered_set]{preordered set} there corresponds exactly one \hyperref[def:category_cardinality]{small} \hyperref[def:thin_category]{thin} category.

  Furthermore, \hyperref[def:preordered_set/supremum_and_infimum]{infima} correspond to categorical \hyperref[def:categorical_product]{products}, suprema to \hyperref[def:categorical_coproduct]{coproducts} and dual \hyperref[def:preordered_set/dual]{preordered sets} correspond to dual \hyperref[def:opposite_category]{categories}.

  Compare this result to \fullref{thm:partial_order_category_correspondence}.
\end{proposition}
\begin{proof}
  \SufficiencySubProof Let \( (P, \leq) \) be a preordered set. We define the category \( \cat{P} \) as follows:
  \begin{refenum}
    \refitem{def:category/C1} The set of objects in \( \cat{P} \) is the set \( P \).
    \refitem{def:category/C2} The set of morphisms between two elements \( x, y \in \cat{P} \) is the singleton \( \{ (x, y) \} \) when \( x \leq y \) and the empty set otherwise.
    \refitem{def:category/C3} The composition of two morphisms \( (x, y) \) and \( (y, z) \) is simply \( (x, z) \) (such a morphism exists by transitivity of \( \leq \)).

    The axiom \ref{def:category/identity} follows from reflexivity of \( \leq \) and the axiom \ref{def:category/associativity} is trivial.
  \end{refenum}

  We showed that \( \cat{P} \) is indeed a category. We will only prove the equivalence of products and infima since the argument for suprema and coproducts is completely analogous.

  Let \( p \) be the categorical product of the set \( A \subseteq P \). Then \( p \leq x \) for all \( x \in A \), hence it is a lower bound. If \( q \) is another lower bound, then by definition of product, there exists a unique morphism \( q \leq p \). Therefore \( p = q \) is the infimum.

  \NecessitySubProof Now assume that \( \cat{P} \) is a thin small category. Define the relation \( \leq \) on the set \( \cat{P} \) as \( x \leq y \iff \cat{P}(x, y) \neq \varnothing \). This is indeed a preorder because
  \begin{itemize}
    \item \( \leq \) is \hyperref[def:binary_relation/reflexive]{reflexivity} because of the existence of identity morphisms.
    \item \( \leq \) is \hyperref[def:binary_relation/transitive]{transitive} since if \( x \leq y \) and \( y \leq z \), composition of morphisms gives us \( x \leq z \).
  \end{itemize}

  Note that the infimum of a set \( A \subseteq \cat{P} \) (if it exists) has a unique morphism \( \inf A \) such that \( \inf A \leq x \) for any \( x \in A \). If \( y \leq x \) for all \( x \in A \) is another \hyperref[def:categorical_cone]{cone}, then necessarily \( y \leq \inf A \). Therefore the infimum is the categorical product.

  We proved that for each partially ordered set there corresponds at least one thin small category and vice versa. The fact that to each poset corresponds at most one category \( \cat{P} \) is obvious. Therefore we have a correspondence between the two.

  Duality is also obvious from our constructions.
\end{proof}

\begin{definition}\label{def:directed_set}\mcite[8]{Engelking1989}
  A \hyperref[def:preordered_set]{preordered set} \( (\mscrP, \leq) \) is called a \term{directed set} if every finite subset of \( \mscrP \) has an upper \hyperref[def:preordered_set/upper_and_lower_bounds]{bound}, i.e. for all \( x, y \in \mscrP \) there exists \( z \in \mscrP \) such that \( x \leq z \) and \( y \leq z \).

  There is no established name for the relation itself.

  Note that \( \{ x, y \} \) may not have a supremum, i.e. the set of its upper bounds may not have a \hyperref[def:preordered_set/maximum_and_minimum]{smallest element}. Thus the upper bound condition is strictly weaker than every two-element set having a supremum.

  Directed sets are used to define nets in topological spaces, see \fullref{def:topological_net}.
\end{definition}
