\subsection{Rings}\label{subsec:rings}

We will start by defining semirings, and to do that we will first motivate distributivity.

\begin{proposition}\label{thm:monoid_distributivity}
  Fix a \hyperref[def:magma/commutative]{commutative} \hyperref[def:monoid]{monoid} \( (R, +, \cdot) \), where \( +: R \times R \to R \) is the monoid operation and \( \cdot: \BbbN \times R \to R \) is \hyperref[rem:additive_magma/multiplication]{recursively defined multiplication}.

  We have the following property, which we call \term{distributivity} of \( \cdot \) over \( + \):
  \begin{equation}\label{eq:thm:monoid_distributivity}
    n \cdot (x + y) = n \cdot x + n \cdot y.
  \end{equation}
\end{proposition}
\begin{proof}
  We use induction on \( n \). The case \( n = 0 \) is trivial. Suppose that \eqref{eq:thm:monoid_distributivity} holds. Then
  \begin{equation*}
    (n + 1) \cdot (x + y)
    \reloset {\eqref{eq:def:magma/exponentiation}} =
    n \cdot (x + y) + (x + y)
    \reloset {\T{ind.}} =
    n \cdot x + n \cdot y + (x + y)
    \reloset {\eqref{eq:def:magma/exponentiation}} =
    (n + 1) \cdot x + (n + 1) \cdot y.
  \end{equation*}
\end{proof}

\begin{definition}\label{def:semiring}\mcite[ch. 3, sec. 2.1]{GondranMinoux1984Graphs}
  A \term{semiring} is a \hyperref[def:magma/commutative]{commutative} \hyperref[def:monoid]{monoid} \( (R, +) \) with a second \hyperref[def:magma/associative]{associative} operation \( \cdot: R \times R \to R \) called \term{multiplication}, which extends multiplication with natural numbers. The precise compatibility axioms are listed in \fullref{def:semiring/theory}.

  Although not strictly necessary, it will be convenient for us to assume that multiplication has an identity. If a multiplicative identity does not exist, we call \( (R, +, \cdot) \) a \term{nonunital semiring}. We will not use nonunital semirings, but it is important to acknowledge their existence. In this context, if an identity exists, we will sometimes call \( (R, + \cdot) \) a \term{unital semiring}.

  We call \( (R, +) \) the \term{additive monoid} and \( (R, \cdot) \) the \term{multiplicative monoid} of the semiring.

  Semirings have the following metamathematical properties:
  \begin{thmenum}
    \thmitem{def:semiring/theory} The \hyperref[def:first_order_theory]{first-order theory} for semirings extends the \hyperref[def:monoid/theory]{theory of monoids}.

    First, we add another \hyperref[rem:first_order_formula_conventions/infix]{infix} binary functional symbol \( \cdot \) and a constant \( 1 \). This notation is justified by \fullref{thm:semiring_characteristic_homomorphism}.

    We then extend the theory of monoids with \hyperref[def:magma/commutative]{commutativity} for \( + \), \hyperref[def:magma/associative]{associativity} for \( \cdot \), and the following axioms:
    \begin{thmenum}
      \thmitem{def:semiring/left_distributivity} Multiplication on the left distributes over addition:
      \begin{equation}\label{eq:def:semiring/left_distributivity}
        \xi \cdot (\eta + \zeta) \doteq \xi \cdot \eta + \xi \cdot \zeta.
      \end{equation}

      \thmitem{def:semiring/right_distributivity} Multiplication on the right also distributes over addition:
      \begin{equation}\label{eq:def:semiring/right_distributivity}
        (\xi + \eta) \cdot \zeta \doteq \xi \cdot \zeta + \eta \cdot \zeta.
      \end{equation}

      If multiplication is commutative, right distributivity follows from left distributivity.

      \thmitem{def:semiring/absorption} Zero is an absorbing element:
      \begin{equation}\label{eq:def:semiring/absorption}
        \xi \cdot 0 \doteq 0 \wedge 0 \cdot \xi \doteq 0.
      \end{equation}

      \thmitem{def:semiring/identity} We also restate the identity axiom \eqref{eq:def:monoid/theory/identity} for the multiplicative unit \( 1 \) to highlight its connection with \eqref{eq:def:semiring/absorption}:
      \begin{equation}\label{eq:def:semiring/identity}
        \xi \cdot 1 \doteq \xi \wedge 1 \cdot \xi \doteq \xi.
      \end{equation}
    \end{thmenum}

    \thmitem{def:semiring/homomorphism} A \hyperref[def:first_order_homomorphism]{first-order homomorphism} between the semirings \( R \) and \( S \) is a function \( \varphi: R \to S \) that is a \hyperref[def:monoid/homomorphism]{monoid homomorphism} both for their additive monoids also for their multiplicative monoids.

    For semirings, \( \varphi \) must be a monoid homomorphism for the multiplicative monoids also. That is, it must preserve the multiplicative identity \( 1 \).

    \thmitem{def:semiring/submodel} The set \( S \subseteq R \) is a \hyperref[thm:substructure_is_model]{submodel} of \( R \) if it is a both \hyperref[def:monoid/submodel]{submonoid} of the additive monoid and also of the multiplicative monoid.

    As a consequence of \fullref{thm:positive_formulas_preserved_under_homomorphism}, the \hyperref[def:multi_valued_function/image]{image} of a group homomorphism \( \varphi: R \to S \) is a subgroup of \( S \).

    For an arbitrary set \( S \), we denote the \hyperref[def:first_order_generated_substructure]{generated submodel} by \( \braket{ S } \).

    \thmitem{def:semiring/trivial} The \hyperref[thm:substructures_form_complete_lattice/bottom]{trivial} semiring is the \hyperref[def:pointed_set/trivial]{trivial pointed set} \( \set{ 0 } \).

    See \fullref{ex:trivial_semiring} for some properties of the trivial semiring.

    \thmitem{def:semiring/category} The \hyperref[def:category_of_small_first_order_models]{category of \( \mscrU \)-small models} for semirings does not have an established name. It is \hyperref[def:monoid]{\( \ucat{CMon} \)}-\hyperref[def:concrete_category]{concrete} with the forgetful functor taking the additive monoids.

    \thmitem{def:semiring/exponentiation} As we shall see in \fullref{thm:semiring_characteristic_homomorphism}, multiplication in \( \cdot \) extends multiplication by natural numbers in the abelian group \( (R, +) \). We do have a third operation, however --- \hyperref[def:monoid/exponentiation]{monoid exponentiation} in \( (R, \cdot) \).

    For any integer \( n \), we have the fundamental property \( 1^n = 1 \).

    \thmitem{def:semiring/commutative} If multiplication is commutative, we call the semiring itself \term{commutative}.
  \end{thmenum}
\end{definition}

\begin{example}\label{ex:trivial_semiring}
  A \hyperref[def:semiring/homomorphism]{semiring} is trivial if and only if \( 0_R = 1_R \). This follows from \eqref{eq:def:semiring/absorption} and \eqref{eq:def:semiring/identity}.

  As a consequence, if \( \varphi: \set{ 0 } \to R \) is a \hyperref[def:semiring/homomorphism]{semiring} homomorphism, \( R \) is a trivial semiring. This is further strengthened by \fullref{thm:semiring_embedding_preserves_characterstic}.
\end{example}

\begin{proposition}\label{thm:semiring_characteristic_homomorphism}
  For every \hyperref[def:semiring/identity]{semiring}, multiplication extends the abelian group multiplication.

  More precisely, denote the additive identity by \( 0_R \) and the multiplicative identity by \( 1_R \). Define the following semiring homomorphism:
  \begin{equation}\label{eq:thm:semiring_characteristic_homomorphism}
    \begin{aligned}
      &\iota: \BbbN \to R \\
      &\iota(n) \coloneqq \begin{cases}
        0_R                &n = 0, \\
        \iota(n - 1) + 1_R &n > 0.
      \end{cases}
    \end{aligned}
  \end{equation}

  This is the unique homomorphism from \( \BbbN \) to \( R \). Furthermore, we have the following analogue to \eqref{eq:def:magma/exponentiation}:
  \begin{equation}\label{eq:thm:semiring_characteristic_homomorphism/multiplication}
    \iota(n) \cdot x \coloneqq \begin{cases}
      0_R,                      &n = 0, \\
      \iota(n - 1) \cdot x + x, &n > 1.
    \end{cases}
  \end{equation}
\end{proposition}
\begin{proof}
  First note that \eqref{eq:thm:semiring_characteristic_homomorphism/multiplication} follows from \eqref{eq:thm:semiring_characteristic_homomorphism} via \hyperref[def:semiring/right_distributivity]{right distributivity}.

  It remains to show that \( \iota \) is a monoid homomorphism, and that it is unique. Clearly \( \iota(0) = 0_R \) and \( \iota(1) = 1_R \). Proving \( \iota(n + m) = \iota(n) + \iota(m) \) and \( \iota(nm) = \iota(n) \cdot \iota(m) \) can be done via nested induction.

  Now suppose \( \varphi: \BbbN \to R \) is a homomorphism. It is clear that \( \varphi(0) = 0_R \) and \( \varphi(1) = 1_R \), and also
  \begin{equation*}
    \varphi(n + 1) = \varphi(n) + \varphi(1) = \varphi(n) + 1_R.
  \end{equation*}

  This implies \( \iota = \varphi \).
\end{proof}

\begin{definition}\label{def:semiring_characteristic}\mimprovised
  We define the \term{characteristic} \( \op{char}(R) \) of a semiring \( R \) via the following equivalent definitions:
  \begin{thmenum}
    \thmitem{def:semiring_characteristic/homomorphism} \( \op{char}(R) \) is the nonnegative integer \( n \) such that
    \begin{equation*}
      n \BbbZ = \ker\iota,
    \end{equation*}
    where \( \iota \) is the homomorphism from the natural numbers defined via \eqref{eq:thm:semiring_characteristic_homomorphism}.

    \thmitem{def:semiring_characteristic/direct} \( \op{char}(R) \) is the smallest positive integer \( n \) such that \( n \cdot 1_R = 0_R \) and \( \op{char}(R) = 0 \) if \( 0_R \) cannot be obtained in this way.
  \end{thmenum}
\end{definition}
\begin{defproof}
  \EquivalenceSubProof{def:semiring_characteristic/homomorphism}{def:semiring_characteristic/direct} Let \( n \) be such that
  \begin{equation*}
    n \BbbZ = \ker\iota.
  \end{equation*}

  In particular, \( \iota(0) = \iota(n) \).

  If \( n = 0 \), \( \ker\iota \) is a trivial group and \( \iota \) is an embedding. Then there cannot exist a positive integer \( n \) such that
  \begin{equation*}
    n \cdot 1_R = 0_R.
  \end{equation*}

  Otherwise, \( n \) is the smallest positive integer such that
  \begin{equation*}
    n \cdot 1_R = 0 \cdot 1_R = 0_R.
  \end{equation*}
\end{defproof}

\begin{proposition}\label{thm:semiring_embedding_preserves_characterstic}
  If \( \varphi: R \to T \) is a \hyperref[def:semiring/homomorphism]{semiring embedding}, then the \hyperref[def:semiring_characteristic]{characteristic} of \( R \) is equal to that of \( T \).
\end{proposition}
\begin{proof}
  First suppose that \( R \) has positive characteristic \( n \). Then \( n \cdot 1_R = 0_R \), which implies \( n \cdot \varphi(1_R) = \varphi(0_R) \), hence \( \op{char}(T) \leq n \). But \( \varphi \) is an embedding, hence if \( k \cdot 1_R \neq 0_R \), then
  \begin{equation*}
    k \cdot \varphi(1_R) \neq \varphi(0_R).
  \end{equation*}

  This implies that \( \op{char}(T) \geq \op{char}(R) \), which in turn shows that \( \op{char}(T) = \op{char}(R) \).

  If \( R \) has characteristic zero, then \( \iota: \BbbN \to R \) is an embedding and thus \( \varphi \bincirc \iota: \BbbN \to T \) is also an embedding. It is unique as shown in \fullref{thm:semiring_characteristic_homomorphism}. Therefore, \( T \) also has characteristic zero.
\end{proof}

\begin{example}\label{ex:semiring_characteristic}
  The following are examples of \hyperref[def:semiring_characteristic]{semiring characteristics}:

  \begin{thmenum}
    \thmitem{ex:semiring_characteristic/nonnegative_integers} The zero-based \hyperref[def:set_of_natural_numbers]{natural numbers} \( \BbbN \) have characteristic \( \op{char}(\BbbN) = 0 \) because \( \iota \) is an isomorphism. Consequently, by \fullref{thm:semiring_embedding_preserves_characterstic}, any supersemiring of \( \BbbN \) has characteristic zero, most notably the ring of integers \( \BbbZ \) and the fields \( \BbbQ \), \( \BbbR \), \( \BbbC \).

    \thmitem{ex:semiring_characteristic/integers_modulo} The ring \hyperref[def:ring_of_integers_modulo]{\( \BbbZ_n \)} of integers modulo \( n \) has characteristic \( \op{char}(\BbbZ_n) = n \) because of \fullref{thm:integers_modulo_isomorphic_to_quotient_group}.

    \thmitem{ex:semiring_characteristic/polynomial_ring} An \hyperref[def:algebra_over_ring]{algebra} \( A \) over a nontrivial commutative unital ring \( R \) has characteristic \( \op{char}(A) = R \) because of the canonical embedding of \( R \) in \( A \). In particular, polynomial \hyperref[def:algebra_of_polynomials]{rings} \( R[X] \) have the same characteristic as their ring.

    \thmitem{ex:semiring_characteristic/galois_fields} The \hyperref[thm:galois_field_existence]{Galois field} \( \BbbF_{p^n} \) has characteristic \( p \) because it is a field extension of \( \BbbF_p \).
  \end{thmenum}
\end{example}

\begin{proposition}\label{thm:category_of_semirings_properties}
  The \hyperref[def:semiring/category]{category of semirings} has the following properties:
  \begin{thmenum}
    \thmitem{thm:category_of_semirings_properties/initial} The \hyperref[def:set_of_integers]{ring of integers} \( \BbbZ \) is an \hyperref[def:universal_objects/initial]{initial object}.

    \thmitem{thm:category_of_semirings_properties/terminal} The trivial semiring \( \set{ 0 } \) is an \hyperref[def:universal_objects/terminal]{terminal object}.
  \end{thmenum}
\end{proposition}
\begin{proof}
  \SubProofOf{thm:category_of_semirings_properties/initial} Follows from \fullref{thm:semiring_characteristic_homomorphism}.
  \SubProofOf{thm:category_of_semirings_properties/terminal} Follows from \fullref{ex:trivial_semiring}.
\end{proof}

\begin{definition}\label{def:semiring_division}
  In a \hyperref[def:semiring]{semiring} \( R \), we say that \( x \) \term{divides} \( y \) or \( x \) is a \term{divisor} of \( y \) if there exists \( z \) such that \( xz = y \). We write \( x \mid y \).

  In particular, we call divisors of \( 1 \) \term{units}. If \( xy = 0 \) and both \( x \) and \( y \) are nonzero, we call them \term{zero divisors}.
\end{definition}

\begin{proposition}\label{thm:semiring_cancellative_iff_no_zero_divisors}
  An element of a commutative ring is a \hyperref[def:semiring_division]{zero divisor} if and only if it is not cancellable. That is, if \( xy = xz \) does not imply \( y = z \).
\end{proposition}
\begin{proof}
  \SufficiencySubProof Suppose that \( x \) is a zero divisor and let \( y \) be such that \( xy = 0 \). For any element \( z \), we have
  \begin{equation*}
    xy = 0 = x(yz).
  \end{equation*}

  But \( y \neq yz \) unless \( z = 1 \). Thus \( x \) is not cancellable.

  \NecessitySubProof Suppose that \( x \neq 0 \) is not cancellable and that \( xy = 0 \) for some nonzero \( y \). Then \( xy = x0 \), which implies \( y = 0 \). But this contradicts out choice of \( y \).

  Thus \( x \) cannot be simultaneously cancellable and a zero divisor.
\end{proof}

\begin{definition}
  The following are special kinds of semirings:
  \begin{thmenum}[resume=def:semiring]
    \thmitem{def:semiring/no_zero_divisor} If \( xy = 0 \) whenever \( x \) and \( y \) are nonzero, we say that the semiring has \term{no zero divisors}.

    \thmitem{def:semiring/ring} \term{Rings} are semirings with invertible addition, i.e. \( (R, +) \) forms an \hyperref[def:abelian_group]{abelian group}. The category of rings is denoted by \( \cat{Ring} \).

    \thmitem{def:semiring/unital_ring} \term{Unital rings} are \hyperref[def:semiring/ring]{rings} in which multiplication is unital, i.e. \( R, \cdot \) is a \hyperref[def:monoid]{monoid} with identity \( 1 \). Some authors define all rings to be unital. Invertible elements under multiplication are called \term{units} (see \fullref{rem:units_in_rings_etymology}) and the operation itself is called \term{division}.

    We may additionally require that the ring is nontrivial, so that \( 0 \neq 1 \) (see \fullref{ex:trivial_semiring}).

    \thmitem{def:semiring/commutative_ring} \term{Commutative rings} are \hyperref[def:semiring/ring]{rings} in which multiplication is \hyperref[def:magma/commutative]{commutative}, i.e. \( R, \cdot \) is a commutative semigroup.

    \thmitem{def:semiring/commutative_unital_ring} \term{Commutative unital rings} are (obviously) both commutative and unital and as usually assumed to be nontrivial. Despite being ubiquitous, they do not have an established one-word name.

    We are mostly interested in these rings because if \( R \) is a commutative unital ring, so is its polynomial \hyperref[def:algebra_of_polynomials]{ring} \( R[X] \) and, by induction, multivariate polynomial \hyperref[def:multivariate_polynomial]{rings} (\( R[X, Y] = R[X][Y] \)). So we are interested in properties of \( R \) that are preserves by \( R[X] \). This is the reason, so many types of commutative rings and ideals are studied. We are interested in integral domains when we speak of divisibility and factorization and since zero is absorbing, it does not interact nicely with factorization. Refer to \fullref{sec:commutative_algebra}.

    \thmitem{def:semiring/integral_domain} \term{Integral domains} are nontrivial commutative unital \hyperref[def:semiring/commutative_unital_ring]{rings} with no zero \hyperref[def:commutative_ring_division]{divisors}. This implies that \( 1 \neq 0 \) because otherwise \( 1 \cdot 1 = 0 \) and \( 1 \) would be a zero divisor. This in turn implies that \fullref{ex:trivial_semiring}.

    \Fullref{thm:semiring_cancellative_iff_no_zero_divisors} shows that a commutative unital ring \( R \) is an integral domain if and only if its multiplication is cancellable.

    \thmitem{def:semiring/unique_factorization_domain} \term{Unique factorization domains} are integral domains in which every element has unique \hyperref[def:factorization_in_ring]{factorization} exists.

    \thmitem{def:semiring/principal_ideal_domain} \term{Principal ideal domains} are integral \hyperref[def:semiring/integral_domain]{domains} in which every \hyperref[def:semiring_ideal]{ideal} is \hyperref[def:principal_ideal]{principal}.

    By \fullref{thm:pid_is_ufd}, every principal ideal domain is a unique factorization domain.

    \thmitem{def:semiring/euclidean_domain} \term{Euclidean domains} are integral \hyperref[def:semiring/integral_domain]{domains} which allow division with remainders (see \fullref{def:euclidean_domain}).

    By \fullref{thm:euclidean_domain_is_pid}, every Euclidean domain is a principal ideal domain.

    \thmitem{def:semiring/division_ring} \term{Division rings} are nontrivial unital \hyperref[def:semiring/unital_ring]{rings} in which all nonzero elements are units, i.e. \( (F \setminus \{ 0 \}, \cdot) \) is a \hyperref[def:group]{group}. The nontriviality is a requirement because we want \( 1 \) not to be a zero divisor (see discussion in \fullref{def:semiring/integral_domain}).

    The multiplicative inverse of an element in a division ring is called its \term{reciprocal}.

    In order to fit multiplicative invertibility as an axiom, we can use the following formula:
    \begin{equation*}
      \forall \xi ((\xi \doteq 0) \lor \exists \eta (\xi \cdot \eta \doteq 1))
    \end{equation*}
    or add an additional operation \( (\cdot)^{-1} \) that inverts all nonzero elements and fixes zero, that is,
    \begin{equation*}
      \forall \xi (\xi \cdot \xi^{-1} \doteq 1),
    \end{equation*}
    where we define \( 0^{-1} = 0 \). This is only a formalism since \( 0 \) is not actually \enquote{invertible}, but it is required if we wish to avoid existential quantifiers.

    \thmitem{def:semiring/field} \term{Fields} are \hyperref[def:magma/commutative]{commutative} division \hyperref[def:semiring/division_ring]{rings}, i.e. \( (F \setminus \{ 0 \}, \cdot) \) is an \hyperref[def:abelian_group]{abelian group}. The category of fields is denoted by \( \cat{Field} \).
  \end{thmenum}
\end{definition}

\begin{example}\label{ex:semirings}
  We give examples and counterexamples of semirings. Note that the order of definitions in \fullref{def:semiring} is not preserved.

  \begin{refenum}
    \refitem{def:semiring/euclidean_domain} The base building block for the examples will be the ring \( (\BbbZ, +, \cdot) \) of \hyperref[def:set_of_integers]{integers}, which itself is an Euclidean domain (see \fullref{def:set_of_integers}).

    Another example of Euclidean domains are the polynomial rings over a field (see \fullref{thm:polynomials_over_field_are_euclidean_domain}).

    \refitem{def:semiring/identity} By removing additive inverses from the integers, we obtain the semiring \( (\BbbN, +, \cdot) \) of \hyperref[def:set_of_natural_numbers]{natural numbers}.

    Take only the commutative monoid \( (\BbbN, +) \) of \hyperref[def:set_of_natural_numbers]{natural numbers} with addition. The endomorphism semiring
    \begin{equation*}
      \End(\BbbN)
    \end{equation*}
    is a noncommutative semiring.

    Tropical \hyperref[def:tropical_semiring]{semirings} are another example noncommutative semirings.

    \refitem{def:semiring} Simple examples of semirings (but not rings) without unity are proper semiring ideals. For example, the ideal \( 2\BbbN \) of positive even numbers is a non-unital semiring.

    Another example can be given by taking a subsemiring, but not a unital subsemiring of the endomorphism semiring \( \End(\BbbN, \oplus) \), for example the functions
    \begin{balign*}
       & f_n: \BbbN \to \BbbN        \\
       & f_n(x) \coloneqq x + n,
    \end{balign*}
    where \( n > 0 \). They are closed under composition and thus form a semiring themselves, but they do not contain the identity function, so the semiring is not unital.

    \refitem{def:semiring/unital_ring} The endomorphism rings \( \End(G) \) for any abelian group \( G \) are unital, but non-commutative rings. This includes the matrix space \( R^{n \times n} \) (see \fullref{thm:finite_dimensional_operators_are_isomorphic_to_matrices}).

    \refitem{def:semiring/ring} Consider the Banach space \( C_0(\BbbC) \) of complex functions vanishing at \hyperref[def:function_spaces/c0]{infinity}. If we take addition to be pointwise addition and multiplication to be composition, then \( C_0(\co) \) becomes a non-commutative ring with no multiplicative identity because \( C_0(\co) \) does not contain the identity function.

    \refitem{def:semiring/commutative_ring} For an example of a commutative ring without unit, consider again the Banach space \( C_0(\co) \), however define multiplication as pointwise function multiplication rather than by composition. The constant function \( f(x) = 1 \) does not vanish at infinity, hence \( (C_0(\co), +, \cdot) \) is a commutative, but not unital ring.

    \refitem{def:semiring/commutative_unital_ring} We are mostly interested in different types of commutative unital rings since the polynomials over well-behaved commutative rings preserve this behavior. This is important because multivariate \hyperref[def:multivariate_polynomial]{polynomials} are defined inductively as polynomials over polynomial rings. See \fullref{thm:geometric_nullstellensatz} for an application.

    An example of a nontrivial commutative unital ring that has zero divisors is the matrix algebra \( \BbbZ^{n \times n} \) over the integers. It is a ring under addition and matrix multiplication. We have
    \begin{equation*}
      \begin{pmatrix}
        1 & 0 \\
        1 & 0
      \end{pmatrix}
      \begin{pmatrix}
        0 & 0 \\
        0 & 1
      \end{pmatrix}
      =
      \begin{pmatrix}
        0 & 0 \\
        0 & 0
      \end{pmatrix},
    \end{equation*}
    thus there are zero divisors in \( \BbbZ^{n \times n} \).

    \refitem{def:semiring/integral_domain}\cite[388]{Knapp2016BasicAlgebra} The integral domain \( \BbbZ[\sqrt{-5}] \) is not a unique factorization domain because
    \begin{equation*}
      6 = (1 + \sqrt{-5}) (1 - \sqrt{-5}) = 2 \cdot 3.
    \end{equation*}

    Note that \( \BbbZ[\sqrt{-5}] \) is an integral domain by \fullref{thm:polynomials_over_integral_domain_are_integral_domain}.

    \refitem{def:semiring/unique_factorization_domain}\cite{ProofWiki:polynomials_in_integers_is_not_principal_ideal_domain} The unique factorization domain \( \BbbZ[X] \) is not a principal ideal domain.

    Note that \( \BbbZ[X] \) is a unique factorization domain by \fullref{thm:polynomials_over_integral_domain_are_integral_domain}.

    Consider the ideal \( I \) of polynomials with an even constant term.

    Assume that \( I \) is generated by the polynomial \( p(X) \in \BbbZ[X] \). Since \( 2 \in I \), then \( p(X) \) divides \( 2 \), so \( p(X) \in \{ -2, -1, 1, 2 \} \), that is \( p(X) \) is a unit of \( \BbbZ[X] \). But then \( I = \braket{p(X)} = \BbbZ[X] \), which contradicts the definition of \( I \).

    The obtained contradiction proves that \( \BbbZ[X] \) is not a principal ideal domain.

    \refitem{def:semiring/principal_ideal_domain} Principal ideal domains are not Euclidean domains in general. Such domains are discussed in \cite{Anderson1988}.

    \refitem{def:semiring/division_ring} The \hyperref[def:ring_localization]{localization} of a noncommutative ring over its nonzero cancellative elements (characterized by \fullref{thm:ring_localization_universal_property}), if it exists, forms a division ring.

    \refitem{def:semiring/field} The canonical examples of fields include the rational \hyperref[def:set_of_rational_numbers]{numbers} \( \BbbQ \), the \hyperref[def:set_of_real_numbers]{real numbers} \( \BbbR \) and the \hyperref[def:set_of_complex_numbers]{complex numbers} \( \BbbC \).

    More generally, any nontrivial commutative unital can be embedded in a field by \fullref{def:field_of_fractions}.
  \end{refenum}
\end{example}

\begin{definition}\label{def:semiring_kernel}
  The \term{kernel} \( \ker(f) \) of a semiring homomorphism \( f: R \to S \) is the \hyperref[def:zero_locus]{zero locus} of \( f \), that is, \hyperref[thm:def:function/properties/preimage]{preimage} \( f^{-1}(0_S) \).

  It is an instance of \fullref{def:zero_morphisms/kernel}.
\end{definition}

\begin{definition}\label{def:quotient_semiring}
  Let \( R \) be a ring and \( I \) be an ideal of \( M \). Define the \term{quotient ring} to be the quotient \hyperref[def:quotient_left_module]{module} when considering \( R \) as a module over itself.
\end{definition}

\begin{theorem}\label{thm:homomorphism_theorem_for_rings}
  Let \( \varphi: R \to T \) be a homomorphism of rings. We have the isomorphism
  \begin{equation*}
    R / \ker \varphi \cong \img \varphi.
  \end{equation*}
\end{theorem}
\begin{proof}
  Special case of \fullref{thm:homomorphism_theorem_for_left_modules}.
\end{proof}

\begin{definition}\label{def:tropical_semiring}\mcite{nLab:tropical_semiring}
  Fix a partially \hyperref[def:partially_ordered_set]{ordered} \hyperref[def:abelian_group]{abelian group} \( (M, +, \leq) \). Let \( \infty \) be a sentinel symbol not in \( M \). Define
  \begin{equation*}
    T \coloneqq M \cup \{ \infty \}
  \end{equation*}
  with operations
  \begin{balign*}
     & \oplus: T \times T \to T                        \\
     & x \oplus y \coloneqq \begin{cases}
      \min \{ x, y \}, & x \neq \infty \text{ and } y \neq \infty \text{ and they are comparable}, \\
      \infty,          & x = \infty \text{ or } y = \infty
    \end{cases} \\
    \\
     & \odot: T \times T \to T                         \\
     & x \odot y \begin{cases}
      x + y,  & x \neq \infty \text{ and } y \neq \infty, \\
      \infty, & x = \infty \text{ or } y = \infty
    \end{cases}
  \end{balign*}

  This makes \( (T, \oplus, \odot) \) into a \hyperref[def:semiring/identity]{semiring} with additive identity \( \infty \) and multiplicative identity \( 0 \). We call \( (T, \oplus, \odot) \) the \( \min \)-\term{tropical semiring} or simply the \term{tropical semiring} over \( M \). We define the \( \max \)-\term{tropical semiring} analogously by simply replacing \( \min \) with \( \max \).
\end{definition}

\begin{definition}\label{def:endomorphism_semiring}
  Let \( (X, +) \) be an monoid and let \( \End(X) \) be set of endomorphism over \( X \). We define two operations:
  \begin{itemize}
    \item Pointwise addition \( [f + g](x) \coloneqq f(x) + g(x) \).
    \item Multiplication by composition \( [fg](x) \coloneqq f(g(x)) \).
  \end{itemize}

  These operations make \( \End(X) \) into a semiring. If \( X \) is a group, then \( \End(X) \) is a ring.

  If \( X \) is a semiring, we define \( \End(X) \) to be a set of semiring endomorphisms (that is, we want the additive group homomorphisms to preserve multiplication and units). Then \( \End(X) \) is again a semiring and, if \( X \) is a unital ring, so is \( \End(X) \).
\end{definition}

\begin{definition}\label{def:ordered_semiring}
  Extending \fullref{def:preordered_magma} to (semi)rings, we define a \term{preordered semiring} to be a semiring \( \BbbR \) with a magma preorder \( \leq \) that additionally satisfies
  \begin{equation}\label{eq:def:ordered_semiring/nonnegativity}
    0 \leq y \T{and} 0 \leq y \T{implies} 0 \leq xy.
  \end{equation}
\end{definition}
