\subsection{Trees}\label{subsec:trees}

In this section, we will regard the edges of simple undirected graphs as sets of unordered tuples and edges of simple directed graphs as sets of ordered tuples.

\begin{definition}\label{def:tree}
  The possibly infinite \hyperref[def:undirected_multigraph]{simple undirected graph} \( T= (E, V, \mscrE) \) is called a \term{tree} if any of the following equivalent conditions hold:
  \begin{thmenum}
    \thmitem{def:tree/connected_acyclic} \( T \) is \hyperref[def:undirected_multigraph_connectedness]{connected} and \hyperref[def:undirected_multigraph_path/cycle]{acyclic}.
    \thmitem{def:tree/maximally_acyclic} \( T \) is \term{maximally acyclic}, meaning that adding an edge between existing vertices would create a \hyperref[def:undirected_multigraph_path/cycle]{cycle}.
    \thmitem{def:tree/minimally_connected} \( T \) is \term{minimally connected}, meaning that removing an edge will make the graph \hyperref[def:quiver_connectedness]{disconnected}.
    \thmitem{def:tree/single_path} For every pair of vertices in \( T \), there exists a unique \hyperref[def:quiver_path]{path} connecting them.

    This condition motivates the definition of \hyperref[def:arborescence]{arborescences}.
  \end{thmenum}

  It is conventional to use the term \term{node} for the vertices of a tree.
\end{definition}
\begin{proof}
  \ImplicationSubProof{def:tree/connected_acyclic}{def:tree/maximally_acyclic} Suppose that \( T \) is connected and acyclic. There is always an edge not in \( T \) because the \hyperref[ex:complete_graph]{complete graph} \( K_V \) contains cycles and \( T \) does not. Let \( \set{ u, v } \) be an edge not in \( T \) and let \( T' \) be the supergraph that adjoins only this new edge.

  Since \( T \) is connected, there exists a path \( p \) connecting \( u \) and \( v \). Appending the edge \( \set{ u, v } \) to this path creates a cycle in \( T' \).

  Therefore, \( T' \) is not acyclic.

  \ImplicationSubProof{def:tree/maximally_acyclic}{def:tree/connected_acyclic} Suppose that \( T \) is maximally acyclic. We will show that it is connected.

  Suppose that \( T \) is not connected. Then there exist vertices \( u \) and \( v \) with no path between them. Let \( T' \) be the supergraph that adjoins only the edge \( \set{ u, v } \).

  Since there is no path between \( u \) and \( v \) in \( T \), \( T' \) is also acyclic. But this contradicts the maximality of \( T \).

  Therefore, \( T \) is connected.

  \ImplicationSubProof{def:tree/connected_acyclic}{def:tree/minimally_connected} Suppose that \( T \) is connected and acyclic. Let \( \set{ u, v } \) be any edge of \( T \) and let \( T' \) be the subgraph that does not contain this edge.

  Suppose that \( T' \) is connected. Let \( w \) be any vertex and let \( p \) be a path in \( T' \) from \( u \) to \( w \). Then adding \( \set{ u, v } \) creates a cycle in \( T \), which contradicts our assumption that \( T \) is acyclic.

  Therefore, \( T' \) is not connected.

  \ImplicationSubProof{def:tree/minimally_connected}{def:tree/connected_acyclic} Suppose that \( T \) is minimally connected. We will show that it is acyclic.

  Suppose that \( T \) has a cycle \( p \). We can thus remove any edge of \( p \) from \( T \) and the resulting subgraph will also be connected. This contradicts the minimality of \( T \).

  Therefore, \( T \) is acyclic.

  \ImplicationSubProof{def:tree/connected_acyclic}{def:tree/single_path} Suppose that \( T \) is connected and acyclic. Since \( T \) is connected, there exists at least one path between any two vertices. Since it is acyclic, this path must be unique because otherwise we could easily create a cycle by joining two such paths.

  \ImplicationSubProof{def:tree/single_path}{def:tree/connected_acyclic} Suppose that, for every pair of vertices in \( T \), there exists a unique path connecting them. It is clear that \( T \) is connected.

  Suppose that \( T \) has a cycle \( p \) from \( u \) to \( u \) passing through \( v \). We thus obtain two paths from \( u \) to \( v \) --- one not containing the first edge in \( p \) and one not containing the last edge. But this contradicts our assumption that any two paths from \( u \) to \( v \) are equal.

  Therefore, \( T \) is acyclic.
\end{proof}

\begin{example}\label{ex:def:tree}
  The towers \eqref{eq:ex:tower_diagram_graph/forward}, \eqref{eq:ex:tower_diagram_graph/backward} and \eqref{eq:ex:tower_diagram_graph/two_sided} from \fullref{ex:tower_diagram_graph} are simple examples of infinite trees.

  Examples of finite trees are the grammar derivations \cref{fig:ex:natural_number_arithmetic_grammar/derivation/unambiguous} and \cref{fig:ex:natural_number_arithmetic_grammar/derivation/ambiguous}.
\end{example}

\begin{definition}\label{def:arborescence}\mcite[ch. 4, sec. 3.1]{GondranMinoux1984Graphs}
  Let \( T = (V, A) \) be a \hyperref[def:quiver/simple]{simple directed graph}. Suppose that there exists a vertex \( r \in V \) such that for every other vertex \( v \in V \), there exists a unique \hyperref[def:quiver_path/directed]{directed path} from \( r \) to \( v \). The pair \( (G, r) \) is called a (directed) \term{arborescence}. It is often more convenient to consider the triple \( T = (V, A, r) \) instead. The vertex \( r \) is called the \term{root} of the arborescence.

  \begin{thmenum}
    \thmitem{def:arborescence/undirected} The undirected analog of an arborescence is an \term{undirected arborescence} \( (T, r) \), where \( T \) is a simple undirected graph such that there exists a unique \hyperref[def:undirected_multigraph_path]{undirected path} from \( r \) to every other vertex.

    If \( T \) is a \hyperref[def:tree]{tree}, then \( (T, r) \) is an arborescence for every vertex \( r \in V \) as a consequence of \fullref{def:tree/single_path}. A tree with a distinguished vertex \( r \) is called a \term{rooted tree}.

    \thmitem{def:arborescence/depth} The \term{depth} of a node \( v \) is the length of the path from \( r \) to \( v \). The term \term{level} is also sometimes used but with a slightly different meaning --- the greater the depth, the lower the level.

    \thmitem{def:arborescence/ancestry} If \( v \) has a strictly greater depth than \( u \), we say that \( u \) is an \term{ancestor} of \( v \) and that \( v \) is a \term{descendant} of \( u \).

    The ancestor of \( v \) at the lowest possible level is called the \term{parent} of \( v \). If \( u \) is a parent of \( v \), \( u \) is a \term{child} of \( v \).

    Finally, if \( u \) and \( v \) are on the same level, we call them \term{siblings}.

    \thmitem{def:arborescence/height} The \term{height} or \term{depth} of the entire tree \( T \), if it exists, is the supremum of the depths of all nodes.

    \thmitem{def:arborescence/width} The \term{width} or \term{breadth} of the entire tree \( T \), if it exists, is the supremum of the number of siblings among all vertices. It is equal to the degree of \( T \) minus \( 1 \).

    We use terminology similar to \fullref{def:relation/arity}, e.g. \enquote{binary tree}, \enquote{ternary tree}, \ldots.

    \thmitem{def:arborescence/sub-arborescence} A \hyperref[def:theory_of_graphs/submodel]{subgraph} of an arborescence that is itself an arborescence is called a \term{sub-arborescence}. The same holds for rooted subtrees.

    For every node \( v \) of \( T \), we define the \term{induced sub-arborescence} of \( T \) with root \( v \) as the subgraph \hyperref[def:hypergraph/submodel]{induced} by \( V \setminus \dom(p) \), where \( p \) is the unique path from \( r \) to \( v \).
  \end{thmenum}
\end{definition}

\begin{theorem}[K\"onig's lemma]\label{thm:konigs_lemma}
  Every \hyperref[def:hypergraph/degree]{locally finite} \hyperref[def:arborescence]{arborescence} of \hyperref[def:hypergraph/order]{infinite} \hyperref[def:hypergraph/order]{order} contains a \hyperref[def:graph_adjancency_chain/simple]{simple path} of infinite length.
\end{theorem}
\begin{proof}
  Let \( T = (V, A, r) \) be a locally finite infinite arborescence. We will build an infinite simple path
  \begin{equation*}
    p = (p_1, p_2, \cdots)
  \end{equation*}
  using \hyperref[rem:natural_number_recursion]{natural number recursion} starting at one. Let \( c \) be a \hyperref[def:choice_function]{choice function} on the family
  \begin{equation*}
    \pow(V) \setminus \set{ \varnothing }.
  \end{equation*}

  Such a choice function exists by the \hyperref[def:zfc/choice]{axiom of choice}.

  \begin{itemize}
    \item Since \( T \) is locally finite, there are finitely many children of \( r \). If we suppose that the \hyperref[def:arborescence/sub-arborescence]{subarborescence} induced by \( v \) is finite for every child \( v \) of \( r \), then we would obtain that \( T \) itself is finite, which is a contradiction. Therefore, for at least one child, the induced sub-arborescence is infinite. Using the choice function \( c \), pick one such child and denote it by \( v_1 \).

    Define \( p_1 \) to be the arc \( r \to v_1 \).

    \item Fix \( k > 1 \). Using the same argument on the children of \( t(e_{k-1}) \), we can pick a child \( v_k \) that has an infinite induced sub-arborescence.

    Define \( e_k \) to be the arc \( t(e_{k-1}) \to v_k \). The path
    \begin{equation*}
      (e_1, e_2, \ldots, e_{k-1}, e_k)
    \end{equation*}
    is simple by construction because \( T \) is an arborescence and thus contains no directed cycles.
  \end{itemize}

  Thus, we have constructed an infinite simple path.
\end{proof}
