\section{Groups}\label{sec:groups}

\paragraph{Inverses in monoids}

\begin{definition}\label{def:monoid_inverse}\mcite[\S 8.5]{Schechter1997AnalysisHandbook}
  Let \( M \) be a \hyperref[def:monoid]{monoid}. We say that \( y \) is a \term{left inverse} (resp. \term{right inverse}) of \( x \) if \( yx = e \) (resp. \( xy = e \)).

  If \( y \) is simultaneously a left and right inverse of \( x \), we call a \term[ru=обратный (\cite[125]{БелоусовТкачёв2004ДискретнаяМатематика})]{two-sided inverse} or simply an \term{inverse} of \( x \) and denote it by \( x^{-1} \). It is unique as shown in \cref{thm:monoid_inverse_unique}. This notation is consistent with \hyperref[def:monoid/exponentiation]{monoid exponentiation}.

  We say that \( x \) is \term[en=invertible (\cite[31]{Jacobson1985BasicAlgebraI})]{invertible} if it has a two-sided inverse.
\end{definition}

\begin{proposition}\label{thm:monoid_inverse_unique}
  For every element \( x \) of any monoid, the (two-sided) \hyperref[def:monoid_inverse]{inverse} \( x^{-1} \) of \( x \), if it exists, is unique.
\end{proposition}
\begin{proof}
  If \( y \) and \( z \) are both inverses of \( x \), then
  \begin{equation*}
    y = ey = zxy = ze = z.
  \end{equation*}
\end{proof}

\begin{proposition}\label{thm:finite_cancellative_monoid_inverse}
  In a finite \hyperref[def:binary_operation/cancellative]{cancellative} monoid, every element has a \hyperref[def:monoid_inverse]{two-sided inverse}.
\end{proposition}
\begin{comments}
  \item Hence, every finite cancellative monoid is a \hyperref[def:group]{group}.
\end{comments}
\begin{proof}
  Let \( M \) be a finite cancellative monoid with \( n \) elements. Suppose that \( x \) has no inverse.

  Then, for any \( y \), the value \( xy \) is distinct from the neutral element \( e \). Furthermore, since \( M \) is cancellative, the map \( a \mapsto xa \) is injective because we can cancel \( x \) in \( xa = xb \) to obtain \( a = b \).

  Then there are at least \( n + 1 \) elements in \( M \), which is a contradiction. It remains for \( x \) to have an inverse.
\end{proof}

\paragraph{Groups}

\begin{definition}\label{def:group}\mcite[79]{Тыртышников2017ОсновыАлгебры}
  A \term[bg=група (\cite[372]{Обрешков1962ВисшаАлгебра}), ru=группа, en=group (\cite[def. 1.2]{Jacobson1985BasicAlgebraI}]{group} is a \hyperref[def:monoid]{monoid} in which every element has a \hyperref[def:monoid_inverse]{two-sided inverse}. They are the most well-studied and most well-behaved \hyperref[def:semigroup]{semigroups}.

  Groups have the following metamathematical properties:
  \begin{thmenum}
    \thmitem{def:group/theory}\mcite[5]{Hinman2005Logic} We can state an \hyperref[def:fol_equational_theory]{equational theory} for groups based on \hyperref[def:monoid/theory]{that of monoids}. Namely, we add a new \hyperref[def:fol_signature/notation]{condensed} unary \hyperref[def:fol_signature/function]{function symbol}, \( \syninv \), and extend the theory of monoid with the axioms
    \begin{align}
      \qforall \synx (\synx \synmult \syninv \synx \syneq \synx), \label{eq:def:group/theory/inverse/left} \\
      \qforall \synx (\syninv \synx \synmult \synx \syneq \synx). \label{eq:def:group/theory/inverse/right}
    \end{align}

    \thmitem{def:group/morphism}\mcite[80]{Тыртышников2017ОсновыАлгебры} Given groups \( G \) and \( H \), we say that the function \( \varphi: G \to H \) is a \term[ru=гомоморфизм (групп), en=homomorphism (of groups) (\cite[150]{Rotman2015AdvancedModernAlgebraPart1})]{group homomorphism} if it is a \hyperref[def:semigroup/morphism]{semigroup homomorphism}.

    We show in \cref{thm:group_homomorphism_single_condition} that semigroup homomorphisms are \hyperref[def:fol_homomorphism]{first-order homomorphism} in the theory of groups.

    \thmitem{def:group/category}\mcite[12]{MacLane1998CategoryTheory} For a fixed \hyperref[def:grothendieck_universe]{Grothendieck universe} \( \mscrU \), the first-order theory of monoids naturally gives rise to a \hyperref[def:fol_theory_model_functor/obj]{category of \( \mscrU \)-small models}.

    We denote this category by \( \ucat{Grp} \).

    We list several properties:
    \begin{itemize}
      \item \( \ucat{Grp} \) is \hyperref[def:concrete_category]{concrete} over the \hyperref[def:monoid/category]{category of \( \mscrU \)-small monoids} and the \hyperref[def:set_with_involution/category]{category of \( \mscrU \)-small sets with involutions}.

      \item As an equational theory, the theory of groups has \hyperref[def:fol_free_term_model]{free term models}, and \fullref{thm:fol_free_term_model_universal_property} provides a \hyperref[def:category_adjunction]{left adjoint} to the \hyperref[def:fol_theory_forgetful_functor/set]{forgetful functor} \( U: \ucat{Grp} \to \ucat{Set} \).

      A more convenient construction is the \hyperref[def:free_group]{free group}, for which we have \fullref{thm:free_monoid_universal_property}.

      The two constructions are necessarily isomorphic because of \cref{thm:functor_adjoint_uniqueness}.

      \item Because of this left adjoint, \cref{thm:concrete_category_function_invertibility/injective} implies that a group homomorphism is a \hyperref[def:morphism_invertibility/left_cancellative]{monomorphism} if and only if it is \hyperref[def:function_invertibility/injective]{injective}.

      \item Unlike in the category \hyperref[def:monoid/category]{\( \ucat{Mon} \)} of monoids, in \( \ucat{Grp} \) every epimorphism is surjective. We will prove this in \cref{thm:group_epimorphisms_are_surjective}.

      To avoid circularity, in this section we will not use that epimorphism are surjective.
    \end{itemize}

    \thmitem{def:group/subobject}\mcite[91]{Тыртышников2017ОсновыАлгебры} A \term[ru=подгрупа (\cite[378]{Обрешков1962ВисшаАлгебра}), ru=подгруппа, en=subgroup (\cite[31]{Jacobson1985BasicAlgebraI})]{subgroup} of \( G \) is a \hyperref[def:monoid/subobject]{submonoid} that is closed with respect to inverses.

    It follows from \cref{thm:fol_equational_theory_models/substructure} that the subgroups of \( G \) are precisely its \hyperref[def:fol_substructure/submodel]{first-order submodels}, and from \cref{thm:fol_categorical_subobjects} that they are its \hyperref[def:categorical_subobject]{categorical subobjects}.

    Furthermore, as a consequence of \cref{thm:fol_equational_theory_models/image}, the image of a group homomorphism is a subgroup of its codomain.

    \thmitem{def:group/generated}\mcite[43]{Jacobson1985BasicAlgebraI} For an arbitrary subset \( A \) of \( G \), we define the \term{generated subgroup} \( \braket{ A } \) of \( A \) as the intersection of subgroups containing \( A \).

    This coincides with the first-order submodel of \( G \) \hyperref[def:fol_generated_substructure]{generated} by \( A \). \Cref{thm:fol_generated_substructure_iteration} implies that, in addition to the elements of \( A \), \( \braket{ A } \) contains their products and inverses, the products of their products and inverses, etc...

    \thmitem{def:group/exponentiation}\mcite[86]{Тыртышников2017ОсновыАлгебры} We extend \hyperref[def:monoid/exponentiation]{monoid exponentiation} to all integers by setting
    \begin{equation}\label{eq:def:group/exponentiation}
      x^{-n} \coloneqq (x^n)^{-1}.
    \end{equation}

    This operation behaves well with respect to inverses, as we will show in \cref{thm:def:group/negative_power}.

    \thmitem{def:group/trivial}\mimprovised In accordance with \cref{def:trivial_object}, we will call a one-element monoid \term{trivial} because, as shown in \cref{thm:equational_theory_zero_morphisms/trivial}, it is a zero object in \( \ucat{Grp} \).

    \thmitem{def:group/opposite}\mcite[example 1.2.2(iii)]{Riehl2016CategoryTheory} The \hyperref[def:monoid/opposite]{opposite monoid} \( G^{\oppos} \) of a group \( G \) is again a group with the exact same inverse operation.

    For this reason, we will call it the \term{opposite group}.

    \thmitem{def:group/kernel}\mcite[def. II.6.5]{Aluffi2009Algebra} We define the \term[bg=ядро (\cite[def. IV.16]{ГеновМиховскиМоллов1991Алгебра}), ru=ядро (\cite[104]{Тыртышников2017ОсновыАлгебры})]{kernel} of a group homomorphism \( \varphi: G \to H \) as the set
    \begin{equation}\label{eq:def:group/kernel}
      \ker \varphi \coloneqq \varphi^{-1}(e_H) = \set{ x \in G \given \varphi(x) = e_H }.
    \end{equation}

    \Cref{thm:equational_theory_zero_morphisms/kernel} implies that this is a subgroup acting as the \hyperref[def:categorical_kernels/kernel]{categorical kernel} of \( \varphi \).

    \thmitem{def:group/quotient}\mimprovised \hyperref[def:fol_congruence]{First-order congruences} in group theory are particularly well-behaved. \Cref{thm:normal_subgroups_and_congruences} shows that congruences correspond exactly to \hyperref[def:normal_subgroup]{normal subgroups}, and \cref{thm:subgroup_cosets} shows how the corresponding equivalence classes are all equinumerous and have a simple form expressible via normal subgroups.

    For this reason, a \term[bg=фактор-група (\cite[97]{ГеновМиховскиМоллов1991Алгебра}), ru=фактор-группа (\cite[94]{Тыртышников2017ОсновыАлгебры})]{quotient group} \( G / N \) is traditionally defined with respect to a normal subgroup \( N \) rather than with respect to the corresponding congruence.

    \Cref{thm:fol_categorical_quotient_objects} implies that the quotient groups are precisely the \hyperref[def:categorical_quotient_object]{categorical quotient objects}.
  \end{thmenum}
\end{definition}

\begin{example}\label{ex:power_set_is_not_a_group}
  The \hyperref[ex:def:monoid/power]{power monoid} \( \pow(G) \) of a group \( G \) is a monoid, but it is not a group unless \( G \) is \hyperref[def:group/trivial]{trivial}.

  Indeed, fix an arbitrary set \( A \) and suppose that there exists a set \( B \) such that \( AB = \set{ e_G } \). Then, for any \( a \) in \( A \) and \( b \) in \( B \), we must have \( ab = e_G \). \Cref{thm:monoid_inverse_unique} implies that inverses are unique, hence \( b = a^{-1} \). Furthermore, there can only be one element in \( A \) because otherwise we would have some \( a' \) in \( A \) distinct from \( a \), and \( a' a^{-1} \) would not equal \( e_G \).

  Therefore, if \( AB = \set{ e_G } \), then \( A \) has a single element and \( B \) contains only the inverse of that element. Any subset of \( G \) that has more than one element does not have an inverse.
\end{example}

\begin{proposition}\label{thm:def:group}
  Every \hyperref[def:group]{group} \( G \) has the following basic properties:
  \begin{thmenum}
    \thmitem{thm:def:group/cancellative} The group operation is \hyperref[def:binary_operation/cancellative]{cancellative}.
    \thmitem{thm:def:group/involution} The inverse operation is an \hyperref[def:involution]{involution}: \( x = (x^{-1})^{-1} \).
    \thmitem{thm:def:group/neutral_inverse} The neutral element \( e \) is its own inverse.
    \thmitem{thm:def:group/neutral_iff_inverse} \( x \) is the neutral element if and only if \( x^{-1} \) is.
    \thmitem{thm:def:group/inverse_contravariant} Inverses are contravariant: \( (xy)^{-1} = y^{-1} x^{-1} \).
    \thmitem{thm:def:group/negative_power} For any nonnegative integer \( n \), \( (x^n)^{-1} = (x^{-1})^n \)
    \thmitem{thm:def:group/equality_of_powers} We have \( x^n = x^m \) if and only if \( x^{n-m} = e \).
    \thmitem{thm:def:group/inverse_isomorphism} The map \( x \mapsto x^{-1} \) is a group isomorphism.
  \end{thmenum}
\end{proposition}
\begin{proof}
  \SubProofOf{thm:def:group/cancellative} If \( x = y \), obviously \( xz = yz \) and \( zx = zy \). Now if \( xz = yz \), we have
  \begin{equation*}
    x = x(zz^{-1}) = (xz)z^{-1} = (yz)z^{-1} = y(zz^{-1}) = y.
  \end{equation*}

  The case \( zx = zy \) is analogous.

  \SubProofOf{thm:def:group/involution}
  \begin{equation*}
    (x^{-1})^{-1}
    =
    (x x^{-1}) (x^{-1})^{-1}
    =
    x (x^{-1} (x^{-1})^{-1})
    =
    x.
  \end{equation*}

  \SubProofOf{thm:def:group/neutral_inverse} \( ee = e \).

  \SubProofOf{thm:def:group/neutral_iff_inverse} If \( x = e \), \cref{thm:def:group/neutral_inverse} implies that \( x^{-1} = e \). Conversely, if \( x^{-1} = e \), we have \( x = (x^{-1})^{-1} = e \) due to \cref{thm:def:group/involution}.

  \SubProofOf{thm:def:group/inverse_contravariant}
  \begin{equation*}
    (xy) (y^{-1} x^{-1})
    =
    x (y y^{-1}) x^{-1}
    =
    e
    =
    y^{-1} (x^{-1} x) y
    =
    (y^{-1} x^{-1}) (xy).
  \end{equation*}

  \SubProofOf{thm:def:group/negative_power} We proceed by induction on \( n \).
  \begin{itemize}
    \item If \( n = 0 \), since \( 0 = -0 \), clearly \( x^{-0} = e = (x^0)^{-1} \) (we have used that \( e \) is its own inverse).
    \item If the inductive hypothesis holds for \( n \), then
    \begin{equation*}
      x^{-(n+1)}
      =
      (x^{n+1})^{-1}
      =
      (x^n \cdot x)^{-1}
      \reloset {\ref{thm:def:group/inverse_contravariant}} =
      x^{-1} (x^n)^{-1}
      \reloset {\T{ind.}} =
      x^{-1} (x^{-1})^n
      \reloset {\eqref{eq:thm:semigroup_exponentiation_properties/distributive}} =
      (x^{-1})^{(n-1)}.
    \end{equation*}
  \end{itemize}

  \SubProofOf{thm:def:group/equality_of_powers}
  \SufficiencySubProof* If \( x^n = x^m \), then, multiplying by \( x^{-m} \), we obtain
  \begin{equation*}
    \underbrace{x^n \cdot x^{-m}}_{x^{n-m} \T*{due to} \eqref{eq:thm:semigroup_exponentiation_properties/multiplication}}
    =
    \underbrace{x^m \cdot x^{-m}}_e.
  \end{equation*}

  \NecessitySubProof* Conversely, if \( x^{n-m} = e \), then similarly we can multiply by \( x^m \) to obtain \( x^n = x^m \).

  \SubProofOf{thm:def:group/inverse_isomorphism} Trivial.
\end{proof}

\begin{proposition}\label{thm:group_homomorphism_single_condition}
  A function between groups satisfies the semigroup homomorphism condition \eqref{eq:def:semigroup/morphism} if and only if it is a \hyperref[def:monoid/morphism]{monoid homomorphism} that preserves inverses.
\end{proposition}
\begin{comments}
  \item This allows to define a group homomorphism in \cref{def:group/morphism} without requiring it to preserve neutral elements and inverses like \cref{def:fol_homomorphism} suggests.
\end{comments}
\begin{proof}
  \SufficiencySubProof \eqref{eq:def:semigroup/morphism} is required to hold by definition.

  \NecessitySubProof Suppose that the function \( \varphi: G \to H \) satisfies \eqref{eq:def:semigroup/morphism}.

  Then it preserves identities, i.e. is a \hyperref[def:pointed_set/morphism]{pointed set homomorphism}, because
  \begin{equation*}
    \varphi(e_G) = \varphi(e_G \cdot e_G) = \varphi(e_G) \cdot \varphi(e_G)
  \end{equation*}
  and
  \begin{equation*}
    \varphi(e_G) = \varphi(e_G) \cdot e_H.
  \end{equation*}

  By \cref{thm:def:group/cancellative}, \( \varphi \) is cancellative, so we can cancel \( \varphi(e_G) \) and obtain \( e_H = \varphi(e_G) \).

  Furthermore, inverses are also also preserved because
  \begin{equation*}
    \varphi(x^{-1})
    =
    \varphi(x^{-1}) \cdot e_H
    =
    \varphi(x^{-1}) \cdot \varphi(x) \cdot \varphi(x)^{-1}
    =
    \varphi(x^{-1} x) \cdot \varphi(x)^{-1}
    =
    e_H \cdot \varphi(x)^{-1}
    =
    \varphi(x)^{-1}.
  \end{equation*}
\end{proof}

\begin{proposition}\label{thm:invertible_submonoid_is_group}
  The set of all \hyperref[def:monoid_inverse]{invertible} elements of a \hyperref[def:monoid]{monoid} is a \hyperref[def:group]{group}.
\end{proposition}
\begin{proof}
  Let \( M \) be a monoid and let \( I \) be the set of invertible elements of \( M \).

  \begin{itemize}
    \item The neutral element is invertible, hence \( I \) is closed under the nullary identity operation.
    \item If \( x \) is invertible with inverse \( x^{-1} \), then \( x^{-1} \) is invertible with inverse \( x \), and thus \( I \) is closed under the unary inversion operation.
    \item If \( x \) and \( y \) are invertible, then \( xy \) is invertible with inverse \( y^{-1} x^{-1} \). Thus, \( I \) is closed under the binary monoid operation.
  \end{itemize}

  Therefore, the set \( I \) of invertible elements is a submonoid of \( M \) and, furthermore, it is a group because it contains an inverse for each of its elements.
\end{proof}

\begin{proposition}\label{thm:ordered_group}
  \hyperref[def:ordered_semiring]{Ordered} \hyperref[def:group]{groups} have the following properties:
  \begin{thmenum}
    \thmitem{thm:ordered_group/inversion} We have \( x < y \) if and only if \( y^{-1} < x^{-1} \).
    \thmitem{thm:ordered_group/sign_inversion} We have \( x > e \) if and only if \( x^{-1} < e \).
  \end{thmenum}
\end{proposition}
\begin{comments}
  \item Some results hold more generally for \hyperref[def:ordered_semigroup]{ordered semigroups} --- see \cref{thm:def:ordered_semigroup}.
\end{comments}
\begin{proof}
  \SubProofOf{thm:ordered_group/inversion} If \( x < y \), \cref{thm:def:ordered_semigroup/cancellative} implies that
  \begin{equation*}
    y^{-1} = x^{-1} x y^{-1} < x^{-1} y y^{-1} = x^{-1}.
  \end{equation*}

  The converse direction follows analogously.

  \SubProofOf{thm:ordered_group/sign_inversion} Follows from \cref{thm:ordered_group/inversion} since, by \cref{thm:def:group/neutral_inverse}, \( e^{-1} = e \).
\end{proof}

\paragraph{Group cosets}

\begin{proposition}\label{thm:subgroup_cosets}
  Fix a group \( G \) and a \hyperref[def:group/subobject]{subgroup} \( H \) of \( G \).

  \begin{thmenum}
    \thmitem{thm:subgroup_cosets/equivalence_relation} The relation
    \begin{equation}\label{eq:thm:subgroup_cosets/equivalence_relation}
      x \sim y \T{if and only if} x^{-1} y \in H\fnote{The preference of \( x^{-1} y \) to \( x y^{-1} \) is largely arbitrary since each implies the other. We follow the convention from \cite[prop. 7.6]{Aluffi2009Algebra}.},
    \end{equation}
    is an \hyperref[def:equivalence_relation]{equivalence relation} (but not in general a congruence --- see \cref{ex:quotient_by_non_normal_subgroup}).

    \thmitem{thm:subgroup_cosets/translations} For every element \( a \) of any coset \( C \) (under the above relation), we have
    \begin{equation*}
      C = a H = \set{ a \cdot h \given h \in H }.
    \end{equation*}

    \thmitem{thm:subgroup_cosets/translation_congruence} We have \( a H = b H \) and only if \( a \sim b \).

    \thmitem{thm:subgroup_cosets/inside} We have \( a H = H \) and only if \( a \in H \).

    \thmitem{thm:subgroup_cosets/equinumerous} Each coset is \hyperref[def:equinumerosity]{equinumerous} with \( H \).
  \end{thmenum}
\end{proposition}
\begin{proof}
  \SubProofOf{thm:subgroup_cosets/equivalence_relation} We will show that \( {\sim} \) is an equivalence relation.

  \SubProofOf*[def:binary_relation/reflexive]{reflexivity} Since \( H \) is closed under the group operations, it contains \( e_G = x^{-1} x \).

  \SubProofOf*[def:binary_relation/symmetric]{symmetry} If \( x \sim y \), then \( xy^{-1} \) is in \( H \), so is \( y^{-1} x = (x^{-1} y)^{-1} \). As a subgroup, \( H \) contains the inverse of all its elements, and thus \( y \sim x \).

  \SubProofOf*[def:binary_relation/transitive]{transitivity} If \( x \sim y \) and \( y \sim z \), then
  \begin{equation*}
    x^{-1} z
    =
    x^{-1} (y y^{-1}) z
    =
    (x^{-1} y) (y^{-1} z).
  \end{equation*}

  The latter factors are both in \( H \), hence their product \( x^{-1} z \) must also be in \( H \). Thus, \( x \sim z \).

  \SubProofOf{thm:subgroup_cosets/translations} Fix an equivalence class \( C \) and let \( a \) be a member of \( C \).

  If \( x \) is in \( C \), then \( x \sim a \) and thus \( a \sim x \). Then \( a^{-1} x \) is in \( H \), so \( x \) is in its translate \( aH \).

  Conversely, if \( x \) is in \( a H \), then \( a^{-1} x \) is in \( a^{-1} a H = H \). Thus, \( a \sim x \), implying that \( x \) is in \( C \).

  We conclude that \( C = a H \).

  \SubProofOf{thm:subgroup_cosets/translation_congruence}

  \SufficiencySubProof* Suppose that \( a H = b H \).

  Since \( b \) belongs to \( b H \), there must exist some \( h \) in \( H \) such that \( a h = b \). Then \( h = a^{-1} b \), that is, \( a \sim b \).

  \NecessitySubProof* Suppose that \( a \sim b \), i.e. \( a^{-1} b \) is in \( H \).

  Fix \( y \) from \( b H \). There must exist some \( h \) in \( H \) such that \( y = bh \). Then \( a^{-1} b h \) is in \( H \), and after multiplying on the left by \( a \), we obtain that
  \begin{equation*}
    y = bh = a \underbrace{a^{-1} b h}_{\in H} \in a H.
  \end{equation*}

  Therefore, \( b H \subseteq a H \).

  We can similarly show that \( a H \subseteq b H \).

  \SubProofOf{thm:subgroup_cosets/inside} \Cref{thm:subgroup_cosets/translation_congruence} implies that \( e H = a H \) if and only if \( e \sim a \), i.e. \( e^{-1} a \in H \).

  \SubProofOf{thm:subgroup_cosets/equinumerous} Suppose we are given a coset \( a H \). Define the function
  \begin{equation*}
    \begin{aligned}
      &f: H \to a H, \\
      &f(x) \coloneqq a x.
    \end{aligned}
  \end{equation*}

  We must show that \( f \) is bijective on \( G \).

  \SubProofOf*[def:function_invertibility/injective/equality]{injectivity} Let \( f(x) = f(y) \). Then \( a x = a y \), which by the cancellative property shown in \cref{thm:def:group/cancellative} implies that \( x = y \).

  Hence, \( f \) is injective.

  \SubProofOf*[def:function_invertibility/surjective/existence]{surjectivity} Fix an element \( y \) from \( a H \). Then \( a^{-1} y \) is in \( H \) and, furthermore,
  \begin{equation*}
    f(a^{-1} y) = a a^{-1} y = y.
  \end{equation*}

  Hence, \( f \) is surjective.
\end{proof}

\begin{definition}\label{def:subgroup_cosets}\mcite[52]{Jacobson1985BasicAlgebraI}
  We have shown in \cref{thm:subgroup_cosets} that each subgroup \( H \) of \( G \) induces a partition of \( G \) into translations of \( H \). We refer to the the family \( \seq{ aH }_{a \in G} \) \term[bg=ляв съседен клас (\cite[91]{ГеновМиховскиМоллов1991Алгебра}), ru=левый смежный класс (\cite[93]{Тыртышников2017ОсновыАлгебры})]{left cosets} and to \( \seq{ Ha }_{a \in G} \) as \term[bg=десен съседен клас (\cite[91]{ГеновМиховскиМоллов1991Алгебра}), ru=правый смежный класс (\cite[93]{Тыртышников2017ОсновыАлгебры})]{right cosets}.

  If \( H \) is \hyperref[def:normal_subgroup]{normal}, \cref{thm:normal_subgroup_left_right_cosets} implies that \( aH = Ha \), so the two notions coincide.
\end{definition}

\paragraph{Group conjugation}

\begin{definition}\label{def:group_conjugation}\mcite[165]{Knapp2016BasicAlgebra}
  Fix a group \( G \) and an element \( x \). For any elements \( g \) and \( y \) of \( G \), we have
  \begin{equation}\label{eq:def:group_conjugation}
    y = g x g^{-1} \T{if and only if} x = g^{-1} y g.
  \end{equation}

  We say that \( y \) is a \term{left conjugate} of \( x \) by \( g \) and that \( x \) is a \term{right conjugate} of \( y \) by \( g \).

  Note also that \( y \) is a left conjugate of \( x \) by \( g \) if and only if it is a right conjugate by \( g^{-1} \). If the element by which we conjugate is irrelevant, we simply say that \( x \) and \( y \) are \term{conjugates}.
\end{definition}
\begin{comments}
  \item The adjectives \enquote{left} and \enquote{right} reflect the fact that \( x \mapsto g x g^{-1} \) is a \hyperref[def:group_action]{left group action}, while \( x \mapsto g^{-1} x g \) is a \hyperref[def:group_action]{right group action} --- see \cref{thm:group_conjugation_action}.
\end{comments}

\begin{example}\label{ex:def:group_conjugation}
  We list some examples of \hyperref[def:group_conjugation]{group conjugation}:
  \begin{thmenum}
    \thmitem{ex:def:group_conjugation/matrix_similarity} \hyperref[def:similar_matrices]{Matrix similarity} is conjugation in the \hyperref[def:linear_groups]{linear groups}.
  \end{thmenum}
\end{example}

\begin{proposition}\label{thm:def:group_conjugation}
  \hyperref[def:group_conjugation]{Group conjugation} is an \hyperref[def:equivalence_relation]{equivalence relation}.
\end{proposition}
\begin{proof}
  Fix a group \( G \).

  \SubProofOf*[def:binary_relation/reflexive]{reflexivity} \( x = e x e^{-1} \).

  \SubProofOf*[def:binary_relation/symmetric]{symmetry} If \( y = g x g^{-1} \), then \( x = g^{-1} y (g^{-1})^{-1} \).

  \SubProofOf*[def:binary_relation/transitive]{transitivity} If \( y = g x g^{-1} \) and \( z = h y h^{-1} \), then
  \begin{equation*}
    z = (hg) x (hg)^{-1},
  \end{equation*}
  where \( (hg)^{-1} = g^{-1} h^{-1} \) due to \cref{thm:def:group/inverse_contravariant}.
\end{proof}

\paragraph{Group quotients}

\begin{definition}\label{def:normal_subgroup}\mcite[def. 1.5]{Jacobson1985BasicAlgebraI}
  We say that a \hyperref[def:group/subobject]{subgroup} is \term[ru=нормальная подгруппа (\cite[94]{Тыртышников2017ОсновыАлгебры})]{normal} if it is closed under (left or right) \hyperref[def:group_conjugation]{conjugation} by arbitrary elements.

  Explicitly, the subgroup \( N \) of \( G \) is normal if, for any \( n \) in \( N \) and an arbitrary \( g \) in \( G \), the left conjugate \( gng^{-1} \) belongs to \( N \). An equivalent condition is for the right conjugate \( gng^{-1} \) to belong to \( N \).
\end{definition}

\begin{proposition}\label{thm:kernel_is_normal_subgroup}
  The \hyperref[def:group/kernel]{kernel} of a \hyperref[def:group/morphism]{group homomorphism} is a \hyperref[def:normal_subgroup]{normal subgroup} of its domain.
\end{proposition}
\begin{proof}
  Consider the homomorphism \( \varphi: G \to H \). Since the kernel \( \ker \varphi \) coincides with the categorical kernel, it is clearly a subgroup of \( G \).

  For any group element \( g \) and every \( n \) in the kernel of \( \varphi \), the left conjugate \( gng^{-1} \) satisfies
  \begin{equation*}
    \varphi(gng^{-1})
    =
    \varphi(g) \cdot \varphi(n) \cdot \varphi(g^{-1})
    =
    \varphi(g) \cdot e_H \cdot \varphi(g^{-1})
    =
    e_H,
  \end{equation*}
  hence \( gng^{-1} \) is also in the kernel.

  Therefore, the kernel is normal.
\end{proof}

\begin{example}\label{ex:quotient_by_non_normal_subgroup}
  This example uses forward references to \fullref{sec:group_actions} to demonstrate that the equivalence relations induced by arbitrary subgroups may not be congruences. Only the multiplication table from \cref{ex:s3} will matter to us.

  Consider the \hyperref[def:symmetric_group]{symmetric group} \( S_3 \) discussed in \cref{ex:s3} and \cref{ex:s3_and_a3} and the subgroup
  \begin{equation*}
    H = \set{ \id, \cycle{ 1, 2 } }.
  \end{equation*}

  It is indeed a subgroup:
  \begin{itemize}
    \item \( H \) contains the neutral element.
    \item \( H \) is closed under inverses because, as a transposition, \( \cycle{ 1, 2 } \) is its own inverse.
    \item \( H \) is closed under products for the same reason.
  \end{itemize}

  It is not \hyperref[def:normal_subgroup]{normal}, however, because the conjugate
  \begin{equation*}
    \cycle{ 1, 3 }
    \cdot
    \cycle{ 1, 2 }
    \cdot
    \cycle{ 1, 3 }
    =
    \cycle{ 2, 3 },
  \end{equation*}
  does not belong to \( H \).

  Consider the equivalence relation \( {\sim} \) induced by \( H \) as per \cref{thm:subgroup_cosets/equivalence_relation}. It is not a congruence as we will now show.

  We have \( \cycle{ 1, 3 } \sim \cycle{ 1, 2, 3 } \) because
  \begin{equation*}
    \cycle{ 1, 3 } \bincirc \cycle{ 1, 3, 2 }
    =
    \underbrace{\cycle{ 1, 2 }}_{\in H}
  \end{equation*}

  Similarly, \( \cycle{ 2, 3 } \sim \cycle{ 1, 3, 2 } \). These are actually the three left cosets:
  \begin{align*}
    \underbrace{ \set{ \id, \cycle{ 1, 2 } } }_{ H },
    &&
    \underbrace{ \set{ \cycle{ 1, 3 }, \cycle{ 1, 2, 3 } } }_{ \cycle{ 1, 3 } H },
    &&
    \underbrace{ \set{ \cycle{ 2, 3 }, \cycle{ 1, 3, 2 } } }_{ \cycle{ 2, 3 } H }.
  \end{align*}

  But \eqref{eq:def:fol_congruence/functions} is not satisfied because
  \begin{equation*}
    \cycle{ 1, 3 } \bincirc \cycle{ 2, 3 }
    =
    \cycle{ 1, 3, 2 }
    \ncong
    \id
    =
    \cycle{ 1, 3, 2 } \bincirc \cycle{ 1, 2, 3 }.
  \end{equation*}

  So, if the subgroup is not normal, \cref{thm:normal_subgroups_and_congruences}, upon which we often rely, does not hold. Furthermore, \cref{thm:normal_subgroup_left_right_cosets}, which gives us certain consistency conditions, also does not hold because the left cosets above differ from the right cosets:
  \begin{align*}
    \underbrace{ \set{ \id, \cycle{ 1, 2 } } }_{ H },
    &&
    \underbrace{ \set{ \cycle{ 1, 3 }, \cycle{ 1, 3, 2 } } }_{ H \cycle{ 1, 3 } },
    &&
    \underbrace{ \set{ \cycle{ 2, 3 }, \cycle{ 1, 2, 3 } } }_{ H \cycle{ 2, 3 } }.
  \end{align*}
\end{example}

\begin{example}\label{ex:image_of_normal_subgroup}
  We will use forward references to \fullref{sec:free_groups} and \fullref{sec:group_actions} to show that the \hyperref[def:set_valued_map/image]{image} of a normal subgroup under a homomorphism is not necessarily normal.

  Consider the \hyperref[def:symmetric_group]{symmetric group} \( S_3 \) discussed in \cref{ex:s3} and the \hyperref[def:cyclic_group]{cyclic group} \( C_2 \).

  Define the homomorphism
  \begin{equation*}
    \begin{aligned}
      &\varphi: C_2 \to S_3, \\
      &\varphi(x) \coloneqq \begin{cases}
        \id,            &x = a^0, \\
        \cycle{ 1, 2 }, &x = a^1.
      \end{cases}
    \end{aligned}
  \end{equation*}

  Based on the \hyperref[def:cayley_table]{multiplication table} in \cref{ex:s3}, we can conclude that \( \cycle{ 1, 2 }^2 = \id \). Hence, \( \varphi \) is indeed a homomorphism.

  Its image is not normal, however, because the left conjugation of \( \cycle{ 1, 2 } \) by \( \cycle{ 1, 2, 3 } \) is not in the image of \( \varphi \):
  \begin{equation*}
    \cycle{ 1, 2, 3 } \bincirc \parens[\big]{ \cycle{ 1, 2 } \bincirc \cycle{ 1, 3, 2 } }
    =
    \cycle{ 1, 2, 3 } \bincirc \cycle{ 1, 3 }
    =
    \cycle{ 2, 3 }.
  \end{equation*}
\end{example}

\begin{proposition}\label{thm:normal_subgroups_and_congruences}
  In a \hyperref[def:group]{group} \( G \), the \hyperref[def:fol_congruence]{first order congruences} and the \hyperref[def:normal_subgroup]{normal subgroup} are related as follows:
  \begin{thmenum}
    \thmitem{thm:normal_subgroups_and_congruences/cong_to_subgroup} For every congruence in \( G \), the coset \( [e_G] \) of the neutral element is a normal subgroup of \( G \).

    \thmitem{thm:normal_subgroups_and_congruences/subgroup_to_cong} Conversely, for every normal subgroup \( N \), the equivalence relation from \eqref{eq:thm:subgroup_cosets/equivalence_relation} is a congruence.

    \thmitem{thm:normal_subgroups_and_congruences/inverse} The procedures for obtaining a congruence from a normal subgroup and vice versa are inverses.
  \end{thmenum}
\end{proposition}
\begin{comments}
  \item This allows us to consider the quotient \( G / N \) with respect to a normal subgroup rather than \( G / {\cong} \) with respect to a congruence.

  \item This fails for more general structures like monoids --- see \cref{ex:congruence_with_cosets_of_different_cardinality}.

  \item A similar statement holds for ring ideals --- see \cref{thm:ideals_and_congruences} --- and submodules --- see \cref{thm:submodules_and_congruences}

  \item \Cref{thm:normal_subgroups_and_congruences/subgroup_to_cong} extends \cref{thm:subgroup_cosets/equivalence_relation}, which shows that, for arbitrary subgroups, \( {\cong} \) is merely an equivalence relation.
\end{comments}
\begin{proof}
  \SubProofOf{thm:normal_subgroups_and_congruences/cong_to_subgroup} Any congruence \( {\cong} \) must respect the group operations.

  In particular, \( [e_G] \) contains the neutral element, the inverse of each of its elements, as well as the product of each pair. This makes \( [e_G] \) a subgroup.

  We will now show that it is a normal subgroup. Fix elements \( g \) from \( G \) and \( n \) from \( [e_G] \). Then
  \begin{equation*}
    g \cdot n \cong \underbrace{g \cdot e_G}_{g},
  \end{equation*}
  hence
  \begin{equation*}
    g \cdot n \cdot g^{-1} \cong e_G.
  \end{equation*}

  \SubProofOf{thm:normal_subgroups_and_congruences/subgroup_to_cong} Let \( N \) be a normal subgroup and consider the relation \( {\sim} \) defined in \eqref{eq:thm:subgroup_cosets/equivalence_relation}.

  We must verify that it is a congruence. \Cref{thm:subgroup_cosets/equivalence_relation} already shows that it is an equivalence relation, so it remains to verify that it respects operations.

  \SubProof*{Proof that \( {\sim} \) respects inverses} If \( x \sim y \), then \( x^{-1} y \) is in \( N \). Since \( N \) is normal, it contains the left conjugate
  \begin{equation*}
    x (x^{-1} y) x^{-1} = y x^{-1}.
  \end{equation*}

  As a subgroup, \( N \) contains the inverses of its elements, hence it contains
  \begin{equation*}
    (x^{-1})^{-1} (y^{-1}).
  \end{equation*}

  Therefore, \( x^{-1} \sim y^{-1} \).

  \SubProof*{Proof that \( {\sim} \) preserves products} Suppose that \( a \sim b \) and \( x \sim y \).

  Then \( a^{-1} b \) belongs to \( N \). Since \( N \) is normal, it also contains the right conjugate \( y^{-1} (a^{-1} b) y \).

  Since \( N \) is a subgroup and contains the products of its elements, \( N \) also contains the following element:
  \begin{equation*}
    (x^{-1} y) y^{-1} (a^{-1} b) y = x^{-1} a^{-1} b y = (a x)^{-1} b y.
  \end{equation*}

  Therefore, \( ax \cong by \).

  \SubProofOf{thm:normal_subgroups_and_congruences/inverse}

  \SubProof*{Congruences can be restored} Suppose we have a congruence \( {\cong} \) on \( G \). Let \( {\approx} \) be congruence generated by the normal subgroup \( [e_G] \). We must show that \( {\cong} \) and \( {\approx} \) coincide.

  Indeed, \( x \cong y \) if and only if \( x^{-1} y \cong e_G \) if and only if \( x^{-1} y \in [e_G] \) if and only if \( x \approx y \).

  \SubProof*{Normal subgroups can be restored} Conversely, suppose we have a normal subgroup \( N \) on \( G \). Let \( {\cong} \) be the corresponding congruence and consider the normal subgroup \( [e_G] \). We must show that \( N = [e_G] \).

  For every \( x \) we have \( x = e_G \cdot x = e_G^{-1} \cdot x \), hence \( x \in [e_G] \) if and only if \( x \in N \).
\end{proof}

\begin{example}\label{ex:congruence_with_cosets_of_different_cardinality}
  Consider the monoid \( (\BbbN, \max) \) of natural numbers with \( \max(a, b) \) as the monoid operation. This is an initial segment of the additive monoid of one of the tropical semirings discussed in \cref{def:tropical_semiring}.

  Define the relation \( a \cong b \) to hold if \( a \) and \( b \) have the same number of decimal digits. It is clearly an equivalence relation. Furthermore, if \( a \cong a' \) and \( b \cong b' \) and \( a \ncong b \), then \( b \) and \( b' \) simultaneously have either more or less digits than \( a \) and \( a' \), hence \( \max(a, b) \) should have the same amount of digits as \( \max(a', b') \). Hence, \( {\cong} \) is a congruence on \( (\BbbN, \max) \).

  We have the following cosets under \( {\cong} \):
  \begin{itemize}
    \item \( 10 \) numbers with \( 1 \) digit: \( 0, 1, \ldots, 8, 9 \),
    \item \( 90 \) numbers with \( 2 \) digits: \( 10, 11, \ldots, 98, 99 \),
    \item \( 900 \) numbers with \( 3 \) digits: \( 100, 11, \ldots, 998, 999 \),
    \item \( \cdots \)
    \item \( 9 \cdot 10^n \) numbers with \( (n - 1) \) digits,
    \item \( \cdots \)
  \end{itemize}

  Clearly the distinct cosets have different sizes. This demonstrates show \cref{thm:subgroup_cosets/equinumerous} may fail to hold more generally.
\end{example}

\begin{definition}\label{def:equivalence_modulo}\mimprovised
  For any \hyperref[def:equivalence_relation]{equivalence relation} \( {\sim} \), if \( x \sim y \), we say that \( x \) and \( y \) are \term{equivalent modulo} \( {\sim} \).

  \begin{thmenum}
    \thmitem{def:equivalence_modulo/congruence} For a \hyperref[def:fol_congruence]{congruence} \( {\cong} \), we say that \( x \) and \( y \) are \term[ru=сравнимые по модулу идеала (\cite[38]{Шафаревич1999ОсновныеПонятияАлгебры}), ru=congruent modulo polynomial ideal (\cite[def. 5.2.1]{CoxLittleOShea2015IdealsVarietiesAlgorithms})]{congruent modulo} \( {\cong} \).

    \thmitem{def:equivalence_modulo/subobject} \Cref{thm:normal_subgroups_and_congruences/subgroup_to_cong} demonstrates how a \hyperref[def:normal_subgroup]{normal subgroup} \( N \) of \( G \) uniquely determines a congruence \( \cong \) on \( G \).

    In this case we write \( x \cong y \pmod N \). If \( N \) is \hyperref[def:group/generated]{generated} by a single element \( n \), we can also write \( x \cong y \pmod n \).

    We proceed similarly for \hyperref[def:semiring_ideal]{ring ideals} and \hyperref[def:module/submodel]{submodules}.

    \thmitem{def:equivalence_modulo/real_number} Over the \hyperref[def:real_numbers]{real numbers}, congruences modulo subgroups are occasionally useful. For example, the \hyperref[def:angle]{angles} \( \alpha \) and \( \beta \coloneqq \alpha + 2k\pi \) correspond to the same \hyperref[def:rigid_motion/rotation]{rotation}, and we can write
    \begin{equation*}
      \alpha \cong \beta \pmod{2\pi}
    \end{equation*}
    to mean that \( \beta - \alpha \) is a multiple of \( 2\pi \).

    This should not be confused with congruence modulo the ideal \( \braket{ 2\pi } \), because over a field the ideal generated by a nonzero number is the field itself.
  \end{thmenum}
\end{definition}
\begin{comments}
  \item This is \hyperref[con:assumed_knowledge]{assumed knowledge}, so it is difficult to find definitions in the literature. For ring ideals, formal definitions are given in \cite[def. 5.2.1]{CoxLittleOShea2015IdealsVarietiesAlgorithms} and \cite[38]{Шафаревич1999ОсновныеПонятияАлгебры}.
\end{comments}

\begin{proposition}\label{thm:normal_subgroup_left_right_cosets}
  Fix a \hyperref[def:normal_subgroup]{normal subgroup} \( N \) of \( G \). Then, for any \( g \) in \( G \), we have \( gN = Ng \).
\end{proposition}
\begin{comments}
  \item The converse may fail --- see \cref{ex:quotient_by_non_normal_subgroup}.
\end{comments}
\begin{proof}
  Fix an element \( n \) of \( N \). Since \( N \) is normal, the left conjugate \( gng^{-1} \) is also in \( N \), hence \( gn \) is in \( Ng \). This shows, \( gN \subseteq Ng \).

  Conversely, every element \( ng \) of \( Ng \) is in \( gN \) because \( g^{-1} n g \) is in \( N \), which in turn implies \( gN \subseteq Ng \).

  Therefore, \( gN = Ng \).
\end{proof}

\begin{lemma}\label{thm:intersection_of_normal_subgroups}
  The \hyperref[def:basic_set_operations/intersection]{intersection} of \hyperref[def:normal_subgroup]{normal subgroups} is also normal.
\end{lemma}
\begin{proof}
  Straightforward.
\end{proof}

\begin{definition}\label{def:normal_closure}\mcite[506]{Jacobson1989BasicAlgebraII}
  We define the \term{normal closure} of a subset \( A \) of a group \( G \) is the intersection of all \hyperref[def:normal_subgroup]{normal subgroups} of \( G \) containing \( A \).
\end{definition}
\begin{comments}
  \item Because of \cref{thm:intersection_of_normal_subgroups}, \cref{thm:moore_family_closure_operator} implies that this is a \hyperref[def:moore_closure_operator]{Moore closure operator} on the lattice of subgroups of \( G \).
\end{comments}

\begin{theorem}[Homomorphism theorem for groups]\label{thm:homomorphism_theorem_for_groups}
  Any \hyperref[def:group/morphism]{group homomorphism} \( \varphi: G \to H \) induces the \hyperref[def:fol_isomorphism]{isomorphism}
  \begin{equation}\label{eq:thm:homomorphism_theorem_for_groups}
    G / \ker \varphi \cong H.
  \end{equation}
\end{theorem}
\begin{comments}
  \item The English name of this theorem is based on \cite[1.23]{LidlNiederreiter1997FiniteFields}, and on the Russian equivalents that can be found in \cite[409]{Курош1968КурсВысшейАлгебры} or \cite[\S D2.4]{Тыртышников2007ЛинейнаяАлгебра}.

  Some authors use the name \enquote{first isomorphism theorem} --- for example, \incite[thm. 2.27]{Knapp2016BasicAlgebra}, \incite[thm A-3.43]{Rotman2015AdvancedModernAlgebraPart1}.
\end{comments}
\begin{comments}
  \item This is a restatement of \cref{thm:fol_homomorphism_factorization} that uses normal subgroups rather than congruences.
\end{comments}

\begin{proposition}\label{thm:group_homomorphism_trivial_kernel}
  A group homomorphism is injective if and only if its \hyperref[def:group/kernel]{kernel} is \hyperref[def:group/trivial]{trivial}.
\end{proposition}
\begin{comments}
  \item This also holds for ring and module homomorphisms since their kernel coincides with the kernel of their additive group.

  \item \Cref{thm:fol_equational_theory_injective_homomorphisms} implies that the injective group homomorphisms are precisely the monomorphisms.
\end{comments}
\begin{proof}
  Fix a homomorphism \( \varphi: G \to H \).

  \SufficiencySubProof Suppose that \( \varphi \) is injective. Then \( \varphi(x) = e_H = \varphi(e_G) \) implies that \( x = e_G \), hence \( \ker \varphi = \set{ e_G } \).

  \NecessitySubProof Suppose that \( \ker \varphi = \set{ e_G } \) and that \( \varphi(x) = \varphi(y) \).

  Then \( e_H = \varphi(xy^{-1}) \), however \( e_G \) is the only value which \( \varphi \) sends to \( e_H \), so \( xy^{-1} = e_G \) and \( x = e_G \cdot y = y \).
\end{proof}

\begin{proposition}\label{thm:group_homomorphism_zero_cokernel}
  A group homomorphism is surjective if and only if its \hyperref[def:categorical_kernels/cokernel]{categorical cokernel} is \hyperref[def:group/trivial]{trivial}.
\end{proposition}
\begin{proof}
  Fix a homomorphism \( \varphi: G \to H \). \Cref{thm:equational_theory_zero_morphisms/cokernel} implies that the cokernel of \( \varphi \) is the quotient of \( H \) by the normal closure of \( \varphi[G] \). This quotient is trivial if and only if \( H = \varphi[G] \), that is, if and only if \( \varphi \) is surjective.
\end{proof}

\begin{theorem}[Lattice theorem for subgroups]\label{thm:lattice_theorem_for_subgroups}\mcite[prop. II.8.9]{Aluffi2009Algebra}
  Given a \hyperref[def:normal_subgroup]{normal subgroup} \( N \) of \( G \), the function \( H \mapsto H / N \) is an \hyperref[def:lattice/homomorphism]{isomorphism of complete lattices} between the \hyperref[def:lattice_ideal/principal]{principal filter} of subgroups of \( G \) containing \( N \) and the \hyperref[thm:fol_lattice_of_substructures]{lattice of subgroups} of the \hyperref[def:group/quotient]{quotient} \( G / N \).

  Furthermore, normal subgroups correspond to normal subgroups.
\end{theorem}
\begin{comments}
  \item See \fullref{thm:lattice_theorem_for_substructures} for a more verbose formulation of this theorem in a general setting.
  \item Similar results hold for \hyperref[def:ring]{ring} \hyperref[def:semiring_ideal]{ideals} --- see \fullref{thm:lattice_theorem_for_ideals} --- and for \hyperref[def:module/submodel]{submodules} --- see \fullref{thm:lattice_theorem_for_submodules}.
\end{comments}
\begin{proof}
  This theorem mostly follows from \fullref{thm:lattice_theorem_for_substructures}, but we need to first show compatibility.

  Let \( N \) be a normal subgroup of \( G \). \Cref{thm:normal_subgroups_and_congruences/subgroup_to_cong} shows that \( N \) induces a congruence \( x \cong y \) that holds if \( x^{-1} y \) is in \( N \).

  \SubProof{Proof that \( S \) is compatible with \( {\cong} \) if and only if it contains \( N \)}

  \SufficiencySubProof* Let \( S \) be a subgroup of \( G \) that is compatible with \( {\cong} \) in the sense of \eqref{eq:thm:lattice_theorem_for_substructures/compatibility}. For any \( n \) in \( N \), we have \( e \cong n \), hence \( n \) is also in \( S \).

  \NecessitySubProof* Let \( S \) be a subgroup of \( G \) containing \( N \).

  Suppose that \( x \) is in \( S \) and that \( x \cong y \). Then \( x^{-1} y \) is in \( N \), but the latter also belongs to \( S \), so \( y = (x x^{-1}) y \) is in \( S \) as well.

  \SubProof{Proof that the join of compatible subgroups is the subgroup generated by their union} This follows directly by noting that the lattice of compatible subgroups of \( G \) is a sublattice of the lattice of all subgroups of \( G \).

  \SubProof{Proof that normal subgroups correspond to normal subgroups}

  \SufficiencySubProof* Let \( S \) be a normal subgroup of \( G \) compatible with \( {\cong} \), that is, containing \( N \). We must show that the quotient \( S / {\cong} \) is normal.

  Fix any member \( g \) of \( G \) and \( s \) of \( S \). Compatibility of \( {\cong} \) with respect to the group operations implies
  \begin{equation*}
    [g] [s] [g]^{-1} = [g s g^{-1}].
  \end{equation*}

  We know that the left conjugate \( g s g^{-1} \) belongs to \( S \), hence the conjugate \( [g] [s] [g]^{-1} \) belongs to \( S / {\cong} \).

  \NecessitySubProof* Let \( Q \) be a normal subgroup of \( G / {\cong} \). Fix again members \( g \) of \( G \) and \( s \) of \( S \). Then \( [g] [s] [g]^{-1} = [g s g^{-1}] \) is a member of \( Q \), hence \( g s g^{-1} \) is a member of \( \bigcup Q \), the subgroup of \( G \) corresponding to \( Q \).

  Therefore, if \( Q \) is normal in \( G / {\cong} \), \( \bigcup Q \) is normal in \( G \).
\end{proof}

\paragraph{Lagrange's subgroup theorem}

\begin{definition}\label{def:subgroup_index}\mcite[II.8.12]{Aluffi2009Algebra}
  We define the \term{index} \( [G : H] \) of a subgroup \( H \) in \( G \) is its (\hyperref[def:cardinal]{cardinal}) number of \hyperref[def:subgroup_cosetss]{cosets}.
\end{definition}
\begin{comments}
  \item \Cref{thm:subgroup_cosets/equinumerous} shows that the left and right cosets are equinumerous, so it does not matter which of them we consider.
\end{comments}

\begin{definition}\label{def:group_order}\mcite[30]{Jacobson1985BasicAlgebraI}
  We define the \term[ru=порядок (\cite[81]{Тыртышников2017ОсновыАлгебры})]{order} \( \ord(G) \) of a group \( G \) as an alias for its \hyperref[thm:cardinality_existence]{cardinality}.
\end{definition}

\begin{theorem}[Lagrange's subgroup theorem]\label{thm:lagranges_subgroup_theorem}
  For any \hyperref[def:group/subobject]{subgroup} \( H \) of \( G \), we have the following equality:
  \begin{equation}\label{eq:thm:lagranges_subgroup_theorem/general}
    \ord(G) = \ord(H) \cdot [G : H].
  \end{equation}

  For a \hyperref[def:normal_subgroup]{normal subgroup} \( N \), the \hyperref[def:subgroup_index]{index} \( [G : H] \) is the order of the quotient \( G / N \), and we have
  \begin{equation}\label{eq:thm:lagranges_subgroup_theorem/normal}
    \ord(G) = \ord(N) \cdot \ord(G / N).
  \end{equation}
\end{theorem}
\begin{comments}
  \item This demonstrates that there exists a bijective function from the \hyperref[def:fol_direct_product]{direct product} \( H \times G / H \) and \( H \), however it may not be a group homomorphism --- see \cref{ex:thm:lagranges_subgroup_theorem/direct_product_zn}.
\end{comments}
\begin{proof}
  \Cref{thm:subgroup_cosets/equinumerous} implies that every coset of \( G \) with respect to \( H \) is equinumerous with \( H \), and by definition there is a total of \( [G : H] \) cosets. The cosets are disjoint, so the theorem follows from \cref{thm:cardinality_product_rule}.
\end{proof}

\begin{example}\label{ex:subgroups_of_integers}
  Consider the group \( \BbbZ \) of integers with respect to addition.

  Let \( 2\BbbZ \) be the subgroup of all even integers. Then both \( \BbbZ \) and \( 2\BbbZ \) are countably infinite, but their quotient group \( \BbbZ / 2\BbbZ \) has two elements --- the set \( 2\BbbZ \) of all even integers and the set \( 2\BbbZ + 1 \) of all odd integers. Generalizations of this quotient group are discussed in \cref{def:group_of_integers_modulo}. \Fullref{thm:lagranges_subgroup_theorem} holds, but it gives no insight.

  Now consider the groups \( 4\BbbZ \subseteq 2\BbbZ \subseteq \BbbZ \). Since \( 2\BbbZ \) is a subgroup of \( \BbbZ \), the quotient \( 2\BbbZ / 4\BbbZ \) must a subgroup of \( \BbbZ / 4\BbbZ \) as a consequence of \fullref{thm:lattice_theorem_for_subgroups}. We may not know the structure of the quotient groups (although we do, see \cref{def:group_of_integers_modulo}), but we know how \( 4\BbbZ \), \( 2\BbbZ \) and \( \BbbZ \) relate to each other, and we are able to determine how the quotient groups relate to each other.
\end{example}

\paragraph{Order of group elements}

\begin{definition}\label{def:group_element_order}
  We define the \term[ru=степень (\cite[86]{Тыртышников2017ОсновыАлгебры}), en=order (\cite[45]{Jacobson1985BasicAlgebraI})]{order} \( \ord(x) \) of a member \( x \) of the group \( G \) via any of the following equivalent definitions:
  \begin{thmenum}
    \thmitem{def:group_element_order/power}\mcite[45]{Knapp2016BasicAlgebra} The smallest positive integer \( n \) such that \( x^n = e \), or \( \infty \) if no such integer exists.

    \thmitem{def:group_element_order/subgroup}\mcite[45]{Jacobson1985BasicAlgebraI} The cardinality (\hyperref[def:group_order]{group order}) of the subgroup \( \braket{ x } \) \hyperref[def:group/subobject]{generated} by \( x \).

    If the order is infinite, it can only be countably infinite. Following the discussion \cref{ex:def:extended_real_numbers/group_order}, we will denote this by \( \infty \) rather than \( \omega \).
  \end{thmenum}

  Elements of order \( 2 \) are called \term[en=involution (\cite[82]{Jacobson1985BasicAlgebraI})]{involutions}.
\end{definition}
\begin{comments}
  \item The term \enquote{involution} is also used more generally for self-inverse morphisms; see \cref{def:morphism_invertibility/involution}.
\end{comments}
\begin{defproof}
  \ImplicationSubProof{def:group_element_order/power}{def:group_element_order/subgroup}

  \SubProof*{Proof of finite case} Suppose first that \( x^n = e \), and that \( n \) is the smallest positive integer with that property.

  It follows that \( e, x, x^2, x^3, \ldots, x^{n-1} \) are elements of the subgroup \( \braket{ x } \). We will show that there are no other elements.

  Indeed, if \( m \geq n \), we can use \fullref{alg:integer_division} to represent \( m \) as \( nq + r \), where \( 0 \leq r < n \). Then
  \begin{equation*}
    x^{nq+r}
    \reloset {\eqref{eq:thm:semigroup_exponentiation_properties/repeated}} =
    (x^n)^q \cdot x^r
    =
    e^q \cdot x^r
    =
    x^r.
  \end{equation*}

  Since \( r < n \), it follows that \( x^m = x^r \) is among those elements we have already enumerated.

  \SubProof*{Proof of infinite case} If \( x^n \neq e \) for every \( n \), then \( \bracks{ x } \) is infinite.

  \ImplicationSubProof{def:group_element_order/subgroup}{def:group_element_order/power}

  \SubProof*{Proof of finite case} Suppose that \( \braket{ x } \) has \( n \) elements.

  Then there exist powers \( j < i \) such that \( x^i = x^j \). By \cref{thm:def:group/equality_of_powers}, we have \( x^{i - j} = e \), so \( x \) has finite order.

  Let \( m \) be the smallest positive integer such that \( x^m = e \).
  \begin{itemize}
    \item If \( m < n \), this leaves us with less than \( n \) elements because \( x^n = x^r \), where \( r \) is the remainder of dividing \( n \) by \( m \).

    \item If \( m > n \), then \( x^m = x^r \), where \( r \) is the remainder of dividing \( n \) by \( m \). Since \( x^m = e \), then \( x^r \), contradicting the minimality of \( m \).

    \item It remains for \( n = m \) to hold.
  \end{itemize}

  \SubProof*{Proof of infinite case} Suppose that \( \braket{ x } \) has infinitely many elements.

  If for some \( i > j \) we have \( x^i = x^j \), then \cref{thm:def:group/equality_of_powers} implies that \( x^{i-j} = e \) and, by what we have already shown, \( \braket{ x } \) must be finite.

  The obtained contradiction shows that \( x^i \neq x^j \) whenever \( i \neq j \).
\end{defproof}

\begin{proposition}\label{thm:def:group_element_order}
  The \hyperref[def:group_element_order]{order} of an element in a \hi{finite} group \( G \) has the following basic properties:

  \begin{thmenum}
    \thmitem{thm:def:group_element_order/group_order} The order of any element of \( G \) is a positive integer dividing the \hyperref[def:group_order]{order} (cardinality) of \( G \).

    \thmitem{thm:def:group_element_order/unique_power} If \( x \) has order \( n \), for every nonnegative integer \( m \), there exists a unique nonnegative integer \( r < n \) such that \( x^r = x^m \). Furthermore, \( n \) divides \( m - r \).

    \thmitem{thm:def:group_element_order/neutral} If \( x \) has order \( n \), then \( x^m = e \) if and only if \( n \) divides \( m \).

    \thmitem{thm:def:group_element_order/power} If \( x \) has order \( n \), for any nonnegative integer \( m \), the order of \( x^m \) is \( n / \gcd(n, m) \).

    \thmitem{thm:def:group_element_order/inverse} The order of an element equals the order of its inverse.
  \end{thmenum}
\end{proposition}
\begin{proof}
  \SubProofOf{thm:def:group_element_order/group_order} Fix an element \( x \) of order \( m \). Then the generated subgroup \( \braket{ x } \) has \( m \) elements.

  \Fullref{thm:lagranges_subgroup_theorem} implies that, since \( \braket{ x } \) is a subgroup of \( G \), \( m \) divides \( n \).

  \SubProofOf{thm:def:group_element_order/unique_power} \Fullref{alg:integer_division} gives us unique nonnegative integers \( q \) and \( r \) such that \( r < n \) and \( m = qn + r \). Then
  \begin{equation*}
    x^m
    =
    x^{qn + r}
    \reloset {\eqref{eq:thm:semigroup_exponentiation_properties/distributive}} =
    x^{qn} \cdot x^r
    \reloset {\eqref{eq:thm:semigroup_exponentiation_properties/repeated}} =
    (x^n)^q \cdot x^r.
    =
    e^q \cdot x^r
    =
    x^r.
  \end{equation*}

  \SubProofOf{thm:def:group_element_order/neutral} If \( x^m = e \), \cref{thm:def:group_element_order/unique_power} implies that \( n \) divides \( m \). Conversely, if \( n \) divides \( m \), then
  \begin{equation*}
    x^m = (x^n)^{n/m} = e^{n/m} = e.
  \end{equation*}

  \SubProofOf{thm:def:group_element_order/power} Suppose that \( x \) has order \( n \) and fix a nonnegative integer \( m \). Denote \( \gcd(n, m) \) by \( g \) and the order of \( x^m \) --- by \( k \).

  Then
  \begin{equation*}
    (x^m)^{n / g}
    \reloset {\eqref{eq:thm:semigroup_exponentiation_properties/repeated}} =
    x^{{mn} / g}
    \reloset {\eqref{eq:thm:semigroup_exponentiation_properties/repeated}} =
    (x^n)^{m / g}
    =
    e^{m / g}
    =
    e.
  \end{equation*}

  Hence, the order \( k \) of \( x^m \) is at most \( n / g \).

  Since \( (x^m)^k = e \), \cref{thm:def:group_element_order/neutral} implies that \( n \) divides \( mk \). Obviously \( m \) also divides \( mk \), thus \( mk \) is a common multiple of \( n \) and \( m \). \Cref{thm:gcd_and_lcm} however implies that \( mn / g \) is a least common multiple, thus \( n / g \leq k \).

  Therefore, \( k = n / g \).

  \SubProofOf{thm:def:group_element_order/inverse} Suppose that \( x \) has order \( n \). Let \( m \) be the order of \( x^{-1} \). We have
  \begin{equation*}
    e = (x^{-1})^m = x^{-m} = (x^m)^{-1},
  \end{equation*}
  thus \( x^m = e \). We cannot have \( m < n \) because of the minimality of \( n \); thus, \( n = m \).
\end{proof}
