\section{Higher-order logic}\label{sec:higher_order_logic}

\begin{remark}\label{rem:higher_order_logic_and_type_theory}
  Type theory allows two unrelated approaches to \hyperref[con:predicate_logic]{predicate logic}:
  \begin{thmenum}
    \thmitem{rem:higher_order_logic_and_type_theory/terms} As mentioned in \cref{rem:type_theory}, Alonzo Church initially introduced typed \( \muplambda \)-calculus as a way to encode logical formulas via \hyperref[con:type_annotation]{type-annotated} \hyperref[def:lambda_term]{\( \muplambda \)-terms}.

    This approach was later extended by his students Leon Henkin and Peter Andrews. Andrews describes a \hyperref[con:logical_system]{logical system} \( \logic{Q}_0 \) in \cite[ch. 5]{Andrews2002Logic} based on his and Henkin's refinements. More recently, William Farmer suggests modernized variants of Church's and Andrews' systems; he briefly outlines such a system in \cite{Farmer2008STTVirtues} and \cite{Farmer1990PartialFunctionSTT}. All these systems presuppose \hyperref[con:classical_logic]{classical logic}.

    Based on the availability of \fullref{ch:lambda_calculus}, which itself grew out of Church's system, we provide a further refinement of their formulations. It is in principle not limited to classical logic, but we will not purpose such generalizations.

    Because of the variety of influences involved, we comment the origin of each definition individually. The relations of our system and those mentioned above are discussed in \cref{rem:hol_notational_shorthands}, \cref{rem:hol_formula_abbreviations} and \cref{rem:hol_axiomatic_derivations}.

    \thmitem{rem:higher_order_logic_and_type_theory/types} A more powerful approach is based on the \hyperref[con:curry_howard_correspondence]{Curry-Howard correspondence}, which hints at how \hi{types} rather than \( \muplambda \)-terms can be used to encode logical formulas.

    The correspondence itself is briefly outlined in \cref{con:curry_howard_correspondence} and formalized for \hyperref[def:propositional_formula]{propositional formulas} via \fullref{alg:type_derivation_to_proof_tree} and \fullref{alg:proof_tree_to_type_derivation}.

    We can use a \hyperref[con:syntax_fragment]{fragment} of \hyperref[def:martin_lof_type_theory]{Martin-L\"of type theory} to emulate Church's original system. We further discuss this possibility in \cref{rem:mltt_hol}.
  \end{thmenum}
\end{remark}

We give here the following definition, and then proceed to explain it:
\begin{definition}\label{def:higher_order_logic}\mimprovised
  We call \term[en=higher-order logic (\cite[267]{Farmer2008STTVirtues})]{higher-order logic} the \hyperref[def:general_logic]{general logic} specified by a (\hyperref[def:truth_value_algebra/classical]{classical}) \hyperref[def:hol_institution]{higher-order logic institution} and the corresponding \hyperref[def:hol_natural_deduction]{natural deduction system}.
\end{definition}
\begin{comments}
  \item The soundness requirement in general logics from \cref{def:general_logic} is satisfied due to \fullref{thm:hol_natural_deduction_soundness}.
\end{comments}

\paragraph{Signatures of higher-order logic}

\begin{definition}\label{def:quantifiable_type}\mimprovised
  As discussed in \cref{con:simple_type_theory/hol}, Church uses restricted \hyperref[def:simple_type]{simple types} in \cite{Church1940STT} for his formulation of higher-order logic. We will call them \term{quantifiable types} to distinguish them from other simple types we consider in this monograph.

  Fix a nonempty finite set \( \op*{Sort} \), whose elements act as \hyperref[def:simple_type]{base types}. Unsurprisingly, we will call them \term{sorts}. Unless explicitly noted otherwise, we presuppose a single sort --- the \term[en=type of individuals (\cite[56]{Church1940STT})]{type of individuals} \( \syn\iota \)\fnote{The term \enquote{individuals} is based on Russell's usage; see \cref{rem:hol_formula_order_origin}. The symbol we use to denote their type is Church's small iota from \cite[56]{Church1940STT} with a dot on top, as per \cref{rem:mathematical_logic_conventions/terminal_dots}.}.

  We also presuppose another base type --- the \term[en=type of propositions (\cite[56]{Church1940STT})]{type of propositions} \( \syn\omicron \)\fnote{\tcite{In #2, where #1}[56]{Church1940STT} pioneers a dedicated syntax for types, formulas have a dedicated type --- the \enquote{type of propositions}, denoted by \( \omicron \). It can be regarded as either a Latin or a Greek letter; here we suppose the latter. In accordance with \cref{rem:mathematical_logic_conventions/terminal_dots}, we place a dot on top.} --- that we do not consider a sort.

  The quantifiable types are then the aforementioned base types and the arrow types generated by them (without type variables, product and sum types).
\end{definition}
\begin{comments}
  \item In practice, quantifiable types depend on a \hyperref[def:hol_signature]{signature of higher-order logic}.

  \item The sorts must not include improper symbols from the \hyperref[def:simple_type_alphabet]{alphabet of simple types}, but are allowed to range over small Greek identifiers because there are no type variables to clash with. In fact, both the type of individuals and the type of propositions are small Greek identifiers.

  \item Quantifiable types are what Church originally calls \enquote{simple types}, but generalized to more than one sort.

  \item We decided to call the non-propositional base types \enquote{sorts} based on similar usage in many-sorted first-order logic, which is discussed in \cite[\S 4.4.13]{TroelstraSchwichtenberg2000BasicProofTheory} and \cite[def. 62]{GoguenBurstall1992Institutions}.

  Sorts, as defined here, are unrelated to sorts in \hyperref[def:pure_type_system]{pure type systems}, which act as \hyperref[con:type_universe]{type universes} rather than types.
\end{comments}

\begin{definition}\label{def:hol_signature}\mimprovised
  A \hyperref[con:logical_system_signature]{signature} for higher-order logic consists of the following:
  \begin{thmenum}
    \thmitem{def:hol_signature/sorts} A nonempty \hyperref[def:formal_language/alphabet]{alphabet} \( \op*{Sort} \), whose elements act as sorts of \hyperref[def:quantifiable_type]{quantifiable types}.

    These sorts must not include improper symbols from the \hyperref[def:simple_type_alphabet]{alphabet of simple types} and the \hyperref[def:inference_rule_alphabet]{alphabet of inference rules}, but are allowed to range over Greek identifiers (because we discard type variables).

    We call the resulting logical system \term[en=one-sorted (predicate calculus) (\cite[84]{Kleene2002Logic})]{one-sorted} or \term[en=many-sorted predicate logic (\cite[\S 4.4.13]{TroelstraSchwichtenberg2000BasicProofTheory})]{many-sorted} depending on the amount of sorts.

    \thmitem{def:hol_signature/nl_const} A possibly empty alphabet \( \op*{NLConst} \), whose elements act as \hyperref[def:lambda_term]{constant \( \muplambda \)-terms}. We call them \term{nonlogical constants}\fnote{We stated in \cref{rem:logical_symbol_terminology} that the separation into \enquote{logical} and \enquote{nonlogical} symbols is generally ambiguous, but in this case there will be no ambiguity because we specify the kind of entity we refer to, e.g. \enquote{nonlogical constant} rather than simply \enquote{nonlogical symbol}.}.

    We disallow the improper symbols from the \hyperref[def:lambda_term_alphabet]{alphabet of \( \muplambda \)-calculus} and the \hyperref[def:inference_rule_alphabet]{alphabet of inference rules}, as well as all Latin identifiers. We also disallow the logical constants from \cref{def:hol_typing_rules}.

    \thmitem{def:hol_signature/nl_type} A set \( \op*{NLType} \), whose elements are \hyperref[def:type_assertion]{type assertions} for the nonlogical constants. We require exactly one assertion per nonlogical constant.

    For each nonlogical constant \( c \), then, if \( c: \tau \) is its assertion, we introduce the \hyperref[def:simple_typing_rule]{simple typing rule}
    \begin{equation*}\taglabel[\ensuremath{ \logic{Ax}_c }]{inf:def:hol_signature/nl_type}
      \begin{prooftree}
        \infer0[\ref{inf:def:hol_signature/nl_type}]{ c: \tau }.
      \end{prooftree}
    \end{equation*}
  \end{thmenum}
\end{definition}
\begin{comments}
  \item We discuss some cardinality considerations in \cref{rem:language_alphabet_cardinality}; we mention here only that both alphabets are by default assumed to be finite. In accordance with \cref{rem:cardinality_auxiliary_terminology}, we use cardinality-related terminology like \enquote{finite} and \enquote{countably infinite} for the signature itself when we mean the larger of the cardinalities of the alphabets.

  \item Explicit signatures are not considered by Church nor Henkin. \incite*[\S 55]{Andrews2002Logic} distinguishes between different \enquote{languages} of his system \( \logic{Q}_0 \), where a single language consists of all well-formed formulas. So Andrews also does not consider signatures, but only the resulting sets of formulas.

  Our definition is instead (loosely) based on \bycite[270]{Farmer2008STTVirtues}:
  \begin{displayquote}
    A \textit{language} of STT is a pair \( L = (\mscrC, \tau) \) where is \( \mscrC \) is a set of symbols called \textit{constants} and \( \tau: \mscrC \to \mscrT \) is a total function. That is, a language is a set of symbols with assigned types (what computer scientists usually call a \enquote{signature}). The constants are the nonlogical primitive symbols that are used to construct the expressions of the language.
  \end{displayquote}

  Farmer assumes here that \( \syn\iota \) is the only sort, but later suggests adding other sorts (which he calls base types).
\end{comments}

\begin{remark}\label{rem:hol_notational_shorthands}
  We will use several notational shorthands for higher-order logic that emulate Church's notation from \cite{Church1940STT} (or rather Farmer's refinement from \cite{Farmer2008STTVirtues}):

  \begin{thmenum}
    \thmitem{rem:hol_notational_shorthands/variable_annotations} There is a discrepancy between Church's theory of types from \cite{Church1940STT} and modernized simple type theory as presented in \fullref{sec:simply_typed_lambda_terms}.

    Namely, as discussed in \cref{rem:typing_style}, Church attaches type annotations to every \( \muplambda \)-term. For convenience, he supposes that every type has a distinct set of variables. This is in contrast with the modernized simple type theory presented in \fullref{sec:simply_typed_lambda_terms}, where only quantifier variables are annotated.

    Namely, the Church-style term
    \begin{equation*}
      (f^{\tau \synimplies \sigma} x^\tau)^\sigma
    \end{equation*}
    consists of two variables of different types\fnote{Church placed the types as subscripts rather than as superscripts; see \cref{rem:type_theory}.}. The variables here are distinguished on the level of syntax.

    In contrast, in a modern theory we have the inference
    \begin{equation*}
      \begin{prooftree}
        \hypo{ f: \tau \synimplies \sigma }
        \hypo{ x: \tau }
        \infer2[\ref{inf:def:arrow_type/elim}]{ f x: \sigma }
      \end{prooftree}
    \end{equation*}

    Here \( f \) and \( x \) are both untyped variables that are parts of type assertions. As long as we do not care about the inference tree, we can express the above as the sequent
    \begin{equation*}
      (f: \tau \synimplies \sigma, x: \tau) \vdash f x: \sigma.
    \end{equation*}

    In \cref{def:hol_expression}, we define an expression of higher-order logic as a triple \( (T, M, \tau) \), where \( T \vdash M: \tau \) and all variables of \( T \) are free in \( M \). \Cref{thm:hol_expression_type_uniqueness} shows that, as long as \( M \) is \hyperref[def:typability]{typable} in \( T \), \( \tau \) is uniquely determined by \( T \) and \( M \). This justifies visually attaching a type to each free variable (somewhat resembling Church's notation):
    \begin{equation*}
      f^{\tau \synimplies \sigma} x^\tau.
    \end{equation*}

    A similar approach is used by \incite[def. 2A7]{Hindley1997BasicSTT}.

    \thmitem{rem:hol_notational_shorthands/arrow_types} Another case where we deviate from Church's notation on purpose is arrow types --- we denote by \( \tau \synimplies \sigma \) what Church denotes by \( \sigma\tau \).

    As discussed in \cref{rem:arrow_type_name}, the arrow notation is preferred by modern authors.

    \thmitem{rem:hol_notational_shorthands/placeholder_annotations} Similarly, we have defined typed \( \muplambda \)-term \hyperref[con:schemas_and_instances]{schemas} in \cref{def:lambda_term_schema} so that neither variable schemas nor placeholders have type annotations attached.

    Since there are no type contexts for schemas, in the object language we instead place type annotations on all variables and placeholders, and disallow the same variable or placeholder with different annotations.

    For example, the schema \( F^{\tau \synimplies \sigma} x^\tau \) designates the types of the placeholder \( F \) and free variable \( x \). For bound variables, we avoid repeating the type annotation as a metalingual convention. So we write \( \qabs {x^\tau} x^\tau \) as \( \qabs {x^\tau} x \).

    Another convention we will use is to denote formula placeholders --- those of type \( \syn\omicron \) --- via letters like \( \varphi \), \( \psi \) and \( \theta \) (without superscripts), in accordance with \cref{rem:mathematical_logic_conventions/greek_alphabet}.

    \thmitem{rem:hol_notational_shorthands/abbreviations} In order to express logical constructs, Church suggests starting with several constant \( \muplambda \)-terms and \hyperref[con:syntactic_abbreviation]{abbreviating} other \( \muplambda \)-terms to obtain a more familiar logical syntax.

    There are thus two languages involved --- the familiar syntax of simply typed \( \muplambda \)-terms, as well as a dedicated syntax for predicate logic. These are appropriately viewed as distinct layers rather than alternatives, since both the syntax and semantics of the logical formulas depends on the underlying \( \muplambda \)-terms, while the usability of the system as an \hyperref[def:general_logic]{general logic} relies on the upper \enquote{logical} layer.

    We describe in \cref{rem:hol_formula_abbreviations} how different authors offer different abbreviations leading to the same logical layer. These abbreviations depend on classical propositional equivalences like those in \cref{thm:classical_equivalences}.

    We prefer a more straightforward approach, where each connective and quantifier has its own dedicated constant --- see \cref{def:hol_typing_rules}.
  \end{thmenum}
\end{remark}

\begin{definition}\label{def:hol_signature_category}\mimprovised
  For a fixed \hyperref[def:grothendieck_universe]{Grothendieck universe} \( \mscrU \), assumed by default to be the universe \hyperref[def:universe_of_hereditary_finite_sets]{\( V_\omega \)} of hereditarily finite sets, we define the \hyperref[def:category]{category} \( \ucat{Sign} \) of \( \mscrU \)-small \hyperref[def:hol_signature]{higher-order signatures} as follows:
  \begin{thmenum}
    \thmitem{def:hol_signature_category/objects} The objects are, unsurprisingly, the \( \mscrU \)-small signatures.

    \thmitem{def:hol_signature_category/morphisms} A signature morphism \( t: \Sigma \to \Sigma' \) maps the symbols in \( \Sigma \) to those in \( \Sigma' \) as follows:
    \begin{thmenum}
      \thmitem{def:hol_signature_category/morphisms/sort} A sort in \( \Sigma \) must map to a sort in \( \Sigma' \).

      \thmitem{def:hol_signature_category/morphisms/nl_const} A nonlogical constant of type \( \tau \) in \( \Sigma \) must map to a \hi{compatible} nonlogical constant in \( \Sigma' \), in which the sorts in \( \tau \) are substituted according to \( t \).
    \end{thmenum}

    \thmitem{def:hol_signature_category/composition} The composition of morphisms is their set-theoretic composition.

    \thmitem{def:hol_signature_category/identity} The identity on \( \Sigma \) is simply the set-theoretic identity.
  \end{thmenum}
\end{definition}
\begin{comments}
  \item We simply generalize the categories of first-order signatures from \cref{def:fol_signature_category}.
\end{comments}

\paragraph{Well-formed expressions}

\begin{definition}\label{def:hol_expression}\mimprovised
  Over a fixed \hyperref[def:hol_signature]{signature} \( \Sigma \) of higher-order logic, suppose that the \( \muplambda \)-term \( M \) is \hyperref[def:typability]{typable} in the \hyperref[def:type_context]{type context} \( T \) with respect to the arrow typing rules \ref{inf:def:arrow_type/elim} and \ref{inf:def:arrow_type/intro/explicit}, the family of constant typing rules \ref{inf:def:hol_signature/nl_type}, and the formula formation that we will discuss in \cref{def:hol_typing_rules}.

  Based on our discussion in \cref{rem:hol_notational_shorthands/variable_annotations}, we call the triple \( (T, M, \tau) \) a \term{well-formed expression} of higher-order logic if all variables in \( T \) are free in \( M \). \Cref{thm:hol_expression_type_uniqueness} shows that \( \tau \) is uniquely determined by \( M \) and \( T \).

  As per \cref{rem:hol_notational_shorthands/placeholder_annotations}, we annotate metalingual placeholders for expressions, so \( M^\tau \) denotes an expression of type \( \tau \).

  We denote the set of all expressions over \( \Sigma \) by \( \op*{Expr}_\Sigma \).

  \begin{thmenum}[resume=def:hol_expression]
    \thmitem{def:hol_expression/formula} If \( \tau \) is the type of propositions, we also call the expression a \term{formula}.

    As per \cref{rem:hol_notational_shorthands/placeholder_annotations}, we use the familiar symbols like \( \varphi \), \( \psi \) and \( \theta \) (without superscripts) for denoting formulas. \Cref{def:hol_typing_rules} lists formation rules for formulas.

    We denote the set of all formulas over \( \Sigma \) by \( \op*{Form}_\Sigma \).

    \thmitem{def:hol_expression/predicate} If \( \tau = \sigma_1 \synimplies \cdots \synimplies \sigma_n \synimplies \syn\omicron \), we instead call the expression a \term[en=predicate (\cite[270]{Farmer2008STTVirtues})]{predicate}.

    Predicates may be full-fledged syntactic statements, but may also be variables annotated with an appropriate type.
  \end{thmenum}
\end{definition}
\begin{comments}
  \item Our usage of the phrase \enquote{well-formed expression} here is aligned with the informal general usage described in \cref{con:expression}.

  \begin{itemize}
    \item \incite{Church1940STT} uses \enquote{well-formed formula}, as does \incite[211]{Andrews2002Logic}. We want to generalize the terminology of first-order logic, so we only use \enquote{formula} for expressions inhabiting type of propositions.

    \item \incite[270]{Farmer2008STTVirtues} uses \enquote{expression}, which we find too general given that we allow expressions to be ill-formed.

    \item Since the crucial component of an expression is a \( \muplambda \)-term, to avoid confusion we refrain from using \enquote{term} for anything else (even though it seems fitting to generalize first-order terms).
  \end{itemize}
\end{comments}

\begin{definition}\label{def:hol_closed_expression}
  As in the case of ordinary \hyperref[def:lambda_term]{\( \muplambda \)-terms}, if a \hyperref[def:hol_expression]{higher-order logic expression} has no \hyperref[def:lambda_variable_freeness]{free variables}, we say that it is \term[en=closed expression (\cite[270]{Farmer2008STTVirtues})]{closed}.

  As in \hyperref[def:first_order_logic]{first-order logic}, we refer to closed formulas as \term{sentences}\fnote{We will not call them \enquote{combinators} even though they technically are uniquely determined by combinators.}.
\end{definition}

\begin{remark}\label{rem:quantifiable_type_uncurrying}
  For \hyperref[def:hol_expression]{higher-order expressions}, we will find it convenient to utilize, within the \hyperref[con:metalogic]{metalanguage}, the different function application notations from \cref{def:function_application_notation}.

  For the term \( M^\tau \), where \( \tau = \sigma_1 \synimplies \cdots \synimplies \sigma_n \synimplies \theta \) and \( \theta \) is atomic, we will find it convenient to write \( M(x_1, \ldots, x_n) \) rather than the conventional \( \muplambda \)-term application syntax \( M x_1 \ldots x_n \), which itself is a simplification of the notation with explicit parentheses, \( (\ldots ((M x_1) x_2) \ldots x_n) \). This acts as \hyperref[def:function_currying]{uncurrying} on the level of syntax.

  Of course, most constants we will encounter will be non-letter symbols denoting binary operations, and for them we will prefer infix notation.
\end{remark}

\begin{definition}\label{def:hol_typing_rules}
  In higher-order logic, we presuppose several \hyperref[def:lambda_term]{constant \( \muplambda \)-terms}, which we call \term{logical constants}, as well as associated \hyperref[def:simple_typing_rule]{simple typing rules}.

  \begin{thmenum}
    \thmitem{def:hol_typing_rules/prop} The simplest are the rules emulating propositional logic. These are based on the constants \( \synH_\top \), \( \synH_\bot \), \( \synH_\neg \), as well as \( \synH_{\syncirc} \) for every \hyperref[def:propositional_alphabet/connectives]{propositional connective} \( {\syncirc} \):
    \columnratio{0.2,0.2,0.2,0.3}
    \begin{paracol}{4}
      \begin{nthcolumn}{0}
        \ParacolAlignmentHack
        \begin{equation*}\taglabel[\ensuremath{ \logic{H}_\top }]{inf:def:hol_typing_rules/prop/top}
          \begin{prooftree}
            \infer0[\ref{inf:def:hol_typing_rules/prop/top}]{ \synH_\top: \syn\omicron }
          \end{prooftree}
        \end{equation*}
      \end{nthcolumn}

      \begin{nthcolumn}{1}
        \ParacolAlignmentHack
        \begin{equation*}\taglabel[\ensuremath{ \logic{H}_\bot }]{inf:def:hol_typing_rules/prop/bot}
          \begin{prooftree}
            \infer0[\ref{inf:def:hol_typing_rules/prop/bot}]{ \synH_\bot: \syn\omicron }
          \end{prooftree}
        \end{equation*}
      \end{nthcolumn}

      \begin{nthcolumn}{2}
        \ParacolAlignmentHack
        \begin{equation*}\taglabel[\ensuremath{ \logic{H}_\neg }]{inf:def:hol_typing_rules/prop/neg}
          \begin{prooftree}
            \infer0[\ref{inf:def:hol_typing_rules/prop/neg}]{ \synH_\neg: \syn\omicron \synimplies \syn\omicron }
          \end{prooftree}
        \end{equation*}
      \end{nthcolumn}

      \begin{nthcolumn}{3}
        \ParacolAlignmentHack
        \begin{equation*}\taglabel[\ensuremath{ \logic{H}_{\syncirc} }]{inf:def:hol_typing_rules/prop/conn}
          \begin{prooftree}
            \infer0[\ref{inf:def:hol_typing_rules/prop/conn}]{ \synH_{\syncirc}: \syn\omicron \synimplies \syn\omicron \synimplies \syn\omicron }
          \end{prooftree}
        \end{equation*}
      \end{nthcolumn}
    \end{paracol}
    \columnratio{}

    We will use \hyperref[con:syntactic_abbreviation]{metalingual abbreviations} to emulate the familiar syntax of predicate logic. For example, we will abbreviate \( \synH_\top \) as \( \syntop \) and \( \synH_\neg \varphi \) as \( \synneg \varphi \), but we will retain \( \synH_\neg \) as a standalone expression.

    \thmitem{def:hol_typing_rules/equality} For equality, we will need a constant \( \synH_= \) and the rule
    \begin{equation*}\taglabel[\ensuremath{ \logic{H}_= }]{inf:def:hol_typing_rules/eq}
      \begin{prooftree}
        \hypo{ \synM: \syn\tau }
        \hypo{ \synN: \syn\tau }
        \infer2[\ref{inf:def:hol_typing_rules/eq}]{ \synH_= \synM \synN: \syn\omicron }
      \end{prooftree}
    \end{equation*}

    We abbreviate \( \synH_= \synM \synN \) as \( \synM^\syn\tau \syneq \synN^\syn\tau \).

    Note that, unlike in \ref{inf:def:hol_typing_rules/prop/conn}, \( \synH_= \) is not by itself typable. We could assign it a type like \( \synH_=:\syn\tau \synimplies \syn\tau \synimplies \syn\omicron \), but then \( \synH_= \) would inhabit more than one type, which would break \cref{thm:hol_expression_type_uniqueness}.

    \thmitem{def:hol_typing_rules/quant} Finally, for every quantifier \( Q \), we need the constant \( \synH_Q \) and the rule
    \begin{equation*}\taglabel[\ensuremath{ \logic{H}_Q }]{inf:def:hol_typing_rules/quant}
      \begin{prooftree}
        \hypo{ \synx: \syn\tau }
        \hypo{ \syn\varphi: \syn\omicron }
        \infer2[\ref{inf:def:hol_typing_rules/quant}]{ \synH_Q (\qabs {\synx^{\syn\tau}} \syn\varphi): \syn\omicron }
      \end{prooftree}
    \end{equation*}

    We rely on all typable terms involving \( \synH_Q \) to have an abstraction as an argument, which allows us to abbreviate \( \synH_Q (\qabs {\synx^{\syn\tau}} \syn\varphi) \) as \( \quantifier Q {\synx^{\syn\tau}} \syn\varphi \).
  \end{thmenum}
\end{definition}
\begin{comments}
  \item As discussed in \cref{rem:hol_notational_shorthands/abbreviations}, unlike Church and his students Henkin and Andrews, we do not pursue minimalism, so we add a dedicated constant for every propositional connective.

  \item We use the symbol \( \synimplies \) both as a connective for arrow types and as a connective for formulas.
\end{comments}

\begin{proposition}\label{thm:hol_expression_type_uniqueness}
  The type of a \hyperref[def:hol_expression]{higher-order logic expression} is unique: for every expression \( (T, M, \tau) \), if \( T \vdash M: \sigma \), then \( \sigma = \tau \).
\end{proposition}
\begin{proof}
  We must extend the recursion in the proof of \cref{thm:typed_term_habitation_uniqueness} to handle the new typing rules from \cref{def:hol_typing_rules}, as well as \ref{inf:def:hol_signature/nl_type}. The actual recursive cases are straightforward.
\end{proof}

\begin{remark}\label{rem:hol_formula_abbreviations}
  The choice of \hyperref[con:primitive_notion]{primitive} constant \( \muplambda \)-terms, used to encode logical constructs, has evolved a lot since the publication of \bycite{Church1940STT}.

  Church originally used several constants --- \( N_{oo} \) for \hyperref[def:propositional_alphabet/negation]{negation}, \( A_{ooo} \) for \hyperref[def:propositional_alphabet/connectives/conjunction]{conjunction}, \( \Pi_{o(o\alpha)} \) for \hyperref[def:predicate_logic_alphabet/quantifiers/universal]{universal quantification} and \( \iota_{\alpha(o\alpha)} \) for \hyperref[con:description_operator/iota]{definite description}. The latter two are in fact families of constants depending on \( \alpha \).

  Church's student \incite{Henkin1950CompletenessInTheoryOfTypes} reused Church's constants, but later in \incite*{Henkin1963TheoryOfPropositionalTypes} he presented a formulation that used only \( Q_{o \alpha \alpha } \) for \hyperref[def:predicate_logic_alphabet/equality]{equality}, acknowledging that one such constant is needed for each type \( \alpha \). This latter formulation was restricted to propositional formulas, but later in \cite{Henkin1975Identity} Henkin extended it to all types supported by Church's original system.

  We show concrete abbreviations in \cref{tab:rem:hol_formula_abbreviations}, based on the constant \( \synH_{\syneq} \), typable via the rule \ref{inf:def:hol_typing_rules/eq}. These abbreviations depend on classical propositional equivalences like those in \cref{thm:classical_equivalences}.

  Most of these abbreviations are based on \bycite[212]{Andrews2002Logic}, with the exception of \( \syntop \), which is instead taken from \bycite[273]{Farmer2008STTVirtues}. In the abbreviation of \( \varphi \synwedge \psi \), as in \cref{def:atomic_lambda_term_substitution/sharp}, the function \( \sharp(V) \) gives the smallest identifier not in \( V \).

  \begin{table}
    \begin{center}
      \begin{tabular}{l l}
        \toprule
        \multicolumn{1}{c}{Abbreviation} & \multicolumn{1}{c}{\( \muplambda \)-term schema}                                                                                                                                                         \\
        \midrule
        \( M^\tau \syneq N^\tau \)       & \( \synH_= M^\tau N^\tau \)                                                                                                                                                                              \\
        \( \syntop \)                    & \( (\qabs {\synp^{\syn\omicron}} \synp) \syneq (\qabs {\synp^{\syn\omicron}} \synp) \)                                                                                                                   \\
        \( \synbot \)                    & \( (\qabs {\synp^{\syn\omicron}} \synp) \syneq (\qabs {\synp^{\syn\omicron}} \syntop) \)                                                                                                                 \\
        \( \synneg \varphi \)            & \( \varphi \syneq \synbot \)                                                                                                                                                                             \\
        \( \varphi \synwedge \psi \)     & \( (\qabs {f^{\syn\omicron \synimplies \syn\omicron \synimplies \syn\omicron}} f \syntop \syntop) \syneq (\qabs {f^{\syn\omicron \synimplies \syn\omicron \synimplies \syn\omicron}} f \varphi \psi) \), \\
                                         & \quad where \( f = \sharp(\op*{Free}(\varphi) \cup \op*{Free}(\psi)) \)                                                                                                                                  \\
        \( \varphi \synvee \psi \)       & \( \synneg (\synneg \varphi \synwedge \synneg \psi) \)                                                                                                                                                   \\
        \( \varphi \synimplies \psi \)   & \( \varphi \syneq \varphi \synwedge \psi \)                                                                                                                                                              \\
        \( \varphi \syniff \psi \)       & \( \varphi \syneq \psi \)                                                                                                                                                                                \\
        \( \qforall {x^\tau} \varphi \)  & \( (\qabs {x^\tau} \varphi) \syneq (\qabs {x^\tau} \syntop) \)                                                                                                                                           \\
        \( \qexists {x^\tau} \varphi \)  & \( \synneg (\qforall {x^\tau} \synneg \varphi) \)                                                                                                                                                        \\
        \bottomrule
      \end{tabular}
    \end{center}

    \caption{Encoding \hyperref[con:predicate_logic]{predicate logic} in \hyperref[def:higher_order_logic]{simply typed higher-order logic}.}\label{tab:rem:hol_formula_abbreviations}
  \end{table}

  \incite*[\S 51]{Andrews2002Logic} presents a refinement of the systems of Church, Henkin and himself. Andrews calls this system \( \logic{Q}_0 \). It is based on two families of primitive constant \( \muplambda \)-terms --- \( Q_{o \alpha \alpha} \) for formal equality and \( \iota_{\iota (o \iota)} \) for definite description. \incite{Farmer2008STTVirtues} instead uses dedicated syntax for equality and definite descriptions.
\end{remark}

\begin{remark}\label{rem:mltt_hol}
  We mentioned in \cref{rem:mltt_curry_howard/arrow_types} that \hyperref[def:martin_lof_type_theory]{Martin-L\"of type theory} can be used as an alternative syntax for higher-order logic.

  This would require us to consider two \hyperref[def:mltt_signature/universe]{type universes} --- \( \syn\BbbQ \) for quantifiable types and \( \syn\omicron \) for formulas, and to restrict the types of bound variables to \( \syn\BbbQ \).

  Several subtleties arise (all of which are easily fixable):
  \begin{thmenum}
    \thmitem{rem:mltt_hol/arrows} The arrow type \( \varphi \synimplies \psi \) inhabiting \( \syn\iota \) is an abbreviation for \( \qprod {x^\varphi} \psi \), where \( \psi \) does not depend on \( x \). We cannot require \( \varphi \) to be a quantifiable type in this situation since it must be a formula.

    One way out of this situation is to regard \( {\synimplies} \) (and also \( {\syntimes} \) and \( {\synplus} \)) as dedicated syntactic entries (as in simple type theory) rather than as abbreviations of dependent products or sums.

    \thmitem{rem:mltt_hol/judgmental_equality} We consider dependent types up to \hyperref[con:equality]{judgmental equality} (encompassing \hyperref[def:beta_eta_reduction]{\( \beta\eta \)-equivalence}), while in logic we regard \( \qforall {x^\tau} \varphi \) and \( \qforall {y^\tau} \varphi[y \mapsto x] \) as distinct.

    So, we must disregard implicit judgmental equality.

    \thmitem{rem:mltt_hol/rules} Some \hyperref[con:typing_rule]{typing rules} like \ref{inf:def:dependent_product/intro} transfer directly to their counterpart \ref{inf:def:hol_quantifier_rules/eigenvariables/forall_intro}, while other like \ref{inf:def:dependent_sum/elim} differ from \ref{inf:def:hol_quantifier_rules/eigenvariables/exists_elim} in subtle ways --- see \cref{rem:dependent_products_and_forall_quantifier_rules}.

    Finally, some logical rules like those in \cref{def:hol_equality_rules} have no type-theoretic counterpart.

    This requires us to state new rules rather than reuse those from \fullref{sec:dependent_types}.
  \end{thmenum}
\end{remark}

\paragraph{Henkin semantics}

\begin{remark}\label{rem:henkin_semantics}
  The general constructions of term models for \fullref{thm:fol_model_existence_theorem} and \fullref{thm:hol_model_existence_theorem} are due to Henkin, as discussed in \cref{rem:fol_henkin_theories} and \cref{rem:hol_henkin_theories}. For this reason, certain authors like \cite[xii]{Hinman2005Logic} refer to them as \enquote{Henkin models}.

  \incite*{Henkin1950CompletenessInTheoryOfTypes} generalizes \hyperref[def:fol_structure]{first-order structures} to \hyperref[def:higher_order_logic]{higher-order logic} and calls them \enquote{standard models} and \enquote{general models}. In accordance with \cref{rem:model_theory_structure_terminology}, in \cref{def:hol_structure} we call these generalizations \term{structures} rather than \enquote{models}.

  This leads to two completely different notions that can be referred to as \enquote{Henkin models} across the literature. We will try to be unambiguous.
\end{remark}

\begin{definition}\label{def:hol_frame}\mimprovised
  A \term[en=frame (\cite[238]{Andrews2002Logic})]{frame} over a \hyperref[def:hol_signature]{signature} \( \Sigma \) of \hyperref[def:higher_order_logic]{higher-order logic} is a \hyperref[def:indexed_family]{family} \( X = \seq{ X_\tau }_{\tau \in \op*{\BbbQ_\Sigma}} \) of sets, indexed by the set \( \BbbQ_\Sigma \) of all \hyperref[def:quantifiable_type]{quantifiable types} over \( \Sigma \), and subject to some constraints. We call \( X_\tau \) the \term{universe} or \term{domain} of type \( \tau \).

  \begin{thmenum}
    \thmitem{def:hol_frame/sorts} For every sort \( S \), the universe \( X_S \) is intended to be chosen freely.

    However, by default we require \( X_S \) to be nonempty. See \cref{rem:hol_empty_universe/semantics} for a broader discussion.

    \thmitem{def:hol_frame/propositions} For the type of propositions \( \syn\omicron \), we require that \( X_{\syn\omicron} \) is a \hyperref[def:truth_value_algebra]{truth value algebra}.

    Unless explicitly noted otherwise, we presuppose that \( X_{\syn\omicron} \) is the classical two-valued Boolean algebra \( \set{ \semtop, \sembot } \).

    \thmitem{def:hol_frame/functions} For every arrow type \( \tau \synimplies \sigma \), we require \( X_{\tau \synimplies \sigma} \) to be a set of functions from \( X_\tau \) to \( X_\sigma \) (also assumed nonempty by default).

    We call the frame \term{standard} if, for every arrow type \( \tau \synimplies \sigma \), the universe \( X_{\tau \synimplies \sigma} \) contains \hi{all} functions from \( X_\tau \) to \( X_\sigma \). If the frame is \hi{not necessarily} standard, we say that it is \term{general}.
  \end{thmenum}
\end{definition}
\begin{comments}
  \item In a standard frame, it is sufficient to specify only the universes of sorts.

  \item We adapt Andrew's definition from \cite[238]{Andrews2002Logic} to many-sorted signatures. Unlike Andrews, we refer to the individual sets as \enquote{universes} rather than \enquote{domains}; see \cref{rem:model_theory_universe_terminology}.

  Additionally, we distinguish between general and standard frames, while Andrews (and Henkin) only distinguish between general and standard \hyperref[def:hol_structure]{structures}.
\end{comments}

\begin{definition}\label{def:hol_structure}\mimprovised
  A \term{structure} \( \mscrX = (X, I) \) over a \hyperref[def:hol_signature]{signature} \( \Sigma \) of higher-order logic consists of the following:
  \begin{thmenum}
    \thmitem{def:hol_structure/frame} A \hyperref[def:hol_frame]{frame} \( X = \seq{ X_\tau }_{\tau \in \op*{\BbbQ_\Sigma}} \) over \( \Sigma \).

    We refer to the structure as \term[en=general model (\cite[84]{Henkin1950CompletenessInTheoryOfTypes})]{general} or \term[en=standard model (\cite[83]{Henkin1950CompletenessInTheoryOfTypes})]{standard} depending on whether \( X \) is.

    \thmitem{def:hol_structure/interpretation} A map \( I \), called an \term{interpretation}, that sends each \hyperref[def:hol_signature/nl_const]{non-logical constant} \( c \) of type \( \tau \) to \( X_\tau \).
  \end{thmenum}
\end{definition}
\begin{comments}
  \item The interpretation is well-defined as a function since it maps \( \op*{NLConst}_\Sigma \) to the set-theoretic union \( \bigcup_{\tau \in \op*{\BbbQ_\Sigma}} X_\tau \).

  \item We will use matching letters used for the frame and for the structure itself, i.e. \( X \) and \( \mscrX \) or \( Y \) and \( \mscrY \). This is discussed in \cref{rem:fol_structure_notation}.

  \item This definition generalizes structures as defined for first-order logic in \cref{def:fol_structure}. It is roughly based on the one by Henkin from \cite[83]{Henkin1950CompletenessInTheoryOfTypes}.

  \incite[238]{Andrews2002Logic} instead calls an \enquote{interpretation} what we call a structure, while \incite[271]{Farmer2008STTVirtues} calls it a \enquote{model}, irrespective of whether a logical theory is involved. We generally avoid calling arbitrary structures \enquote{models} without logical theories involved; see \cref{rem:model_theory_structure_terminology}.
\end{comments}

\begin{remark}\label{rem:hol_empty_universe/semantics}
  We discussed in \cref{rem:fol_empty_universe/semantics} how, over a \hyperref[def:fol_signature]{first-order signature} with individual constants, for any \hyperref[def:fol_structure]{first-order structure} \( \mscrX = (X, I) \), either the universe \( X \) must be nonempty or the interpretation \( I \) must be a partial function.

  For \hyperref[def:higher_order_logic]{higher-order logic}, this is further amplified by the requirement that all universes in the frame \( X = \seq{ X_\tau }_{\tau \in \op*{\BbbQ_\Sigma}} \) must be nonempty. An immediate peculiarity is that, if \( X_{\syn\iota} \) is empty, then the function universe \( X_{\tau \synimplies \syn\iota} \) is empty if and only if \( X_\tau \) is \hi{not} empty.

  We in first-order logic, we simply refrain from dealing with empty universes.
\end{remark}

\begin{definition}\label{def:hol_variable_assignment}\mcite[238]{Andrews2002Logic}
  A \term{variable assignment} in a given \hyperref[def:hol_frame]{frame} \( X = \seq{ X_\tau }_{\tau \in \op*{\BbbQ_\Sigma}} \) is a function \( v \) that sends each possible annotated variable \( x^\tau \) to an element of \( X_\tau \).

  For every variable \( x^\tau \) and every \( a \) from \( X_\tau \), we also define the \term{modified assignment} of \( v \) at \( x \) with \( a \):
  \begin{equation*}
    v_{x^\tau \mapsto a}(y^\tau) \coloneqq \begin{cases}
      a,         &y = x, \\
      v(y^\tau), &\T{otherwise.}
    \end{cases}
  \end{equation*}
\end{definition}
\begin{comments}
  \item We use Andrews' definition from \cite[238]{Andrews2002Logic}, but elaborate on modified assignments so that they more closely resemble their counterparts in first-order logic defined in \cref{def:fol_variable_assignment}.
\end{comments}

\begin{algorithm}[Higher-order logic denotation]\label{alg:hol_denotation}\mimprovised
  Fix a \hyperref[def:hol_structure]{structure} \( \mscrX = (X, I) \) of \hyperref[def:higher_order_logic]{higher-order logic}, as well as a \hyperref[def:hol_variable_assignment]{variable assignment} \( v \) into the frame \( X = \seq{ X_\tau }_{\tau \in \op*{\BbbQ_\Sigma}} \).

  We define the \term{denotation} \( \Bracks{ M^\tau }_\mscrX^v \) of a generic \hyperref[def:hol_expression]{expression} \( M^\tau \) as follows:
  \begin{empheq}[left={\Bracks{ M^\tau }_\mscrX^v} \coloneqq \empheqlbrace]{align}
    &I(M),                                                                                          &&M \in \op*{NLConst},                                                     \label{eq:alg:hol_denotation/nl_const} \\
    &v(M^\tau),                                                                                     &&M \in \op*{Var},                                                         \label{eq:alg:hol_denotation/var} \\
    &\parens[\big]{ a \mapsto \Bracks{ N^\rho }_\mscrX^{v_{x^\sigma \mapsto a}} },                  &&\tau = \sigma \synimplies \rho \T{and} M^\tau = \qabs {x^\sigma} N^\rho, \label{eq:alg:hol_denotation/abs} \\
    &\Bracks{ N^{\sigma \synimplies \tau} }_\mscrX^v \parens[\big]{ \Bracks{ K^\sigma }_\mscrX^v }, &&M^\tau = N^{\sigma \synimplies \tau} K^\sigma.                           \label{eq:alg:hol_denotation/app}
  \end{empheq}

  Here \eqref{eq:alg:hol_denotation/abs} determines a function whose value for \( a \in X_\sigma \) is the denotation of \( N^\rho \) with respect to \( v \) modified at \( x^\sigma \) with \( a \). This ensures that the denotation of an arrow type is a function, so in \eqref{eq:alg:hol_denotation/app} we apply the denotation of the left (function) subterm to the denotation of the right subterm.

  Based on \fullref{alg:fol_formula_denotation}, we also define denotations specifically for the logical constants from \cref{def:hol_typing_rules}:
  \begin{empheq}[left={\Bracks{c}_\mscrX} \coloneqq \empheqlbrace]{align}
    &\semtop,                              &&c = \synH_\top,       \label{eq:alg:hol_denotation/l_const/top} \\
    &\sembot,                              &&c = \synH_\bot,       \label{eq:alg:hol_denotation/l_const/bot} \\
    &p \mapsto \oline{p},                  &&c = \synH_\neg,       \label{eq:alg:hol_denotation/l_const/neg} \\
    &p \mapsto (q \mapsto (p \bincirc q)), &&c = \synH_{\syncirc}, \label{eq:alg:hol_denotation/l_const/conn} \\
    &a \mapsto (b \mapsto \delta(a, b)),   &&c = \synH_=.          \label{eq:alg:hol_denotation/l_const/eq}
  \end{empheq}

  Since we are working only with well-formed expressions, the denotations of standalone propositional constants like \( \synH_\top \) or \( \synH_\vee \) are always well-defined and do not depend on the structure \( \mscrX \) nor variable assignment \( v \).

  The denotation of \( \synH_= \) formally depends on the type of its arguments, but this is a negligible technicality since \( \synH_= \) always has two arguments in well-formed formulas.

  Quantifier formulas require special treatment. Well-formed quantifier formulas always have the form
  \begin{equation*}
    \quantifier Q {x^{\sigma}} \varphi = \synH_Q (\qabs {x^{\sigma}} \varphi).
  \end{equation*}

  The denotation of \( \synH_Q \) must be a function, but it is not clear how to define it. We can rely on its argument always being \( \Bracks{\qabs {x^\sigma} \varphi}_\mscrX^v \) for some abstraction \( \qabs {x^\sigma} \varphi \) and some assignment \( v \), which leads to the following definition:
  \begin{empheq}[left={\Bracks{c}_\mscrX^v} \coloneqq \empheqlbrace]{align}
    &\Bracks{\qabs {x^\sigma} \varphi}_\mscrX^v \mapsto \bigwedge\set[\big]{ \Bracks{\varphi}_\mscrX^{v^{x^\sigma \mapsto a}} \given* a \in X_\sigma }, &&c = \synH_\forall, \label{eq:alg:hol_denotation/l_const/forall} \\
    &\Bracks{\qabs {x^\sigma} \varphi}_\mscrX^v \mapsto \bigvee\set[\big]{ \Bracks{\varphi}_\mscrX^{v^{x^\sigma \mapsto a}} \given* a \in X_\sigma }.   &&c = \synH_\exists. \label{eq:alg:hol_denotation/l_const/exists}
  \end{empheq}
\end{algorithm}
\begin{comments}
  \item We take the definition of denotation for arbitrary terms from \cite[238]{Andrews2002Logic}. For formulas, we generalize the denotations of first-order formulas from \fullref{alg:fol_formula_denotation}.

  \item As per \cref{rem:hol_parametrized_denotation}, due to the availability of \( \muplambda \)-abstraction, we will not be interested in parameterized denotations like in \cref{def:fol_parameterized_formula_denotation}.

  \item This algorithm can be found as \identifier{math.lambda_.hol.evaluate_hol_expression} in \cite{notebook:code}.
\end{comments}

\begin{remark}\label{rem:hol_definite_description}
  Compared to Andrews and Farmer, we not only prefer the explicit typing rules from \cref{inf:def:hol_typing_rules} to the abbreviations from \cref{tab:rem:hol_formula_abbreviations}, but we also avoid the definite description operator \( \syndesc \) discussed in \cref{con:description_operator/iota}.

  Introducing this operator has some nontrivial consequences:
  \begin{thmenum}
    \thmitem{rem:hol_definite_description/typing} Analogously to how we handled quantifiers in \cref{def:hol_typing_rules/quant}, we must introduce a new logical constant \( \synH_\turnediota \), as well as a typing rule
    \begin{equation*}\taglabel[\ensuremath{ \logic{H}_\turnediota }]{inf:rem:hol_definite_description/typing}
      \begin{prooftree}
        \hypo{ \synx: \syn\tau }
        \hypo{ \syn\varphi: \syn\omicron }
        \infer2[\ref{inf:rem:hol_definite_description/typing}]{ \underbrace{\synH_{\turnediota} (\qabs {\synx^{\syn\tau}} \syn\varphi)}_{\qdesc {\synx^{\syn\tau}} \syn\varphi}: \syn\tau }
      \end{prooftree}
    \end{equation*}

    This allows us to produce a term inhabiting any type.

    \thmitem{rem:hol_definite_description/interpretation} The interpretation of a \hyperref[def:hol_structure]{structure} must account for the possibility that no unique value assigned to \( \synx \) possibly satisfies \( \syn\varphi \).

    Farmer suggests endowing each structure \( \mscrX = (X, I) \) with a \hyperref[def:choice_function]{choice function} \( e \) on the frame \( X = \seq{ X_\tau }_{\tau \in \op*{\BbbQ_\Sigma}} \) that provides us with \enquote{error values}, i.e. fallback values for cases where unique existence fails\fnote{This is yet another reason to disallow empty universes.}.

    An alternative is to leave the denotation of \( \qdesc {\synx^{\syn\tau}} \syn\varphi \) \hyperref[con:undefinedness]{possibly undefined}. This is more complicated unfortunately because any value becomes possibly undefined. \incite*{Farmer1990PartialFunctionSTT} introduces 6 additional axioms to deal with such a situation.

    \thmitem{rem:hol_definite_description/axiom} \incite*[thm. 5404]{Andrews2002Logic} proves that, if we allow definite descriptions, every finite structure is standard.

    We will sketch his proof. Consider the arrow type \( \tau \synimplies \rho \). Fix any function \( g: X_\tau \to X_\rho \). We must show that \( g \) belongs to \( X_{\tau \synimplies \rho} \).

    Let \( a_1, \ldots, a_n \) be the elements of \( X_\tau \). Let \( v \) be any assignment such that \( v(\synb_k) = a_k \) and \( v(\sync_k) = g(a_k) \) for any \( k = 1, \ldots, n \).

    Andrews then proceeds to prove that \( g \) coincides with the denotation under \( v \) of the predicate
    \begin{equation*}
      \qabs {\synx^{\syn\tau}} \qdesc {\syny^{\syn\sigma}} (\synx \syneq \synb_1 \synwedge \syny \syneq \sync_1) \synvee \cdots \synvee (\synx \syneq \synb_n \synwedge \syny \syneq \sync_n).
    \end{equation*}
  \end{thmenum}
\end{remark}

\paragraph{Institution}

\begin{remark}\label{rem:def:hol_homomorphism}
  In order to discuss many semantical concepts, we must have a notion of homomorphism. \incite*[245]{Andrews2002Logic} discusses isomorphisms, and we state his definition in \cref{def:hol_isomorphism/direct}. Unfortunately, neither he nor the other authors whose work we rely upon discuss general homomorphisms.

  So, in \cref{def:hol_homomorphism}, we try to generalize Andrews' definition by allowing the functions to not be bijective. We try to justify our definition in \cref{ex:fol_to_hol_homomorphism}, and we show how it relates to Andrews' isomorphisms in \cref{def:hol_isomorphism}.
\end{remark}

\begin{definition}\label{def:hol_homomorphism}\mimprovised
  We define a \term{homomorphism} between the \hyperref[def:hol_structure]{structures} \( \mscrX = (X, I) \) and \( \mscrY = (Y, J) \) over the same signature \( \Sigma \) as a family \( h = \seq{ h_\tau }_{\tau \in \op*{\BbbQ_\Sigma}} \) of functions, where \( h_\tau \) maps \( X_\tau \) to \( Y_\tau \), such that the following holds:
  \begin{thmenum}
    \thmitem{def:hol_homomorphism/prop} For the type of propositions, we have \( h_{\syn\omicron}(\semtop) = \semtop \) and \( h_{\syn\omicron}(\sembot) = \sembot \).

    \thmitem{def:hol_homomorphism/base} For every nonlogical constant \( (c: \tau) \), we have \( h_\tau(I(c)) = J(c) \).

    \thmitem{def:hol_homomorphism/arrow} For every function type \( \tau = \sigma \synimplies \rho \), every function \( f \) from \( X_{\sigma \synimplies \rho} \) and every value \( a \) from \( X_\sigma \), we have
    \begin{equation}\label{eq:def:hol_homomorphism/arrow}
      h_{\rho \synimplies \pi}(f) \parens[\big]{ h_\rho(a) } = h_\pi\parens[\big]{ f(a) }.
    \end{equation}
  \end{thmenum}
\end{definition}
\begin{comments}
  \item This definition is not established --- see \cref{rem:def:hol_homomorphism}.
\end{comments}

\begin{example}\label{ex:fol_to_hol_homomorphism}
  We demonstrate how \hyperref[def:fol_homomorphism]{first-order homomorphisms} related to \hyperref[def:hol_homomorphism]{higher-order homomorphisms}:
  \begin{thmenum}
    \thmitem{ex:fol_to_hol_homomorphism/fol_nullary} Consider a \hyperref[def:fol_signature]{first-order signature} with a single nullary function symbol \( c \). Consider some \hyperref[def:fol_homomorphism]{first-order homomorphism} \( h: \mscrX \to \mscrY \) between the \hyperref[def:fol_structure]{first-order structures} \( \mscrX = (X, I) \) and \( \mscrY = (Y, J) \) over \( \Sigma \). As a homomorphism, it must preserve \( c \):
    \begin{equation*}
      h(I(c)) = J(c).
    \end{equation*}

    For the \hyperref[def:hol_structure]{higher-order structures}, this condition reduces to \cref{def:hol_homomorphism/base}.

    \thmitem{ex:fol_to_hol_homomorphism/fol_unary} Now suppose that \( c \) is a unary function symbol. The condition for \( h \) to be a homomorphism becomes
    \begin{equation}\label{eq:ex:fol_to_hol_homomorphism/fol_unary/fol_homomorphism}
      h\parens[\big]{ I(c)(a) } = J(c)\parens[\big]{ h(a) }.
    \end{equation}

    Let \( h_{\syn\iota \synimplies \syn\iota} \) be \hi{some} operator sending endofunctions on \( X \) to endofunctions on \( Y \). Suppose that \( h_{\syn\iota \synimplies \syn\iota} \) satisfies \cref{def:hol_homomorphism/base}. That is, suppose that, for any \( b \) in \( Y \), we have
    \begin{equation}\label{eq:ex:fol_to_hol_homomorphism/fol_unary/intermediate}
      h_{\syn\iota \synimplies \syn\iota}\parens[\big]{ I(c) }(b) = J(c)(b).
    \end{equation}

    Then
    \begin{equation*}
      h_{\syn\iota \synimplies \syn\iota}\parens[\big]{ I(c) }\parens[\big]{ h(a) }
      \reloset{\eqref{eq:ex:fol_to_hol_homomorphism/fol_unary/fol_homomorphism}} =
      J(c)\parens[\big]{ h(a) }
      \reloset{\eqref{eq:ex:fol_to_hol_homomorphism/fol_unary/intermediate}} =
      h\parens[\big]{ I(c)(a) }.
    \end{equation*}

    Generalizing this condition yields \cref{def:hol_homomorphism/arrow}.
  \end{thmenum}
\end{example}

\begin{definition}\label{def:hol_institution}\mimprovised
  For a fixed \hyperref[def:grothendieck_universe]{Grothendieck universe}, we can define an \hyperref[def:institution]{institution} for \hyperref[def:hol_structure]{higher-order structures} and \hyperref[def:hol_homomorphism]{their homomorphisms} with componentwise composition, similarly to the institutions for \hyperref[def:first_order_logic]{first-order logic} from \cref{def:fol_institution}.

  For \fullref{thm:hol_weak_completeness}, we must additionally distinguish between institutions allowing only standard models and ones allowing general models.
\end{definition}

\begin{definition}\label{def:hol_isomorphism}
  Over the \hyperref[def:hol_signature]{signature} \( \Sigma \), we say that a \hyperref[def:hol_homomorphism]{higher-order homomorphism} \( h: \mscrX \to \mscrY \) is an \term[en=isomorphism (\cite[245]{Andrews2002Logic})]{isomorphism} if any of the hollowing equivalent conditions hold:
  \begin{thmenum}
    \thmitem{def:hol_isomorphism/categorical}\mimprovised \( h \) is a \hyperref[def:morphism_invertibility/isomorphism]{categorical isomorphism} in the \hyperref[def:hol_institution]{category of models} \( \cat{Mod}(\Sigma) \): there exists an inverse homomorphism \( g \) such that \( g \bincirc h \) is the identity in \( \mscrX \) and \( h \bincirc g \) is the identity in \( \mscrY \).

    \thmitem{def:hol_isomorphism/direct}\mcite[245]{Andrews2002Logic} All components of \( h \) are \hyperref[def:function_invertibility/bijective]{bijective}.
  \end{thmenum}
\end{definition}
\begin{comments}
  \item \Cref{def:hol_isomorphism/direct} is the only established notion for higher-order logic related to homomorphisms. The others are our generalizations. See \cref{rem:def:hol_homomorphism}.
\end{comments}
\begin{defproof}
  \EquivalenceSubProof{def:hol_isomorphism/categorical}{def:hol_isomorphism/direct} If \( g \) is the inverse of \( h \), for any quantifiable type \( \tau \), the components \( g_\tau \) and \( h_\tau \) are inverses. Thus, they must be bijective.

  \EquivalenceSubProof{def:hol_isomorphism/direct}{def:hol_isomorphism/direct} Suppose that all components of \( h \) are bijective.

  We can easily define a family
  \begin{equation*}
    h^{-1} = \seq{ h^{-1}_\tau }_{\tau \in \op*{\BbbQ_\Sigma}}
  \end{equation*}
  that can be composed componentwise with \( h \) to produce the identities on \( \mscrX \) or \( \mscrY \).

  It remains to show that \( h^{-1} \) is a homomorphism. The conditions \cref{def:hol_homomorphism/prop} and \cref{def:hol_homomorphism/base} are vacuous, so we must verify \cref{def:hol_homomorphism/arrow}.

  Fix a function \( g \) from \( Y_{\tau \synimplies \sigma} \) and some value \( b \) from \( Y_\tau \) .

  Let \( f \) be \( h^{-1}_{\tau \synimplies \sigma}(g) \) and let \( a \) be \( h^{-1}_\tau(b) \). Since \( h \) is a homomorphism, we have
  \begin{equation*}
    h_{\tau \synimplies \sigma}(f)\parens[\big]{ h_\tau(a) } = h_\sigma\parens[\big]{ f(a) },
  \end{equation*}
  whose left side we can expand to
  \begin{equation*}
    h_{\tau \synimplies \sigma}\parens[\big]{h^{-1}_{\tau \synimplies \sigma}(g)}\parens[\big]{ h_\tau\parens[\big]{ h^{-1}_\tau(b) } } = h_\sigma\parens[\big]{ f(a) },
  \end{equation*}
  and simplify to
  \begin{equation*}
    g(b) = h_\sigma\parens[\big]{ f(a) }.
  \end{equation*}

  We can then apply \( h^{-1}_\sigma \) to both sides:
  \begin{equation*}
    h^{-1}_\sigma\parens[\big]{ g(b) } = f(a).
  \end{equation*}

  Expanding the right side, we obtain
  \begin{equation*}
    h^{-1}_\sigma\parens[\big]{ g(b) } = h^{-1}_{\tau \synimplies \sigma}(g) \parens[\big]{ h^{-1}_\tau(b) }.
  \end{equation*}

  This is precisely the condition \cref{def:hol_homomorphism/arrow} that we needed to verify.
\end{defproof}

\begin{definition}\label{def:hol_semantics}\mimprovised
  By the \enquote{\hyperref[con:syntax_semantics_duality]{semantics} of higher-order logic} we mean the \hyperref[def:hol_institution]{higher-order logic institutions} and all related notions like \hyperref[def:institution/models]{models}, \hyperref[def:institution/satisfaction]{satisfaction}, \hyperref[def:institutional_entailment]{semantic entailment} and \hyperref[def:semantic_equivalence]{semantic equivalence}.

  \begin{thmenum}
    \thmitem{def:hol_semantics/model} As in the case of first-order logic in \cref{def:fol_semantics/model}, we say that the structure \( \mscrX \) is a \term{model} \hi{of} the set of \hi{closed} formulas \( \Gamma \) and write \( \mscrX \vDash \Gamma \) if and only if \( \Bracks{\varphi}_\mscrX = \semtop \) for every (closed) formula \( \varphi \) in \( \Gamma \).

    \thmitem{def:hol_semantics/satisfaction} Again, as in the case of first-order logic in \cref{def:fol_semantics/satisfaction}, if \( \Bracks{\varphi}_\mscrX^v = \semtop \) for some formula \( \varphi \) and some assignment \( v \), we say that \( v \) \term{satisfies} \( \varphi \).

    If all assignments of \( \mscrX \) satisfy \( \varphi \), we say that \( \mscrX \) itself satisfies \( \varphi \).

    \thmitem{def:hol_semantics/tautology} Based on \cref{def:tautology}, We call the closed formula \( \varphi \) a \term{tautology} if \( \varphi \) is satisfied in every model, and a \term{contradiction} if it is satisfied in no model.

    \thmitem{def:hol_semantics/entailment} We use the terminology and notation from \cref{def:institutional_entailment} regarding entailment: \( \Gamma \vDash \varphi \) if and only if \( \mscrX \) satisfies \( \varphi \) whenever it satisfies \( \Gamma \).

    Here \( \varphi \), as well as the formulas of \( \Gamma \), are assumed to be closed.

    \thmitem{def:hol_semantics/standard} We distinguish between \term{standard} and \term{general} semantics depending on whether only standard models are allowed in the institution.
  \end{thmenum}
\end{definition}

\begin{proposition}\label{thm:hol_expression_semantics_of_compatible_substitutions}
  Fix an \hyperref[def:hol_expression]{expression} \( M^\tau \) of higher-order logic and \hyperref[def:atomic_lambda_term_substitution]{atomic substitutions} \( \Bbbs \) and \( \Bbbt \). Fix also \hyperref[def:hol_variable_assignment]{variable assignments} \( v \) and \( w \) in some \hyperref[def:hol_structure]{structure} \( \mscrX = (X, I) \).

  If \( \Bracks{\Bbbs(x^\sigma)}_\mscrX^v = \Bracks{\Bbbt(x^\sigma)}_\mscrX^w \) for every \hi{free} typed variable \( x^\sigma \) in \( M^\tau \), then \( \Bracks{M^\tau[\Bbbs]}_\mscrX^v = \Bracks{M^\tau[\Bbbt]}_\mscrX^w \).
\end{proposition}
\begin{comments}
  \item An analogous statement holds for first-order terms --- see \cref{thm:fol_term_semantics_of_compatible_substitutions} --- and for first-order formulas --- see \cref{thm:fol_formula_semantics_of_compatible_substitutions}.
\end{comments}
\begin{proof}
  We will use \fullref{thm:induction_on_rooted_trees} on \( M^\tau \), simultaneously on all substitutions and assignments.
  \begin{itemize}
    \item If \( M^\tau \) is a constant (either logical or nonlogical), it is invariant under substitution, so there is nothing to prove.
    \item If \( M^\tau \) is a variable, say \( x \), by assumption
    \begin{equation*}
      \Bracks{M^\tau}_\mscrX^v
      =
      \Bbbs(x^\tau)
      =
      \Bbbt(x^\tau)
      =
      \Bracks{M^\tau}_\mscrX^w.
    \end{equation*}

    \item If \( M^\tau = N^{\sigma \synimplies \tau} K^\sigma \), where the inductive hypothesis holds for \( N^{\sigma \synimplies \tau} \) and \( K^\sigma \), then
    \begin{equation*}
      \Bracks{M^\tau}_\mscrX^v
      \reloset{\eqref{eq:alg:hol_denotation/app}} =
      \Bracks{N^{\sigma \synimplies \tau}}_\mscrX^v \parens[\big]{ \Bracks{K^\sigma}_\mscrX^v }
      \reloset{\T{ind.}} =
      \Bracks{N^{\sigma \synimplies \tau}}_\mscrX^w \parens[\big]{ \Bracks{K^\sigma}_\mscrX^w }
      \reloset{\eqref{eq:alg:hol_denotation/app}} =
      \Bracks{M^\tau}_\mscrX^w.
    \end{equation*}

    \item Finally, if \( M^\tau = \qabs {x^\sigma} N^\rho \), where \( \tau = \sigma \synimplies \rho \) and the inductive hypothesis holds for \( N^\rho \), then
    \begin{equation*}
      \Bracks{M^\tau}_\mscrX^v
      \reloset{\eqref{eq:alg:hol_denotation/abs}} =
      \parens[\Big]{ a \mapsto \Bracks{N^\rho}_\mscrX^{v_{x^\sigma \mapsto a}} }
      \reloset{\T{ind.}} =
      \parens[\Big]{ a \mapsto \Bracks{N^\rho}_\mscrX^{w_{x^\sigma \mapsto a}} }
      \reloset{\eqref{eq:alg:hol_denotation/abs}} =
      \Bracks{N^\tau}_\mscrX^w.
    \end{equation*}
  \end{itemize}
\end{proof}

\begin{theorem}[Higher-order semantic deduction theorem]\label{thm:hol_semantic_deduction_theorem}
  In any \hyperref[def:hol_institution]{higher-order logic institution}, we have
  \begin{equation*}
    \Gamma, \varphi \vDash \psi \T{if and only if} \Gamma \vDash \varphi \synimplies \psi.
  \end{equation*}
\end{theorem}
\begin{comments}
  \item As in the case of first-order logic, it is important here that the formulas are closed --- see \cref{rem:implicit_quantification_and_deduction}.

  \item This is one of several deduction theorems presented here --- see \cref{rem:deduction_theorem_list}.
\end{comments}
\begin{proof}
  Analogous to \fullref{thm:fol_semantic_deduction_theorem}.
\end{proof}

\paragraph{Natural deduction}

\begin{remark}\label{rem:hol_natural_deduction_formalization}
  We will adapt the proof system from \fullref{sec:first_order_natural_deduction} to \hyperref[def:higher_order_logic]{higher-order logic}. As in \fullref{sec:dependent_types}, and for similar reasons, we will not formalize the rule schemas and proof trees completely.

  One thing should be noted --- placeholders have a type attached, so we place the eigenvariable marker as a subscript, i.e. \( x^\tau_* \).

  An alternative proof system is discussed in \cref{rem:hol_axiomatic_derivations}.
\end{remark}

\begin{definition}\label{def:hol_natural_deduction_system}\mimprovised
  An \term{abstract higher-order natural deduction system} consists of a nonempty collection of \hyperref[def:natural_deduction_rule]{natural deduction rules} for \hyperref[def:hol_expression]{higher-order logic formulas} over the same \hyperref[def:hol_signature]{signature}. We require the rules to have distinct names, but otherwise impose no restrictions on them.
\end{definition}
\begin{comments}
  \item We do not formalize these rules via schemas, as per \cref{rem:hol_natural_deduction_formalization}.
\end{comments}

\begin{definition}\label{def:hol_natural_deduction_proof_tree}
  We can define natural deduction \hyperref[def:proof_tree]{proof trees} for \hyperref[def:higher_order_logic]{higher-order logic} based on how we defined \hyperref[def:proof_tree]{proof trees} proof trees for \hyperref[def:first_order_logic]{first-order logic} in \cref{def:fol_natural_deduction_proof_tree}.
\end{definition}

\begin{definition}\label{def:hol_natural_deduction_consequence}\mimprovised
  Based on first-order \hyperref[def:consequence_relation]{consequence relation} from \cref{def:fol_natural_deduction_consequence}, for every \hyperref[def:hol_natural_deduction_system]{higher-order natural deduction system}, we define the following relation on \hi{closed} formulas: we let \( \Gamma \vdash \varphi \) if there exists a \hyperref[def:hol_natural_deduction_proof_tree]{proof tree} with conclusion \( \varphi \) whose \hyperref[def:propositional_natural_deduction_proof_tree/open_assumptions]{open assumptions} are all in \( \Gamma \).

  In accordance with \cref{def:general_logic}, we say that \( \varphi \) is \term{derivable} from \( \Gamma \) if this relation holds.
\end{definition}

\begin{proposition}\label{thm:hol_natural_deduction_derivation_compact}
  Every \hyperref[def:hol_natural_deduction_consequence]{higher-order natural deduction consequence relation} is \hyperref[def:consequence_relation/compactness]{compact}.
\end{proposition}
\begin{comments}
  \item This is one of several compactness theorems presented here --- see \cref{rem:logical_compactness_theorems}.
\end{comments}
\begin{proof}
  A proof tree is can have at most finitely many assumptions.
\end{proof}

\begin{definition}\label{def:hol_natural_deduction}\mimprovised
  We will now extend the rules for minimal, intuitionistic and classical systems for propositional natural deduction from \cref{def:propositional_natural_deduction}.

  \begin{thmenum}
    \thmitem{def:hol_natural_deduction/equality} We add more complicated rules for \hyperref[def:predicate_logic_alphabet/equality]{equality} than those for first-order logic from \cref{def:fol_natural_deduction/equality}:
    \columnratio{0.3,0.3,0.4}
    \begin{paracol}{3}
      \begin{nthcolumn}{0}
        \ParacolAlignmentHack
        \begin{equation*}\taglabel[\ensuremath{ =_+ }]{inf:def:hol_natural_deduction/equality/intro}
          \begin{prooftree}
            \infer0[\ref{inf:def:hol_natural_deduction/equality/intro}]{ M^\tau \syneq N^\tau },
          \end{prooftree}
        \end{equation*}
        where \( M^\tau \equivrel{\beta\eta} N^\tau \).
      \end{nthcolumn}

      \begin{nthcolumn}{1}
        \ParacolAlignmentHack
        \begin{equation*}\taglabel[\ensuremath{ =_\leftrightarrow }]{inf:def:hol_natural_deduction/equality/iff}
          \begin{prooftree}
            \hypo{ \varphi \syniff \psi }
            \infer1[\ref{inf:def:hol_natural_deduction/equality/iff}]{ \varphi \syneq \psi }
          \end{prooftree}
        \end{equation*}
      \end{nthcolumn}

      \begin{nthcolumn}{2}
        \ParacolAlignmentHack
        \begin{equation*}\taglabel[\ensuremath{ =_- }]{inf:def:hol_natural_deduction/equality/elim}
          \begin{prooftree}
            \hypo{ \varphi[x^\tau \mapsto M^\tau] }
            \hypo{ M^\tau \syneq N^\tau }
            \infer2[\ref{inf:def:hol_natural_deduction/equality/elim}]{ \varphi[x^\tau \mapsto N^\tau] }
          \end{prooftree}
        \end{equation*}
      \end{nthcolumn}
    \end{paracol}
    \columnratio{}

    In the first second rule \ref{inf:def:hol_natural_deduction/equality/intro}, we utilize the fact from \fullref{alg:simply_typed_reduction} that \( \beta\eta \)-equivalent \( \muplambda \)-terms inhabit identical types.

    The additional second rule \ref{inf:def:hol_natural_deduction/equality/iff} is simply auxiliary.

    \thmitem{def:hol_natural_deduction/forall} The rules for \hyperref[def:predicate_logic_alphabet/quantifiers/universal]{universal quantifiers} are straightforward adaptations of those for first-order logic from \cref{def:fol_natural_deduction/forall}:
    \begin{paracol}{2}
      \begin{leftcolumn}
        \ParacolAlignmentHack
        \begin{equation*}\taglabel[\ensuremath{ \forall_+ }]{inf:def:hol_natural_deduction/forall/intro}
          \begin{prooftree}
            \hypo{ \varphi[x^\tau \mapsto y_*^\tau] }
            \infer1[\ref{inf:def:hol_natural_deduction/forall/intro}]{ \qforall {x^\tau} \varphi }
          \end{prooftree}
        \end{equation*}
      \end{leftcolumn}

      \begin{rightcolumn}
        \ParacolAlignmentHack
        \begin{equation*}\taglabel[\ensuremath{ \forall_- }]{inf:def:hol_natural_deduction/forall/elim}
          \begin{prooftree}
            \hypo{ \qforall {x^\tau} \varphi }
            \infer1[\ref{inf:def:hol_natural_deduction/forall/elim}]{ \varphi[x^\tau \mapsto M^\tau] }
          \end{prooftree}
        \end{equation*}
      \end{rightcolumn}
    \end{paracol}

    \thmitem{def:hol_natural_deduction/exists} Similarly, the rules for \hyperref[def:predicate_logic_alphabet/quantifiers/existential]{existential quantifiers} are adaptations of those for first-order logic from \cref{def:fol_natural_deduction/exists}:
    \begin{paracol}{2}
      \begin{leftcolumn}
        \ParacolAlignmentHack
        \begin{equation*}\taglabel[\ensuremath{ \exists_+ }]{inf:def:hol_natural_deduction/exists/intro}
          \begin{prooftree}
            \hypo{ \varphi[x^\tau \mapsto M^\tau] }
            \infer1[\ref{inf:def:hol_natural_deduction/exists/intro}]{ \qexists {x^\tau} \varphi }
          \end{prooftree}
        \end{equation*}
      \end{leftcolumn}

      \begin{rightcolumn}
        \ParacolAlignmentHack
        \begin{equation*}\taglabel[\ensuremath{ \exists_- }]{inf:def:hol_natural_deduction/exists/elim}
          \begin{prooftree}
            \hypo{ \qexists {x^\tau} \varphi }

            \hypo{ \varphi[x^\tau \mapsto y_*^\tau] }
            \infer[dashed]1{ \psi }

            \infer2[\ref{inf:def:hol_natural_deduction/exists/elim}]{ \psi }
          \end{prooftree}
        \end{equation*}
      \end{rightcolumn}
    \end{paracol}
  \end{thmenum}
\end{definition}

\begin{proposition}\label{thm:hol_natural_deduction_admissible_rules}
  In \hyperref[def:hol_natural_deduction]{minimal higher-order natural deduction}, the hollowing rules are \hyperref[con:inference_rule_admissibility]{admissible}:

  \begin{thmenum}
    \thmitem{thm:hol_natural_deduction_admissible_rules/equality} Similarly to \cref{thm:fol_natural_deduction_equality}, equality acts like an \hyperref[def:equivalence_relation]{equivalence relation} compatible with \( \muplambda \)-term construction:

    \columnratio{0.3,0.3,0.4}
    \begin{paracol}{3}
      \begin{nthcolumn}{0}
        \ParacolAlignmentHack
        \begin{equation*}\taglabel[\ensuremath{ =_R }]{inf:thm:hol_natural_deduction_admissible_rules/equality/refl}
          \begin{prooftree}
            \infer0[\ref{inf:thm:hol_natural_deduction_admissible_rules/equality/refl}]{ M^\tau \syneq M^\tau }
          \end{prooftree}
        \end{equation*}
      \end{nthcolumn}

      \begin{nthcolumn}{1}
        \ParacolAlignmentHack
        \begin{equation*}\taglabel[\ensuremath{ =_S }]{inf:thm:hol_natural_deduction_admissible_rules/equality/symm}
          \begin{prooftree}
            \hypo{ M^\tau \syneq N^\tau }
            \infer1[\ref{inf:thm:hol_natural_deduction_admissible_rules/equality/symm}]{ N^\tau \syneq M^\tau }
          \end{prooftree}
        \end{equation*}
      \end{nthcolumn}

      \begin{nthcolumn}{2}
        \ParacolAlignmentHack
        \begin{equation*}\taglabel[\ensuremath{ =_T }]{inf:thm:hol_natural_deduction_admissible_rules/equality/trans}
          \begin{prooftree}
            \hypo{ M^\tau \syneq N^\tau }
            \hypo{ N^\tau \syneq K^\tau }
            \infer2[\ref{inf:thm:hol_natural_deduction_admissible_rules/equality/trans}]{ M^\tau \syneq K^\tau }
          \end{prooftree}
        \end{equation*}
      \end{nthcolumn}
    \end{paracol}
    \columnratio{}

    \begin{paracol}{2}
      \begin{leftcolumn}
        \ParacolAlignmentHack
        \begin{equation*}\taglabel[\ensuremath{ =_{\logic{App}} }]{inf:thm:hol_natural_deduction_admissible_rules/equality/app}
          \begin{prooftree}
            \hypo{ A^{\tau \synimplies \sigma} \syneq C^{\tau \synimplies \sigma} }
            \hypo{ B^\tau \syneq D^\tau }
            \infer2[\ref{inf:thm:hol_natural_deduction_admissible_rules/equality/app}]{ AB \syneq CD }
          \end{prooftree}
        \end{equation*}
      \end{leftcolumn}

      \begin{rightcolumn}
        \ParacolAlignmentHack
        \begin{equation*}\taglabel[\ensuremath{ =_{\logic{Abs}} }]{inf:thm:hol_natural_deduction_admissible_rules/equality/abs}
          \begin{prooftree}
            \hypo{ M^\sigma \syneq N^\sigma }
            \infer1[\ref{inf:thm:hol_natural_deduction_admissible_rules/equality/abs}]{ \qabs {x^\tau} M^\sigma \syneq \qabs {x^\tau} N^\sigma }.
          \end{prooftree}
        \end{equation*}
      \end{rightcolumn}
    \end{paracol}

    \thmitem{thm:hol_natural_deduction_admissible_rules/eq_to_iff} The converse rule of \ref{inf:def:hol_natural_deduction/equality/iff} is also admissible:
    \begin{equation*}\taglabel[\ensuremath{ \leftrightarrow_= }]{inf:thm:hol_natural_deduction_admissible_rules/eq_to_iff}
      \begin{prooftree}
        \hypo{ \varphi \syneq \psi }
        \infer1[\ref{inf:thm:hol_natural_deduction_admissible_rules/eq_to_iff}]{ \varphi \syniff \psi }
      \end{prooftree}
    \end{equation*}

    \thmitem{thm:hol_natural_deduction_admissible_rules/prop_eq} Derivable propositions equal verum, negations equal falsum \hi{in intuitionistic logic}, but both cannot hold simultaneously:
    \columnratio{0.3,0.3,0.4}
    \begin{paracol}{3}
      \begin{nthcolumn}{0}
        \ParacolAlignmentHack
        \begin{equation*}\taglabel[\ensuremath{ =_\top }]{inf:thm:hol_natural_deduction_admissible_rules/prop_eq/top}
          \begin{prooftree}
            \hypo{ \varphi }
            \infer1[\ref{inf:thm:hol_natural_deduction_admissible_rules/prop_eq/top}]{ \varphi \syneq \syntop }
          \end{prooftree}
        \end{equation*}
      \end{nthcolumn}

      \begin{nthcolumn}{1}
        \ParacolAlignmentHack
        \begin{equation*}\taglabel[\ensuremath{ =_\bot }]{inf:thm:hol_natural_deduction_admissible_rules/prop_eq/bot}
          \begin{prooftree}
            \hypo{ \synneg \varphi }
            \infer1[\ref{inf:thm:hol_natural_deduction_admissible_rules/prop_eq/bot}]{ \varphi \syneq \synbot }
          \end{prooftree}
        \end{equation*}
      \end{nthcolumn}

      \begin{nthcolumn}{2}
        \ParacolAlignmentHack
        \begin{equation*}\taglabel[\ensuremath{ \bot_= }]{inf:thm:hol_natural_deduction_admissible_rules/prop_eq/both}
          \begin{prooftree}
            \hypo{ \varphi \syneq \syntop }
            \hypo{ \varphi \syneq \synbot }
            \infer2[\ref{inf:thm:hol_natural_deduction_admissible_rules/prop_eq/both}]{ \synbot }
          \end{prooftree}
        \end{equation*}
      \end{nthcolumn}
      \columnratio{}
    \end{paracol}
  \end{thmenum}
\end{proposition}
\begin{proof}
  \SubProofOf{thm:hol_natural_deduction_admissible_rules/equality} The equivalence relation rules are analogous to \cref{thm:fol_natural_deduction_equality}.

  \SubProofOf{inf:thm:hol_natural_deduction_admissible_rules/equality/app}
  \begin{equation*}
    \begin{prooftree}
      \infer0[\ref{inf:def:hol_natural_deduction/equality/intro}]{ AB \syneq AB }
      \hypo{ [A^{\tau \synimplies \sigma} \syneq C^{\tau \synimplies \sigma}]^u }
      \infer2[\ref{inf:def:hol_natural_deduction/equality/elim}]{ AB \syneq CB }

      \hypo{ [B^\tau \syneq D^\tau]^v }
      \infer2[\ref{inf:def:hol_natural_deduction/equality/elim}]{ AB \syneq CD }
    \end{prooftree}
  \end{equation*}

  \SubProofOf{inf:thm:hol_natural_deduction_admissible_rules/equality/abs} Let \( y \) be a variable not free in \( M^\sigma \) nor \( N^\sigma \). Then
  \begin{equation*}
    \begin{prooftree}
      \infer0[\ref{inf:def:hol_natural_deduction/equality/intro}]{ \parens[\big]{ \qabs {x^\tau} M^\sigma \syneq \qabs {x^\tau} y^\sigma}[y^\sigma \mapsto M^\sigma] }

      \hypo{ [M^\sigma \syneq N^\sigma]^u }
      \infer2[\ref{inf:def:hol_natural_deduction/equality/elim}]{ \parens[\big]{ \qabs {x^\tau} M^\sigma \syneq \qabs {x^\tau} y^\sigma}[y^\sigma \mapsto N^\sigma] }
    \end{prooftree}
  \end{equation*}

  \SubProofOf{inf:thm:hol_natural_deduction_admissible_rules/eq_to_iff}
  \begin{equation*}
    \begin{prooftree}
      \infer0[\ref{inf:thm:propositional_admissible_rules/self_biconditional}]{ \varphi \syniff \varphi }
      \hypo{ [\varphi \syneq \psi]^v }
      \infer2[\ref{inf:def:hol_natural_deduction/equality/elim}]{ \varphi \syniff \psi }
    \end{prooftree}
  \end{equation*}

  \SubProofOf{inf:thm:hol_natural_deduction_admissible_rules/prop_eq/top}
  \begin{equation*}
    \begin{prooftree}
      \hypo{ [\varphi]^u }
      \infer0[\ref{inf:def:propositional_natural_deduction/top/intro}]{ \syntop }
      \infer2[\ref{inf:def:propositional_natural_deduction/iff/intro}]{ \varphi \syneq \syntop }
    \end{prooftree}
  \end{equation*}

  \SubProofOf{inf:thm:hol_natural_deduction_admissible_rules/prop_eq/bot}
  \begin{equation*}
    \begin{prooftree}
      \hypo{ [\synbot]^u }
      \infer1[\ref{inf:def:propositional_natural_deduction/bot/efq}]{ \varphi }

      \hypo{ [\synneg \varphi]^v }
      \hypo{ [\varphi]^w }
      \infer2[\ref{inf:def:propositional_natural_deduction/neg/elim}]{ \synbot }

      \infer[left label={\( u, w \)}]2[\ref{inf:def:propositional_natural_deduction/iff/intro}]{ \varphi \syniff \synbot }
      \infer1[\ref{inf:def:hol_natural_deduction/equality/iff}]{ \varphi \syneq \synbot }
    \end{prooftree}
  \end{equation*}

  \SubProofOf{inf:thm:hol_natural_deduction_admissible_rules/prop_eq/both}
  \begin{equation*}
    \begin{prooftree}
      \infer0[\ref{inf:def:propositional_natural_deduction/top/intro}]{ \syntop }

      \hypo{ [\varphi \syneq \syntop]^u }
      \infer1[\ref{inf:thm:hol_natural_deduction_admissible_rules/equality/symm}]{ \syntop \syneq \varphi }

      \infer2[\ref{inf:def:hol_natural_deduction/equality/elim}]{ \varphi }

      \hypo{ [\varphi \syneq \synbot]^v }
      \infer2[\ref{inf:def:hol_natural_deduction/equality/elim}]{ \synbot }
    \end{prooftree}
  \end{equation*}
\end{proof}

\begin{remark}\label{rem:hol_axiomatic_derivations}
  We will outline the proof system of Andrews' \( \logic{Q}_0 \) from \cite[\S 51]{Andrews2002Logic} since it is the most refined of the Church-like systems listed in \cref{rem:higher_order_logic_and_type_theory/types}.

  It is based on \hyperref[def:propositional_axiomatic_derivation_system]{axiomatic derivation systems}, but instead of \ref{inf:def:axiomatic_derivation_as_natural_deduction/mp}, it features a peculiar rule:
  \begin{displayquote}
    \textit{Rule R}. From \( \mathbf{C} \) and \( \mathbf{A_\alpha} = \mathbf{B_\alpha} \) to infer the result of replacing one occurrence of \( \mathbf{A_\alpha} \) in \( \mathbf{C} \) by an occurrence of \( \mathbf{B_\alpha} \), provided that the occurrence of \( \mathbf{A_\alpha} \) in \( \mathbf{C} \) is not (an occurrence of a variable) immediately preceded by \( \muplambda \).
  \end{displayquote}

  For conditional hypotheses, Andrews further refines this rule as follows:
  \begin{displayquote}
    (Rule R') If \( \mscrH \vdash \mathbf{A_\alpha} = \mathbf{B_\alpha} \), and \( \mscrH \vdash \mathbf{C} \), and \( \mathbf{D} \) is obtained from \( \mathbf{C} \) by replacing one occurrence of \( \mathbf{A_\alpha} \) in \( \mathbf{C} \) by an occurrence of \( \mathbf{B_\alpha} \), then \( \mscrH \vdash \mathbf{A_\alpha} \), provided that the occurrence of \( \mathbf{A_\alpha} \) in \( \mathbf{C} \) is not a variable immediately preceded by \( \mathbf{\muplambda} \), and the occurrence of \( \mathbf{A_\alpha} \) in \( \mathbf{C} \) is not in a wf part \( [\mathbf{\muplambda x_\beta E_\gamma}] \) of \( \mathbf{C} \), where \( \mathbf{x_\beta} \) is free in a member of \( \mscrH \) and free in \( \mathbf{A_\alpha} = \mathbf{B_\alpha} \).
  \end{displayquote}

  We will show in \cref{ex:hol_equality_rules_and_rule_r} that this rule is a special case of rule \ref{inf:def:hol_natural_deduction/equality/elim} above.

  Note that Andrews uses the abbreviations that we discuss in \cref{tab:rem:hol_formula_abbreviations}. Andrews spends considerable effort demonstrating that part of the familiar natural deduction rules from \cref{def:propositional_natural_deduction} are \hyperref[con:inference_rule_admissibility]{admissible}.

  The axioms of the system are constructed so that they allow deriving many useful proofs. For example, axiom \( 4 \) allow performing \( \beta \)-reduction similarly to \ref{inf:def:hol_equality_rules/intro}. This axiom, along with rule \logic{R'}, allows proving \cref{thm:hol_equality_is_equivalence_relation}.

  An immediate downside to his system is that hypothesis discharging in rules such as \ref{inf:def:propositional_natural_deduction/imp/intro} requires rewriting the entire proof, akin to \fullref{alg:derivation_conclusion_hypothesis_introduction}, but more complicated.

  There are other subtleties. For example, we must be careful so that the choice of variable \( f \) from the abbreviation of \( \varphi \synwedge \psi \) so that it is compatible with substitution via rule \logic{R'}.
\end{remark}

\begin{example}\label{ex:hol_equality_rules_and_rule_r}
  Andrews' rule \logic{R'} described in \cref{rem:hol_axiomatic_derivations} is also \hyperref[con:inference_rule_admissibility]{admissible} with respect to \ref{inf:def:hol_natural_deduction/equality/elim}, but the latter is more general.

  Indeed, fix a formula \( \varphi \) and let \( \psi \) be the result of substituting a single occurrence of \( A^\tau \) for \( B^\tau \) in \( \varphi \). Rule \logic{R'} then allows us to infer \( \psi \) from the open hypotheses \( \Gamma \) as long as the substituted occurrence of \( A^\tau \) is not in the scope of a variable in
  \begin{equation*}
    \parens[\Bigg]{ \bigcup_{\chi \in \Gamma} \op*{Free}(\chi) } \cap \parens[\Bigg]{ \op*{Free}(A) \cup \op*{Free}(B) }.
  \end{equation*}

  As an example, let
  \begin{align*}
    \varphi &\coloneqq \qabs {f^{\tau \synimplies \tau \synimplies \syn\omicron}} f A^\tau A^\tau, \\
    \psi    &\coloneqq \qabs {f^{\tau \synimplies \tau \synimplies \syn\omicron}} f B^\tau A^\tau  \\
  \intertext{and}
    \theta  &\coloneqq \qabs {f^{\tau \synimplies \tau \synimplies \syn\omicron}} f x^\tau A^\tau.
  \end{align*}

  Clearly
  \begin{equation*}
    \begin{prooftree}
      \hypo{ \theta[x^\tau \mapsto A^\tau] }
      \hypo{ A^\tau \syneq B^\tau }
      \infer2[\ref{inf:def:hol_natural_deduction/equality/elim}]{ \theta[x^\tau \mapsto A^\tau] }
    \end{prooftree}
  \end{equation*}

  If \( f \) is not free in \( A^\tau \) nor \( B^\tau \), the substitution in both \( \varphi \) and \( \psi \) must use the non-renaming rule \eqref{eq:alg:lambda_term_substitution/abstraction/direct}, and thus \( \varphi = \theta[x^\tau \mapsto B^\tau] \) and \( \psi = \theta[x^\tau \mapsto B^\tau] \). In this case,
  \begin{equation*}
    \begin{prooftree}
      \hypo{ \varphi }
      \hypo{ A^\tau \syneq B^\tau }
      \infer2[\ref{inf:def:hol_natural_deduction/equality/elim}]{ \psi }
    \end{prooftree}
  \end{equation*}

  This is precisely an application of \( \logic{R'} \).

  On the other hand, if \( f \) is free in \( B^\tau \) but not \( A^\tau \), then the substitution in \( \theta[x^\tau \mapsto B^\tau] \) must use the renaming rule \eqref{eq:alg:lambda_term_substitution/abstraction/renaming}, and thus
  \begin{equation*}
    \theta[x^\tau \mapsto B^\tau] = \qabs {g^{\tau \synimplies \tau \synimplies \syn\omicron}} g B^\tau (B^\tau[f \mapsto g]),
  \end{equation*}
  where \( g \) is not free in \( B^\tau \). Then
  \begin{itemize}
    \item \( \theta[x^\tau \mapsto B^\tau] \) does not equal \( \psi \), so we cannot deduce \( \psi \) from \( \varphi \) and \( A^\tau \syneq B^\tau \) using \ref{inf:def:hol_natural_deduction/equality/elim}.

    \item \( \theta[x^\tau \mapsto B^\tau] \) is not \enquote{\( \varphi \) with a single instance of \( A^\tau \) substituted \( B^\tau \)}, and rule \logic{R'} is inapplicable.
  \end{itemize}

  Admissibility of rule \logic{R'} for arbitrary \( \varphi \) and \( \psi \) can be proven by induction on the abstract syntax tree of \( \varphi \).
\end{example}

\begin{theorem}[Higher-order syntactic deduction theorem]\label{thm:hol_syntactic_deduction_theorem}
  With respect to \hyperref[def:fol_natural_deduction]{minimal higher-order natural deduction}, we have
  \begin{equation*}
    \Gamma, \varphi \vdash \psi \T{if and only if} \Gamma \vdash \varphi \synimplies \psi.
  \end{equation*}
\end{theorem}
\begin{comments}
  \item The formulas here are assumed closed, although this is not strictly necessary.
  \item This is one of several deduction theorems presented here --- see \cref{rem:deduction_theorem_list}.
\end{comments}
\begin{proof}
  Analogous to \fullref{thm:fol_syntactic_deduction_theorem}.
\end{proof}

\begin{theorem}[Higher-order natural deduction soundness]\label{thm:hol_natural_deduction_soundness}
  \hyperref[def:hol_natural_deduction]{Classical higher-order natural deduction} is \hyperref[def:general_logic]{sound} with respect to \hyperref[def:truth_value_algebra/classical]{classical} \hyperref[def:hol_semantics]{semantics}.
\end{theorem}
\begin{comments}
  \item This provides a proof that the higher-order general logic in \cref{def:first_order_logic} is well-defined.
  \item See \cref{rem:soundness_and_completeness_theorem_list} for a list of soundness and completeness theorems.
\end{comments}
\begin{proof}
  We can proceed analogously to \fullref{thm:fol_natural_deduction_soundness} to show that, whenever a proof tree \( P \) whose open assumptions are universally valid in some \hyperref[def:hol_structure]{structure} \( \mscrX = (X, I) \), its conclusion is also universally valid in \( \mscrX \).

  We will use \fullref{thm:induction_on_rooted_trees} on \( P \), but only consider the rules that are principally different from \fullref{thm:fol_natural_deduction_soundness}.

  \SubProofOf{inf:def:hol_natural_deduction/equality/intro} Let \( M^\tau \syneq N^\tau \) be the conclusion of \( P \). Then \( M^\tau \) and \( N^\tau \) are \( \beta\eta \)-equivalent expressions.

  Since equality is an equivalence relation, there is no need to verify closure conditions for the \( \beta\eta \)-equivalence relation. It is sufficient to show that, for any assignment \( v \), \( \Bracks{M^\tau}_\mscrX^v = \Bracks{N^\tau}_\mscrX^v \) whenever \( M^\tau \redrel{\beta\eta} N^\tau \).

  We proceed via \fullref{thm:induction_on_recursively_defined_relations}:
  \begin{itemize}
    \item If \( M^\tau \redrel{\beta\eta} N^\tau \) due to \ref{inf:def:lambda_term_reduction/app_left}, then \( M^\tau = A^{\sigma \synimplies \tau} B^\sigma \) and \( N^\tau = C^{\sigma \synimplies \tau} B^\sigma \), where \( A^{\sigma \synimplies \tau} \redrel{\beta\eta} C^{\sigma \synimplies \tau} \) and the inductive hypothesis holds for the latter.

    Then
    \begin{equation*}
      \Bracks{M^\tau}_\mscrX^v
      \reloset{\eqref{eq:alg:hol_denotation/app}} =
      \Bracks{A^{\sigma \synimplies \tau}}_\mscrX^v \parens[\Big]{ \Bracks{B^\sigma}_\mscrX^v }
      \reloset{\T{ind.}} =
      \Bracks{C^{\sigma \synimplies \tau}}_\mscrX^v \parens[\Big]{ \Bracks{B^\sigma}_\mscrX^v }
      \reloset{\eqref{eq:alg:hol_denotation/app}} =
      \Bracks{N^\tau}_\mscrX^v.
    \end{equation*}

    \item If \( M^\tau \redrel{\beta\eta} N^\tau \) due to \ref{inf:def:lambda_term_reduction/app_right}, we proceed similarly.

    \item If \( M^\tau \redrel{\beta\eta} N^\tau \) due to \ref{inf:def:typed_term_reduction/abs}, then \( \tau = \sigma \synimplies \rho \) for some types \( \sigma \) and \( \rho \), and \( M^\tau = \qabs {x^\sigma} A^\rho \) and \( N^\tau = \qabs {x^\sigma} B^\rho \), where \( A^\rho = B^\rho \) and the inductive hypothesis holds for the latter.

    Then
    \begin{equation*}
      \Bracks{M^\tau}_\mscrX^v
      \reloset{\eqref{eq:alg:hol_denotation/abs}} =
      \parens[\Big]{ a \mapsto \Bracks{A^\rho}_\mscrX^{v_{x^\sigma \mapsto a}} }
      \reloset{\T{ind.}} =
      \parens[\Big]{ a \mapsto \Bracks{B^\rho}_\mscrX^{v_{x^\sigma \mapsto a}} }
      \reloset{\eqref{eq:alg:hol_denotation/abs}} =
      \Bracks{N^\tau}_\mscrX^v.
    \end{equation*}

    \item If \( M^\tau \redrel{\beta\eta} N^\tau \) due to \ref{inf:def:lambda_term_reduction/alpha}, then \( M^\tau \aequiv A^\tau \) and \( N^\tau \aequiv N^\tau \), where \( A^\tau \redrel{\beta\eta} B^\tau \) and the inductive hypothesis holds for the latter.

    We can check by induction on \( M^\tau \aequiv A^\tau \) that \( \Bracks{M^\tau}_\mscrX^v = \Bracks{A^\tau}_\mscrX^v \) since the denotations do not depend on the names of bound variables, and similarly for \( N^\tau \aequiv B^\tau \).

    Therefore,
    \begin{equation*}
      \Bracks{M^\tau}_\mscrX^v = \Bracks{A^\tau}_\mscrX^v = \Bracks{B^\tau}_\mscrX^v = \Bracks{N^\tau}_\mscrX^v.
    \end{equation*}

    \item If \( M^\tau \redrel{\beta\eta} N^\tau \) due to \ref{inf:def:typed_term_reduction/beta}, then \( M^\tau = (\qabs {x^\sigma} A^\tau) B^\sigma \) and \( N^\tau = A^\tau[x^\sigma \mapsto B^\sigma] \).

    Denote \( \Bracks{ B^\sigma }_\mscrX^v \) by \( a_0 \). We have
    \begin{balign*}
      \Bracks{ M^\tau }_\mscrX^v
      &=
      \Bracks{ (\qabs {x^\sigma} A^\tau) B^\sigma }_\mscrX^v
      \reloset{\eqref{eq:alg:hol_denotation/app}} = \\ &=
      \Bracks{ \qabs {x^\sigma} A^\tau }_\mscrX^v \parens[\big]{ \underbrace{\Bracks{ B^\sigma }_\mscrX^v}_{a_0} }
      \reloset{\eqref{eq:alg:hol_denotation/abs}} = \\ &=
      \parens[\Big]{ a \mapsto \Bracks{ A^\tau }_\mscrX^{v_{x^\sigma \mapsto a}} } (a_0)
      = \\ &=
      \Bracks{ A^\tau }_\mscrX^{v_{x^\sigma \mapsto a_0}}.
    \end{balign*}

    On the other hand, \cref{thm:hol_expression_semantics_of_compatible_substitutions} implies that
    \begin{equation*}
      \Bracks{ N^\tau }_\mscrX^v
      =
      \Bracks{ A^\tau[x^\sigma \mapsto B^\sigma] }_\mscrX^v
      =
      \Bracks{ A^\tau }_\mscrX^{v_{x^\sigma \mapsto a_0}}.
    \end{equation*}

    Therefore, as desired,
    \begin{equation*}
      \Bracks{ M^\tau }_\mscrX^v = \Bracks{ N^\tau }_\mscrX^v.
    \end{equation*}

    \item If \( M^\tau \redrel{\beta\eta} N^\tau \) due to \ref{inf:def:typed_term_reduction/eta}, then \( M^\tau = \qabs {x^\sigma} N^\tau x^\sigma \) for some variable \( x \) not free in \( N \).

    Clearly
    \begin{equation*}
      \Bracks{M^\tau}_\mscrX^v
      =
      \Bracks{ \qabs {x^\sigma} (N^\tau x^\sigma) }_\mscrX^v
      \reloset{\eqref{eq:alg:hol_denotation/abs}} =
      \parens[\Big]{ a \mapsto \Bracks{ N^\tau }_\mscrX^{v_{x^\sigma \mapsto a}} }
      \reloset{\ref{thm:lambda_substitution_noop}} =
      \Bracks{ N^\tau }_\mscrX^v.
    \end{equation*}

    \SubProofOf{inf:def:hol_natural_deduction/equality/iff} Let \( \varphi \syneq \psi \) be the conclusion of \( P \). Then the premise \( P_1 \) of \( P \) has conclusion \( \varphi \syniff \psi = \synH_\leftrightarrow \varphi \psi \).

    By the inductive hypothesis, we have \( \Bracks{ \synH_\leftrightarrow \varphi \psi }_\mscrX^v = \semtop \). But
    \begin{equation*}
      \Bracks{ \synH_\leftrightarrow \varphi \psi }_\mscrX^v
      \reloset{\eqref{eq:alg:hol_denotation/abs}} =
      \Bracks{ \synH_\leftrightarrow }_\mscrX^v \parens[\big]{ \Bracks{ \varphi }_\mscrX^v } \parens[\big]{ \Bracks{ \psi }_\mscrX^v }
      \reloset{\eqref{eq:alg:hol_denotation/l_const/conn}} =
      \Bracks{ \varphi }_\mscrX^v \leftrightarrow \Bracks{ \psi }_\mscrX^v,
    \end{equation*}
    so \cref{thm:def:heyting_algebra/eq} implies that \( \Bracks{ \varphi }_\mscrX^v = \Bracks{ \psi }_\mscrX^v \), and thus \( \Bracks{ \varphi \syneq \psi }_\mscrX^v = \semtop \).
  \end{itemize}
\end{proof}

\paragraph{Completeness}

\begin{remark}\label{rem:hol_henkin_theories}
  We will present a proof of \fullref{thm:hol_weak_completeness} based on \bycite{Henkin1950CompletenessInTheoryOfTypes}.

  The term model construction, which we present in \cref{def:hol_term_model}, is based on Henkin's work. Henkin himself bases it on earlier construction for first-order logic presented in \cite{Henkin1949FOLCompleteness}.

  We also present Henkin's construction for first-order logic --- see \cref{rem:hol_henkin_theories}.
\end{remark}

\begin{definition}\label{def:hol_theory}\mimprovised
  We restate here some definitions related to \hyperref[def:logical_theory]{logical theories}, adapted to \hyperref[def:first_order_logic]{first-order logic}.

  First and foremost, a (syntactic or semantic) \term{theory} is, as in \cref{def:general_logic_theory}, a set of sentences closed under logical consequence. For any set of sentences \( \Gamma \), we denote its consequence closure by \( \cat{Th}(\Gamma) \), and say that \( \Gamma \) \term{axiomatizes} \( \cat{Th}(\Gamma) \).

  \begin{thmenum}
    \thmitem{def:hol_theory/morphism} As in \cref{def:entailment_system_theory/morphism}, we call the \hyperref[def:hol_signature_category/morphisms]{first-order signature morphism} \( t: \Sigma \to \Sigma' \) a \term{theory morphisms} from \( (\Sigma, T) \) to \( (\Sigma', T') \) if the translation of the formulas in \( T \) belong to \( T' \).

    \thmitem{def:hol_theory/extension} As for first-order theories, we call \( T^+ \) an \term{extension} of \( T \) if \( t \) is an inclusion map that only adds new signature symbols.

    \thmitem{def:hol_theory/category} Based on theories and their morphisms, as per \cref{def:category_of_theories}, we have a category of (syntactic or semantic) theories \( \ucat{Th} \) for every \hyperref[def:grothendieck_universe]{Grothendieck universe} \( \mscrU \).

    \thmitem{def:hol_theory/conservative} Based on \cref{def:entailment_system_theory/conservative}, we say that a theory morphism \( t: (\Sigma, T) \to (\Sigma', T') \) is \term{conservative} when \( \varphi \) belongs to \( T \) if and only if \( \op*{Sen}(t)(\varphi) \) belongs to \( T' \).

    Sufficiency is obvious for both semantic entailment and natural deduction, so it usually suffices to check that \( \op*{Sen}(t)(\varphi) \in T' \) implies \( \varphi \in T \).

    \thmitem{def:hol_theory/consistent} We call a higher-order theory \term{consistent} if \( T \) contains no \hyperref[def:hol_semantics/tautology]{contradictions}, i.e. if \( T \) does not contain \( \synbot \).

    See \cref{def:fol_theory/consistent} for remarks on consistent first-order theories that are also valid here.

    \thmitem{def:hol_theory/complete} We call a higher-order theory \term{complete} if, for every sentence \( \varphi \), \( T \) contains either \( \varphi \) or \( \synneg \varphi \) (or possibly both, if the theory is inconsistent).

    See \cref{def:fol_theory/complete} for remarks on consistent first-order theories that are also valid here.
  \end{thmenum}
\end{definition}

\begin{definition}\label{def:hol_henkin_completeness}\mimprovised
  We extend the notion of Henkin-completeness from \cref{def:fol_henkin_completeness} to \hyperref[def:hol_theory]{higher-order logic theories} as follows: we say that the theory \( T \) over \( \Sigma \) is \term{Henkin complete} if, for every closed existentially quantified formula \( \qexists {x^\tau} \varphi \) over \( \Sigma \) (not necessarily in \( T \)), there exists a constant \( c \) such that
  \begin{equation}\label{eq:def:hol_henkin_completeness/axiom}
    \qexists {x^\tau} \varphi \synimplies \varphi[x^\tau \mapsto c]
  \end{equation}
  belongs to \( T \).
\end{definition}
\begin{comments}
  \item Compare this to the Henkin-completeness for first-order theories, which we define in \cref{def:fol_henkin_completeness}.
\end{comments}

\begin{definition}\label{def:hol_term_model}\mcite[thm. 1]{Henkin1950CompletenessInTheoryOfTypes}
  We adapt \hyperref[def:fol_term_model]{first-order term models} as follows. With respect to \hyperref[def:hol_natural_deduction_system]{higher-order natural deduction}, fix a \hyperref[def:hol_henkin_completeness]{Henkin-complete}, \hyperref[def:hol_theory/complete]{complete} and \hyperref[def:hol_theory/consistent]{consistent} theory \( T \) over the \hyperref[def:hol_signature]{signature} \( \Sigma \).

  For every quantifiable type \( \tau \) over \( \Sigma \), let \( E_\tau \) be the set of all closed expressions over \( \Sigma \). This set is nonempty because \( T \) contains a witness for \( \qexists {x^\tau} \syntop \).

  We will now use these sets to define a \hyperref[def:hol_frame]{frame} \( X = \seq{ X_\tau }_{\tau \in \op*{\BbbQ_\Sigma}} \) and maps \( \Phi_\tau: E_\tau \to X_\tau \) for every quantifiable type \( \tau \). Our term model will then be \( \mscrT = (X, I) \), where the interpretation \( I(c) \) of the nonlogical constant \( (c: \tau) \) is \( \Phi_\tau(c) \).

  We will additionally need the relation
  \begin{equation*}
    M^\tau \cong_\tau N^\tau \T{if and only if} (M^\tau \syneq N^\tau) \in T.
  \end{equation*}

  \Cref{thm:hol_natural_deduction_admissible_rules/equality} implies that \( {\cong_\tau} \) is an \hyperref[def:equivalence_relation]{equivalence relation} on \( E_\tau \) compatible with \( \muplambda \)-term construction. We will make sure that \( \Phi_\tau \) is compatible with \( {\cong_\tau} \), i.e. that \( M^\tau \cong_\tau N^\tau \) implies \( \Phi_\tau(M) = \Phi_\tau(N) \).

  \begin{thmenum}
    \thmitem{def:hol_term_model/prop} For type of propositions \( \syn\omicron \), the universe \( X_{\syn\omicron} \) is by definition the two-valued Boolean algebra \( \set{ \semtop, \sembot } \).

    We let
    \begin{equation*}
      \Phi_{\syn\omicron}(\varphi) \coloneqq \begin{cases}
        \semtop, &\varphi \cong_{\syn\omicron} \syntop, \\
        \sembot, &\varphi \cong_{\syn\omicron} \synbot.
      \end{cases}
    \end{equation*}

    This map is well-defined, as can be shown easily.

    Since \( T \) is complete, for every formula \( \varphi \) over \( \Sigma \), either \( T \) contains \( \varphi \) and \ref{inf:thm:hol_natural_deduction_admissible_rules/prop_eq/top} implies that \( \varphi \cong_{\syn\omicron} \syntop \), or \( T \) contains \( \varphi \) and \ref{inf:thm:hol_natural_deduction_admissible_rules/prop_eq/bot} implies that \( \varphi \cong_{\syn\omicron} \synbot \).

    Furthermore, since \( T \) is consistent, we cannot have both \( \varphi \cong_{\syn\omicron} \syntop \) and \( \varphi \cong_{\syn\omicron} \synbot \) because \ref{inf:thm:hol_natural_deduction_admissible_rules/prop_eq/both} would allow deriving \( \synbot \).

    \thmitem{def:hol_term_model/sort} For every sort \( \tau \), we define the universe \( X_\tau \) as the quotient of \( E_\tau \) by \( {\cong_\tau} \). In this case \( \Phi_\tau \) is simply the quotient map.

    \thmitem{def:hol_term_model/arrow} For the arrow type \( \tau \synimplies \sigma \), every term \( M^{\tau \synimplies \sigma} \) induces a function as follows:
    \begin{equation*}
      \Phi_{\tau \synimplies \sigma}(M) \coloneqq \parens[\Big]{ \Phi_\tau(N) \mapsto \Phi_\sigma(MN) }.
    \end{equation*}

    Well-definedness of \( \Phi_{\tau \synimplies \sigma} \) follows the consistency properties in \cref{thm:hol_natural_deduction_admissible_rules/equality}.
  \end{thmenum}
\end{definition}
\begin{comments}
  \item This construction is taken from Henkin's \cite[thm. 1]{Henkin1950CompletenessInTheoryOfTypes} and is loosely based on his earlier construction for first-order logic presented in \cite{Henkin1949FOLCompleteness}. Our term models for first-order logic are also based on Henkin's.
\end{comments}

\begin{definition}\label{def:hol_frugal_structure}\mcite[128]{Andrews2002Logic}
  We say that a \hyperref[def:hol_structure]{higher-order structure} \( \mscrX = (X, I) \) over \( \Sigma \) is \term{frugal} if the cardinality of each universe \( X_\tau \) in the \hyperref[def:hol_frame]{frame} \( X \) does not exceed the number of expressions over \( \Sigma \).
\end{definition}
\begin{comments}
  \item Compare this to \cref{def:fol_frugal_structure}, where for first-order structures frugality restricts only the universe of individuals.
\end{comments}

\begin{proposition}\label{thm:hol_term_model_is_frugal}
  \hyperref[def:hol_term_model]{Higher-order logic term models} are \hyperref[def:hol_frugal_structure]{frugal}.
\end{proposition}
\begin{proof}
  The cardinality of each universe is restricted by the number of closed expressions of the corresponding type.
\end{proof}

\begin{proposition}\label{thm:hol_frugality_characterization}
  A \hyperref[def:hol_structure]{higher-order logic structure} \( \mscrX = (X, I) \) over \( \Sigma \) is \hyperref[def:hol_frugal_structure]{frugal} if and only if, for each quantifiable type \( \tau \), we have
  \begin{equation*}
    \card(X_\tau) \leq \max\set{ \aleph_0, \kappa },
  \end{equation*}
  where \( \kappa \) is the amount of symbols in \( \Sigma \).
\end{proposition}
\begin{proof}
  Analogous to \cref{thm:fol_frugality_characterization}, with the insignificant caveat that each purely syntactic \( \muplambda \)-term possibly has multiple type contexts.
\end{proof}

\begin{theorem}[Higher-order model existence theorem]\label{thm:hol_model_existence_theorem}\mimprovised
  As in \cref{def:hol_term_model}, fix a \hyperref[def:hol_henkin_completeness]{Henkin-complete}, \hyperref[def:hol_theory/complete]{complete} and \hyperref[def:hol_theory/consistent]{consistent} theory \( T \) (with respect to \hyperref[def:hol_natural_deduction_system]{natural deduction}).

  We claim that the term model \( \mscrT \) is indeed a model of \( T \).
\end{theorem}
\begin{proof}
  We will use induction on the \hyperref[def:graph_cardinality/order]{graph order} of the \hyperref[def:lambda_term_ast]{abstract syntax tree} of the \hi{closed} expression \( M^\tau \) to show that \( \Bracks{M^\tau}_\mscrT = \Phi(M^\tau) \). Since \( \Phi_\tau(\varphi) = \semtop \) if and only if \( T \vdash \varphi \), it will follow that \( \mscrT \) is a model of \( T \).

  \begin{itemize}
    \item If \( M^\tau \) is a constant, either logical or nonlogical, its denotation is by definition \( \Phi_\tau(M) \).

    \item Since \( M^\tau \) is closed, it cannot be a variable.

    \item If \( M^\tau = N^{\sigma \synimplies \tau} K^\sigma \), where the inductive hypothesis holds for \( N^{\sigma \synimplies \tau} \) and \( K^\sigma \), then
    \begin{equation*}
      \Bracks{M^\tau}_\mscrX^v
      \reloset{\eqref{eq:alg:hol_denotation/app}} =
      \Bracks{N^{\sigma \synimplies \tau}}_\mscrX^v \parens[\big]{ \Bracks{K^\sigma}_\mscrX^v }
      \reloset{\T{ind.}} =
      \Phi_{\sigma \synimplies \tau}(N) \parens[\big]{ \Phi_\sigma(K) }
      =
      \Phi_\tau(NK).
    \end{equation*}

    \item Finally, suppose that \( M^\tau = \qabs {x^\sigma} N^\rho \), where \( \tau = \sigma \synimplies \rho \) and the inductive hypothesis holds for \( N^\rho \).

    Since \( T \) is Henkin-complete, for every closed expression \( K^\sigma \), there exists a witness \( c_K \) for the formula
    \begin{equation*}
      \qexists {y^\sigma} (y \syneq K^\sigma).
    \end{equation*}

    Then \( \Phi_\sigma(c_K) = \Phi_\sigma(K) \). Furthermore, since \( c_K \) is a constant, we have \( \Bracks{c_K^\sigma}_\mscrT = \Phi_\sigma(c_K) \).

    For any variable assignment \( v \), we have
    \begin{equation*}
      \Bracks{M^\tau}_\mscrX^v\parens[\big]{ \Phi_\sigma(K) }
      =
      \Bracks{M^\tau}_\mscrX^v\parens[\big]{ \Phi_\sigma(c_K) }
      \reloset{\eqref{eq:alg:hol_denotation/abs}} =
      \Bracks{N^\rho}_\mscrX^{v_{x^\sigma \mapsto \Phi_\sigma(c_K)}}
      \reloset{\ref{thm:hol_expression_semantics_of_compatible_substitutions}} =
      \Bracks{N^\rho[x^\sigma \mapsto c_K^\sigma]}_\mscrX^v.
    \end{equation*}

    By the inductive hypothesis, the above equals \( \Phi_\rho(N^\rho[x^\sigma \mapsto c_a^\sigma]) \). Since
    \begin{equation*}
      M^\tau c_K^\sigma \bred N^\rho[x^\sigma \mapsto c_K^\sigma],
    \end{equation*}
    via an application of \ref{inf:def:hol_natural_deduction/equality/intro} it follows that the two are congruent in \( \mscrT \), thus
    \begin{equation*}
      \Phi_\sigma(N[x^\sigma \mapsto c_K^\sigma])
      =
      \Phi_\sigma(M c_K)
      =
      \Phi_\tau(M) \parens[\big]{ \Phi_\sigma(c_K) }
      =
      \Phi_\tau(M) \parens[\big]{ \Phi_\sigma(K) }.
    \end{equation*}

    We have obtained that \( \Phi_\tau(M) = \Bracks{M^\tau}_\mscrT \), as desired.
  \end{itemize}
\end{proof}

\begin{corollary}\label{thm:hol_consistent_implies_satisfiable}
  If a set of closed formulas is \hyperref[def:hol_theory/consistent]{consistent} with respect to \hyperref[def:hol_natural_deduction_system]{classical natural deduction}, it has a \hyperref[def:hol_frugal_structure]{frugal} \hyperref[def:fol_semantics/model]{model}.
\end{corollary}
\begin{comments}
  \item This statement has an analog for both \hyperref[def:propositional_logic]{propositional logic} and \hyperref[def:first_order_logic]{first-order logic} --- see \cref{thm:propositional_consistent_implies_satisfiable} and \cref{thm:fol_consistent_implies_satisfiable}, respectively.
\end{comments}
\begin{proof}
  We have skipped several steps of the proof compared to \cref{thm:fol_consistent_implies_satisfiable} because they are similar to the case of first-order logic.

  Let \( T \coloneqq \op*{Th}(\Gamma) \) be the theory generated by \( \Gamma \). We have not defined Henkin completions, but the process is analogous to the first-order case, which we study in detail in \fullref{sec:first_order_completeness}. Anyhow, let \( T^{*\omega} \) be the similarly constructed Henkin completion of \( T \).

  We can then use an analog of \fullref{thm:fol_lindenbaums_lemma} to build an extension \( T^+ \) of \( T^{*\omega} \) that is consistent, complete and Henkin-complete. As an extension of \( T \), \( T^+ \) is \hyperref[def:hol_theory/conservative]{conservative}.

  This allows us to construct a \hyperref[def:hol_term_model]{term model} \( \mscrT \), which satisfies \( T^+ \) by \fullref{thm:hol_model_existence_theorem}. Its reduct along the inclusion \( \iota: T \to T^+ \) is thus a model of \( T \).

  Furthermore, \cref{thm:fol_term_model_is_frugal} shows that \( \mscrT \) is frugal over the Henkin completion \( \Sigma^{*\omega} \), so similarly to the first-order case we can show that it is also frugal over \( \Sigma \).
\end{proof}

\begin{theorem}[Weak higher-order completeness]\label{thm:hol_weak_completeness}
  The \hyperref[def:fol_natural_deduction]{classical higher-order natural deduction system} is \hyperref[def:general_logic/completeness]{complete} with respect to \hyperref[def:fol_semantics/standard]{general higher-order semantics}.
\end{theorem}
\begin{comments}
  \item As discussed by \incite{Henkin1950CompletenessInTheoryOfTypes}, this does not hold for standard semantics.

  \item This is the higher-order counterpart of \fullref{thm:classical_propositional_completeness} and \fullref{thm:fol_completeness}.
\end{comments}
\begin{proof}
  We will spell out the details to highlight where frugality is used.

  Suppose that \( \Gamma \vDash \varphi \). Then the set \( \Gamma \cup \set{ \neg \varphi } \) is \hyperref[def:hol_semantics/satisfaction]{unsatisfiable} because \( \Bracks{\neg \varphi}_\mscrX = \sembot \) for every model \( \mscrX \) of \( \Gamma \).

  The set \( \Gamma \cup \set{ \neg \varphi } \) must be \hyperref[def:hol_theory/consistent]{inconsistent} since otherwise \cref{thm:hol_consistent_implies_satisfiable} would imply that it has a \hi{frugal} model.

  Since is inconsistent, there exists a proof tree deriving \( \bot \) from \( \Gamma \cup \set{ \neg \varphi } \). Then we can apply \ref{inf:def:propositional_natural_deduction/bot/raa} and obtain a proof of \( \varphi \) from \( \Gamma \).
\end{proof}

\paragraph{Compactness}

\begin{theorem}[Weak higher-order semantic compactness]\label{thm:hol_weak_semantic_compactness}
  All of the hollowing equivalent statements hold for the \hyperref[def:hol_semantics/entailment]{higher-order entailment relation} for \hyperref[def:hol_semantics/standard]{general semantics} over finite signatures:
  \begin{thmenum}
    \thmitem{thm:hol_weak_semantic_compactness/relation} The relation is \hyperref[def:consequence_relation/compactness]{compact}: if \( \Gamma \vDash \varphi \), there exists a finite subset \( \Gamma_0 \) of \( \Gamma \) such that \( \Gamma_0 \vDash \varphi \).

    \thmitem{thm:hol_weak_semantic_compactness/satisfiable} A set of sentences is \hyperref[def:hol_semantics/model]{satisfiable} (has a model) if and only if it is \hyperref[def:finitely_satisfiable_set_of_sentences]{finitely satisfiable} (every finite subset has a model).

    \thmitem{thm:hol_weak_semantic_compactness/consistent} A set of sentences is \hyperref[def:hol_theory/consistent]{consistent} (does not entail \( \synbot \)) if and only if it is \hyperref[def:finitely_consistent_set_of_sentences]{finitely consistent} (no finite subset entails \( \synbot \)).
  \end{thmenum}
\end{theorem}
\begin{comments}
  \item \Cref{thm:propositional_semantic_inconsistency} gives several equivalent conditions for a set of closed formulas to be semantically consistent. Satisfiability is among them.

  \item This is one of several compactness theorems presented here --- see \cref{rem:logical_compactness_theorems}.
\end{comments}
\begin{proof}
  \Cref{thm:hol_weak_semantic_compactness/relation} Follows from \fullref{thm:hol_compactness} via \cref{thm:completeness_implies_compactness}.

  The equivalence between the different conditions can be shown as in \fullref{thm:classical_propositional_semantic_compactness}.
\end{proof}
