\subsection{Net convergence}\label{subsec:net_convergence}

\subsubsection{Nets}

\begin{definition}\label{def:topological_net}\mcite[65]{Kelley1975}
  A \term{net} or \term{generalized sequence} is a family of elements of a \hyperref[def:topological_space]{topological space} \hyperref[def:cartesian_product/indexed_family]{indexed} by a \hyperref[def:directed_set]{directed set}.

  We reuse the notation for indexed families from \fullref{def:cartesian_product/indexed_family}:
  \begin{equation*}
    \set{ x_k }_{k \in \mscrK}
  \end{equation*}
\end{definition}
\begin{comments}
  \item This definition does not require a topology, but, unlike sequences, this concept of a net is only discussed in the context of topological spaces.
\end{comments}

\begin{definition}\label{def:net_eventually_in}\mcite[65]{Kelley1975}
  We say that the \hyperref[def:topological_net]{net} \( \set{ x_k }_{k \in \mscrK} \) is \term{eventually in} the set \( A \) if there exists an index \( k_0 \in \mscrK \) such that, for every \( k \geq k_0 \), we have \( x_k \in A \).
\end{definition}

\begin{proposition}\label{thm:eventually_in_implies_nonempty}
  If the net \( \set{ x_k }_{k \in \mscrK} \) is \hyperref[def:net_eventually_in]{eventually in} some set \( A \), then both \( \mscrK \) and \( A \) are nonempty.
\end{proposition}
\begin{proof}
  By definition, there exists some index \( k_0 \in \mscrK \) such that \( x_{k_0} \in A \).
\end{proof}

\begin{proposition}\label{thm:sequence_eventally_in}
  An \hyperref[def:sequence]{infinite sequence} is eventually in some set if and only if only finitely many sequence elements are outside the set.
\end{proposition}
\begin{proof}
  A sequence \( \seq{ x_k }_{k=1}^\infty \) is eventually in \( A \) if there exists a positive integer \( k_0 \) such that \( x_k \in A \) whenever \( k \geq k_0 \). Since the positive integers are totally ordered, the condition is equivalent to requiring that there exists some upper limit for the set \( \set{ k \given x_k \not\in A } \) --- every such limit can be taken to be \( k_0 \). \Fullref{thm:natural_number_cofinal_subsets} implies that the latter condition is equivalent to requiring that only finitely many sequence elements are outside \( A \).
\end{proof}

\begin{definition}\label{def:net_frequently_in}\mcite[65]{Kelley1975}
  We say that the \hyperref[def:topological_net]{net} \( \set{ x_k }_{k \in \mscrK} \) is \term{frequently in} the set \( A \) if, for every index \( k_0 \in \mscrK \), there exists some \( k \geq k_0 \) such that \( x_k \in A \).
\end{definition}

\begin{proposition}\label{thm:empty_index_set_implies_frequently_in}
  A net \( \set{ x_k }_{k \in \varnothing} \) over the empty set is \hyperref[def:net_frequently_in]{frequently in} any set.
\end{proposition}
\begin{proof}
  Any condition with quantifier \enquote{for every index in \( \varnothing \)} is vacuously true.
\end{proof}

\begin{proposition}\label{thm:eventually_in_implies_frequently_in}
  If a net is \hyperref[def:net_eventually_in]{eventually in} a set, it is also \hyperref[def:net_frequently_in]{frequently in} it.
\end{proposition}
\begin{proof}
  Suppose that \( \seq{ x_k }_{k \in \mscrK} \) is eventually in \( A \). Then there exists some index \( k_e \in \mscrK \) such that \( x_k \in A \) for \( k \geq k_e \).

  Fix some arbitrary index \( k_f \in \mscrK \). Then, since \( \mscrK \) is directed, there must exist an upper bound \( k_u \) for \( k_e \) and \( k_f \).

  Since \( k_u \geq k_e \), we have \( x_{k_u} \in A \). Therefore, \( \seq{ x_k }_{k \in \mscrK} \) is frequently in \( A \).
\end{proof}

\begin{proposition}\label{thm:eventually_and_frequently_in}
  The net \( \seq{ x_k }_{k \in \mscrK} \) in the space \( X \) is \hyperref[def:net_eventually_in]{eventually in} \( A \) if and only if it is not \hyperref[def:net_frequently_in]{frequently in} \( X \setminus A \).
\end{proposition}
\begin{proof}
  \Fullref{thm:relativized_first_order_quantifiers_are_dual} implies that
  \begin{equation*}
    \qexists {k_0 \in \mscrK} \qforall {k \geq k_0} x_k \in A
  \end{equation*}
  if and only if
  \begin{equation*}
    \neg \qforall {k_0 \in \mscrK} \qexists {k \geq k_0} \neg (x_k \in A).
  \end{equation*}
\end{proof}

\begin{proposition}\label{thm:cofinal_iff_frequently_in}
  Let \( \mscrK \) be a directed set, let \( A \subseteq \mscrK \) be a subset and let \( \seq{ x_k }_{k \in \mscrK} \) be a net.

  Then \( A \) is \hyperref[def:cofinal_set]{cofinal} if and only if \( x \) is \hyperref[def:net_frequently_in]{frequently in} the \( \set{ x_a \given a \in A } \).
\end{proposition}
\begin{proof}
  The net \( \seq{ x_k }_{k \in \mscrK} \) is frequently in \( \set{ x_a \given a \in A } \) if
  \begin{equation*}
    \qforall {k_0 \in \mscrK} \qexists {a_0 \in \mscrK} k \leq a_0 \T{and} x_{a_0} \in A,
  \end{equation*}
  which can be rewritten as
  \begin{equation*}
    \qforall {k_0 \in \mscrK} \qexists {a_0 \in A} k \leq a_0.
  \end{equation*}

  This is exactly the condition for \( A \) to be cofinal in \( \mscrK \).
\end{proof}

\begin{example}\label{ex:def:topological_net}
  We list several examples of \hyperref[def:topological_net]{nets}:
  \begin{thmenum}
    \thmitem{ex:def:topological_net/empty} The most basic example of a net is the empty set. Indeed, an empty set is vacuously directed.

    It is also vacuously \hyperref[def:net_frequently_in]{frequently in} any set as a consequence of \fullref{thm:empty_index_set_implies_frequently_in}, but it is not \hyperref[def:net_eventually_in]{eventually in} any set because of \fullref{thm:eventually_in_implies_nonempty}.

    \thmitem{ex:def:topological_net/finite_sequence} Any \hyperref[def:sequence]{finite sequence} \( x_1, \ldots, x_n \) is a net.

    It is eventually in (and hence also frequently in) \( \set{ x_n } \) and any of its supersets.

    \thmitem{ex:def:topological_net/sequence} For a finite sequence it suffices for \( A \) to contain the last element, but for an infinite sequence \( x_1, x_2, x_3, \ldots \), this condition is more complicated. \Fullref{thm:cofinal_iff_frequently_in} implies that the sequence is frequently in some set \( A \) if and only if \( A \) contains a \hyperref[def:cofinal_set]{cofinal subset} of the sequence (when regarding the sequence as a well-ordered set).

    \thmitem{ex:def:topological_net/reverse} Given any point \( x_0 \), its \hyperref[def:neighborhood_system]{neighborhood filter} \( \mscrT(x_0) \) is naturally a \hyperref[def:semilattice/lattice]{lattice} with respect to inclusion. Indeed, the union and intersection of any pair of neighborhoods of \( x_0 \) is again a neighborhood of \( x_0 \).

    We can also consider the \hyperref[def:semilattice/duality]{dual lattice} of \( \mscrT(x_0) \), that is, \( V \leq U \) if \( V \supseteq U \).

    The latter is useful as an index set of nets, for example in the proofs of \fullref{thm:def:net_limit_point/cluster_point_subnet_limit} and \fullref{thm:limit_point_iff_in_closure} or the proof of equivalence in \fullref{def:local_convergence}.
  \end{thmenum}
\end{example}

\subsubsection{Subnets}

It is unfortunately an established convention for subsequences to be defined differently compared to subnets. There are actually different definitions of subnets --- see \cite{Lehuta2009} --- but we will base our exposition on that of John Kelley in \cite[ch. 2]{Kelley1975}.

\begin{definition}\label{def:subsequence}\mcite[63]{Kelley1975}
  We say that
  \begin{equation}\label{eq:def:subsequence/sub}
    y_1, y_2, y_3, y_4, \ldots
  \end{equation}
  is a \term{subsequence} of
  \begin{equation}\label{eq:def:subsequence/original}
    x_1, x_2, x_3, x_4, \ldots
  \end{equation}
  if there exists a strictly increasing sequence of indices
  \begin{equation*}
    i_1 < i_2 < i_3 < i_4 < \cdots
  \end{equation*}
  such that \( y_k = x_{i_k} \) for every \( k = 1, 2, \ldots \).
\end{definition}
\begin{comments}
  \item While Kelley in \cite[63]{Kelley1975} calls this definition \enquote{unnecessarily stringent} before introducing nets, other authors like Walter Rudin in \cite[def. 3.5]{Rudin1976Principles} prefer this definition.

  \item If we regard \( x \) and \( y \) as functions from \( \BbbZ_{>0} \) to \( X \), then \( y \) is a subsequence of \( x \) if and only if there exists a \hyperref[def:order_homomorphism/increasing]{strictly increasing} inclusion \hyperref[def:function/endofunction]{endofunction} \( \iota: \BbbZ_{>0} \to \BbbZ_{>0} \) such that \( y = x \cdot \iota \).
\end{comments}

\begin{definition}\label{def:subnet}\mcite[70]{Kelley1975}
  We say that the net \( \seq{ y_m }_{m \in \mscrM} \) is a \term{subnet} of \( \seq{ x_k }_{k \in \mscrK} \) if there exists a family \( \seq{ k_m }_{m \in \mscrM} \) of indices from \( \mscrK \) such that
  \begin{thmenum}
    \thmitem[def:subnet/composition]{SN1} For every \( m \in \mscrM \), we have \( x_{k_m} = y_m \).

    This condition ensures that the subnet consists of members of the supernet.

    \thmitem[def:subnet/cofinality]{SN2} For every \( k_0 \in \mscrK \), there exists some index \( m_0 \in \mscrM \) such that, for every \( m \geq m_0 \), we have \( k_m \geq k_0 \).

    This condition ensures that the subnet \enquote{follows} the direction of the supernet. It involves only the index sets and not the actual nets.
  \end{thmenum}
\end{definition}
\begin{comments}
  \item \ref{def:subnet/cofinality} states that the subnet must be eventually in the tail \( \set{ x_k \given k \geq k_0 } \) for every \( k_0 \in \mscrK \).
\end{comments}

\begin{example}\label{ex:def:subnet}
  We list several examples of \hyperref[def:subnet]{subnets}:
  \begin{thmenum}
    \thmitem{ex:def:subnet/empty} The empty net has no empty subnet.

    \thmitem{ex:def:subnet/subsequence} Consider the \hyperref[def:sequence]{sequence}
    \begin{equation*}
      x_1, x_2, x_3, \ldots
    \end{equation*}

    The sequence
    \begin{equation*}
      x_1, x_3, x_5, \ldots
    \end{equation*}
    with only odd indices is a \hyperref[def:subsequence]{subsequence}, and so is
    \begin{equation*}
      x_2, x_3, x_4, \ldots
    \end{equation*}

    The above are also subnets.

    \thmitem{ex:def:subnet/padded_sequence} The following is a subnet but not subsequence:
    \begin{equation*}
      \underbrace{0}_{y_1}, \underbrace{x_1}_{y_2}, \underbrace{x_2}_{y_3}, \underbrace{x_3}_{y_4}, \underbrace{x_4}_{y_5}, \cdots.
    \end{equation*}

    Indeed, consider the family of indices
    \begin{equation*}
      k_m = \begin{cases}
        0,     &m = 1, \\
        m - 1, &m > 1.
      \end{cases}
    \end{equation*}

    The family is not monotone, yet it defines a subnet because it satisfies \fullref{def:subnet}:
    \begin{itemize}
      \item Since \( y_m = x_{k_m} \) for every \( m = 1, 2, \ldots \), \ref{def:subnet/composition} holds.
      \item For every \( k_0 \), we have an index \( m_0 = k_0 + 1 \) such that \( k_m = m - 1 \geq m_0 - 1 = k_0 \) for \( m \geq m_0 \); hence \ref{def:subnet/cofinality} also holds.
    \end{itemize}

    \thmitem{ex:def:subnet/real_subnet} Consider again some sequence
    \begin{equation*}
      x_1, x_2, x_3, \ldots
    \end{equation*}

    A subnet needs not be a sequence. For example, consider the \hyperref[def:totally_ordered_set]{totally ordered set} \( \BbbR_{>0} \) of positive \hyperref[def:real_numbers]{real numbers}. For every positive real number \( r \), let \( k_r \coloneqq \ceil(r) \) be the smallest integer not smaller than \( r \), and define the net \( y_r \coloneqq x_{k_r} \).

    Then \ref{def:subnet/composition} is automatically satisfied.

    Fix some positive integer \( k_0 \). Then we have an index \( r_0 \coloneqq k_0 \) such that, for \( r \geq r_0 \),
    \begin{equation*}
      k_r = \ceil(r) \geq \ceil(r_0) = r_0 = k_0.
    \end{equation*}

    Therefore, \ref{def:subnet/cofinality} holds and \( \seq{ y_r }_{r \in \BbbR} \) is a subnet.

    \thmitem{ex:def:subnet/divisibility_subnet} Furthermore, a subnet of a sequence needs not even be totally ordered. Given some sequence, if we order its index set by divisibility\footnote{The natural number divisibility lattice is described in \fullref{thm:natural_number_divisibility_lattice}.} rather than by the conventional order, we will obtain a subnet of the original.

    Indeed, for every positive integer \( k_0 \), we have an index \( m_0 \coloneqq k_0 \) such that, for \( m_0 \mid m \), we have \( m \geq k_0 \).
  \end{thmenum}
\end{example}

\subsubsection{Cluster points}

We will now work in an ambient topological space \( X \).

\begin{definition}\label{def:net_cluster_point}\mcite[71]{Kelley1975}
  We say that the point \( x_0 \) is an \term{cluster point} or \term{accumulation point} of the net \( \seq{ x_k }_{k \in \mscrK} \) if it is \hyperref[def:net_frequently_in]{frequently in} every neighborhood of \( x_0 \).
\end{definition}
\begin{comments}
  \item \Fullref{thm:net_cluster_point_base} shows that it is sufficient to only consider bases.
\end{comments}

\begin{example}\label{ex:def:net_cluster_point}
  We give several examples of \hyperref[def:net_cluster_point]{net cluster points}:

  \begin{thmenum}
    \thmitem{ex:def:net_cluster_point/empty} \Fullref{thm:empty_index_set_implies_frequently_in} implies that the empty net \( \seq{ x_k }_{k \in \varnothing} \) is frequently in every set, including every neighborhood of every point.

    Therefore, every point is a cluster point of the empty net.

    \thmitem{ex:def:net_cluster_point/finite_sequence} As discussed in \fullref{ex:def:topological_net/finite_sequence}, any finite sequence \( x_1, x_2, \ldots, x_n \) is frequently in \( \set{ x_n } \) and its supersets, hence also in every neighborhood of \( x_n \).

    The last element of a finite sequence is thus a cluster point. Furthermore, it is the only cluster point of the sequence.

    \thmitem{ex:def:net_cluster_point/sequence} An easy example of a sequence with two cluster points is
    \begin{equation*}
      0, 1, 0, 1, 0, 1, 0, 1, \cdots
    \end{equation*}

    It is frequently in \( \set{ 0 } \) and in \( \set{ 1 } \), hence in all of their neighborhoods (independent of the topology). Thus, both \( 0 \) and \( 1 \) are cluster points.

    \thmitem{ex:def:net_cluster_point/sierpinski} Consider the \hyperref[def:sierpinski_space]{Sierpinski space} \( \set{ \top, \bot } \). The constant sequence
    \begin{equation*}
      \top, \top, \top, \top, \top, \top
    \end{equation*}
    obviously has \( \top \) as a cluster point since it is frequently in the neighborhoods \( \set{ \top } \) and \( \set{ \top, \bot } \) of \( \top \).

    Furthermore, since \( \set{ \top, \bot } \) is the only neighborhood of \( \bot \), the latter is also a cluster point of the sequence (despite not belonging to it).

    \thmitem{ex:def:net_cluster_point/natural_divisibility} Consider the sequence
    \begin{equation*}
      p_1, p_2, p_3, \cdots,
    \end{equation*}
    where \( p_k \) is the smallest \hyperref[def:prime_number]{prime number} that divides \( k \). For completeness, let \( p_1 \coloneqq 1 \).

    The first terms are
    \begin{equation*}
      1, 2, 3, 2, 5, 2, 7, 2, 3, 2, 11, 2, 13, 2, 3, \ldots
    \end{equation*}

    The sequence is frequently in \( \set{ 2 } \) since, for any number \( n \), \( 2n \) is a strictly larger even number. Analogously, it is frequently in \( \set{ p } \) for every prime \( p \) because, for any number \( n \), \( pn \) is a strictly larger number divisible by \( p \).

    Therefore, under any topology, the prime numbers are cluster points, and no other numbers except \( 0 \) and \( 1 \) occur.

    \thmitem{ex:def:net_cluster_point/sin} Consider the net \( \seq{ \sin(r) }_{r \in \BbbR} \).

    Every real number in the interval \( [-1, 1] \) is a cluster point under any topology. Indeed, fix \( x_0 \in [-1, 1] \).

    For any index \( r_0 \in \BbbR \), let \( n_0 \) be the smallest integer such that \( r_0 \leq 2 n_0 \pi \). Then
    \begin{equation*}
      \sin(r_0) = \sin(\underbrace{2 n_0 \pi + \arcsin(x_0)}_r).
    \end{equation*}

    Thus, \( r \geq r_0 \) and \( \sin(r) = x_0 \), meaning that the net \( \seq{ \sin(r) }_{r \in \BbbR} \) is frequently in \( \set{ x_0 } \) (and hence any neighborhood of \( x_0 \)).

    Therefore, every number in the interval \( [-1, 1] \) is a cluster point for the net.
  \end{thmenum}
\end{example}

\begin{proposition}\label{thm:net_cluster_point_base}
  Let \( \mscrB \) be a \hyperref[def:topological_base]{base} in some topological space.

  The point \( x_0 \) is a \hyperref[def:net_cluster_point]{cluster point} of the net \( \seq{ x_k }_{k \in \mscrK} \) if and only if it is \hyperref[def:net_frequently_in]{frequently in} every neighborhood of \( x_0 \) in \( \mscrB \).
\end{proposition}
\begin{proof}
  \SufficiencySubProof Every set in \( \mscrB \) is itself a neighborhood of \( x_0 \), hence if \( x_0 \) is a cluster point, the net is frequently in every set from \( \mscrB \) containing \( x_0 \).

  \NecessitySubProof Suppose that the net is frequently in every set in \( \mscrB \) containing \( x_0 \).

  Let \( U \) be a neighborhood of \( x_0 \). By definition of base, there exists some set \( V \) in \( \mscrB \) such that
  \begin{equation*}
    x_0 \in V \subseteq U.
  \end{equation*}

  For every index \( k_0 \in \mscrK \) then there exists some \( k \geq k_0 \) such that \( x_k \in V \subseteq U \).

  Since \( U \) was arbitrary, we conclude that \( x_0 \) is a cluster point of the net.
\end{proof}

\subsubsection{Limit points}

Assume again that we work in an ambient topological space \( X \).

\begin{definition}\label{def:net_limit_point}\mcite[66]{Kelley1975}
  We say that the point \( x_0 \) is a \term{limit point} of the net \( \seq{ x_k }_{k \in \mscrK} \) if it is \hyperref[def:net_eventually_in]{eventually in} every neighborhood of \( x_0 \).

  We say that a net is \term{convergent} if it has a limit point and \term{divergent} otherwise.
\end{definition}
\begin{comments}
  \item \Fullref{thm:net_limit_point_subbase} shows that it is sufficient to only consider subbases.

  \item In general, there can exist multiple limit points --- see \fullref{ex:def:net_limit_point}. In Hausdorff spaces, however, limits are unique as shown in \fullref{thm:t2_iff_singleton_limits}.
\end{comments}

\begin{example}\label{ex:def:net_limit_point}
  We give several examples of \hyperref[def:net_limit_point]{net limit points}:

  \begin{thmenum}
    \thmitem{ex:def:net_limit_point/empty} The empty net cannot have limit points as a consequence of \fullref{thm:eventually_in_implies_nonempty}.

    \thmitem{ex:def:net_limit_point/finite_sequence} As in the case of cluster points in \fullref{ex:def:net_cluster_point/finite_sequence}, a finite sequence \( x_1, \ldots, x_n \) can have only one limit point --- \( x_n \).

    \thmitem{ex:def:net_limit_point/sin} We saw in \fullref{ex:def:net_cluster_point/sin} that the net \( \seq{ \sin(r) }_{r \in \BbbR} \) has a lot of cluster points. It does not have a limit point, however.

    Indeed, suppose that \( x_0 \) is a limit point. Fix some neighborhood \( U \) of \( x_0 \) in the topological space \( [-1, 1] \) that is not the whole space.

    There exists some \( r_0 \) such that \( \sin(r) \in U \) for \( r \geq r_0 \). But
    \begin{equation*}
      \sin([r_0, r_0 + 2\pi)) = [-1, 1]
    \end{equation*}
    and we have explicitly assumed that \( U \) is a strict subset of \( [-1, 1] \).

    Therefore, no point in \( [-1, 1] \) is a limit point of the net \( \seq{ \sin(r) }_{r \in \BbbR} \).

    \thmitem{ex:def:net_limit_point/sierpinski_constant} As in \fullref{ex:def:net_cluster_point/sierpinski}, consider the \hyperref[def:sierpinski_space]{Sierpinski space} \( \set{ \top, \bot } \) and the constant sequence
    \begin{equation*}
      \top, \top, \top, \top, \top, \top.
    \end{equation*}

    Both \( \top \) and \( \bot \) are cluster points, but they are also limit points. They are distinct and even \hyperref[def:topologically_indistinguishable]{topologically distinguishable}.

    \thmitem{ex:def:net_limit_point/sierpinski_alternating} Consider instead the Sierpinski space with the alternating sequence
    \begin{equation*}
      \top, \bot, \top, \bot, \top, \bot.
    \end{equation*}

    Both \( \top \) and \( \bot \) are cluster points. Furthermore, \( \bot \) is a limit point, while \( \top \) is not.

    \thmitem{ex:def:net_limit_point/multiple} Limits of sequences need not be unique in general topological spaces. Let \( X = \set{ a, b } \) be a binary set with the \hyperref[def:indiscrete_topology]{indiscrete topology} \( \set{ \varnothing, X } \).

    Define the following \hyperref[def:sequence]{sequence}
    \begin{equation*}
      x_k \coloneqq \begin{cases}
        a, &k \T{is even}, \\
        b, &k \T{is odd}.
      \end{cases}
    \end{equation*}

    The only neighborhood of \( y \), the whole space \( X \), contains all members of the sequence, therefore \( y \) is a limit point of the sequence. The same is true for \( z \), however.
  \end{thmenum}
\end{example}

\begin{proposition}\label{thm:def:net_limit_point}
  \hyperref[def:net_limit_point]{Net limit points} have the following basic properties:

  \begin{thmenum}
    \thmitem{thm:def:net_limit_point/limit_point_is_cluster_point} Every limit point is a cluster point.

    The converse is true for converging nets in Hausdorff spaces, but not in general --- see \fullref{ex:def:net_limit_point/sierpinski_alternating}.

    \thmitem{thm:def:net_limit_point/cluster_point_in_subnet} Every cluster point of a subnet is also a cluster point of the entire net.

    \thmitem{thm:def:net_limit_point/limit_and_subnet_limit} A net has a limit point if and only if it is a limit point of all of its subnets.

    \thmitem{thm:def:net_limit_point/cluster_point_subnet_limit} A point is a cluster point of a net if and only if it is a limit point of some \hyperref[def:subnet]{subnet}.

    \thmitem{thm:def:net_limit_point/cluster_and_limit_point} Given a limit point and a cluster point of the same net, every neighborhood of the first intersects every neighborhood of the second and vice versa.

    In a general topological space, the two points may still be \hyperref[def:topologically_indistinguishable]{topologically distinguishable} as shown in \fullref{ex:def:net_limit_point/sierpinski_constant}.

    \thmitem{thm:def:net_limit_point/sequence_converges_iff_almost_entirely_in_neighborhood} Follows from \fullref{thm:sequence_eventally_in}.
  \end{thmenum}
\end{proposition}
\begin{proof}
  \SubProofOf{thm:def:net_limit_point/limit_point_is_cluster_point} Follows from \fullref{thm:eventually_in_implies_frequently_in}.

  \SubProofOf{thm:def:net_limit_point/cluster_point_in_subnet} Let \( \seq{ y_m }_{k \in \mscrK} \) be a subnet of \( \seq{ x_k }_{k \in \mscrK} \) with index translation \( \seq{ k_m }_{m \in \mscrM} \). Suppose that \( y_0 \) is a cluster point of \( \seq{ y_m }_{k \in \mscrK} \).

  Let \( U \) be a neighborhood of \( y_0 \) and fix some index \( k_0 \in \mscrK \). \ref{def:subnet/cofinality} implies that there exists some index \( m_0 \) such that \( k_m \geq k_0 \) for \( m \geq m_0 \).

  Since \( y_0 \) is a cluster point of \( \seq{ k_m }_{m \in \mscrM} \), there exists some index \( m' \geq m_0 \) such that \( y_{m'} = x_{k_m'} \in U \). Since we have \( k_m' \geq k_0 \), we conclude that the entire net \( \seq{ x_k }_{k \in \mscrK} \) is frequently in \( U \).

  The neighborhood \( U \) was chosen arbitrarily, hence we conclude that \( y_0 \) is a cluster point of \( \seq{ x_k }_{k \in \mscrK} \).

  \SubProofOf{thm:def:net_limit_point/limit_and_subnet_limit}

  \SufficiencySubProof* Let \( x_0 \) be a limit point of \( \seq{ x_k }_{k \in \mscrK} \) and let \( \seq{ y_m }_{m \in \mscrM} \) be a subnet with index translation \( \seq{ k_m }_{m \in \mscrM} \).

  Let \( U \) be a neighborhood of \( x_0 \). Then there exists an index \( k_0 \) such that \( x_k \in U \) for \( k \geq k_0 \). Since \( \seq{ y_m }_{m \in \mscrM} \) is a subnet, \ref{def:subnet/cofinality} states that there exists some index \( m_0 \) such that \( k_m \geq k_0 \) for \( m \geq m_0 \).

  We showed that the subnet is frequently in an arbitrary neighborhood of \( x_0 \), implying that \( x_0 \) is a limit point of the subnet.

  \NecessitySubProof* Suppose that \( x_0 \) is a limit point of all subnets of \( \seq{ x_k }_{k \in \mscrK} \). Every net is a subnet of itself, hence \( x_0 \) is a limit point of \( \seq{ x_k }_{k \in \mscrK} \).

  \SubProofOf{thm:def:net_limit_point/cluster_point_subnet_limit}

  \SufficiencySubProof* Let \( x_0 \) be a cluster point of the net \( \seq{ x_k }_{k \in \mscrK} \).

  For every neighborhood \( U \) of \( x_0 \), there exists some index \( k_U \) such that \( x_{k_U} \in U \).

  Consider the \hyperref[ex:def:topological_net/reverse]{dual filter} \( (\mscrT(x_0), \supseteq) \). For every concrete neighborhood \( U_0 \), there exists some index \( k_{U_0} \) such that \( x_{k_V} \in U_0 \) whenever the neighborhood \( V \) is a subset of \( U_0 \), that is, whenever \( V \geq U_0 \).

  Therefore, \( \seq{ k_U }_{U \in \mscrT(x_0)} \) is a subnet of \( \seq{ x_k }_{k \in \mscrK} \).

  Furthermore, it converges to \( x_0 \) because it is eventually in each of its neighborhoods.

  \NecessitySubProof* Follows from \fullref{thm:def:net_limit_point/limit_point_is_cluster_point} and \fullref{thm:def:net_limit_point/cluster_point_in_subnet}.

  \SubProofOf{thm:def:net_limit_point/cluster_and_limit_point} Let \( l \) be a limit point of \( \seq{ x_k }_{k \in \mscrK} \) and let \( c \) be a cluster point.

  Let \( U \) be a neighborhood of \( l \). Then there exists an index \( k_l \) such that \( x_k \in U \) for \( k > k_l \).

  Let \( V \) be a neighborhood of \( c \). Then there exists an index \( k_c \geq k_l \) such that \( x_{k_c} \in V \). Furthermore, \( x_{k_c} \in U \cap V \).

  Therefore, \( U \) and \( V \) intersect. Since they are arbitrary neighborhoods, we conclude that the proposition holds.

  \SubProofOf{thm:def:net_limit_point/sequence_converges_iff_almost_entirely_in_neighborhood} This is simply a restatement of \fullref{def:net_limit_point} for the special case of sequences.
\end{proof}

\begin{proposition}\label{thm:net_limit_point_subbase}
  Let \( \mscrS \) be a \hyperref[def:topological_base]{base} in some topological space.

  The point \( x_0 \) is a \hyperref[def:net_limit_point]{limit point} of the net \( \seq{ x_k }_{k \in \mscrK} \) if and only if it is \hyperref[def:net_eventually_in]{eventually in} every neighborhood of \( x_0 \) in \( \mscrB \).
\end{proposition}
\begin{proof}
  \SufficiencySubProof Every set in \( \mscrS \) is itself a neighborhood of \( x_0 \), hence if \( x_0 \) is a limit point, the net is eventually in every set from \( \mscrS \) containing \( x_0 \).

  \NecessitySubProof Suppose that the net is eventually in every set in \( \mscrS \) containing \( x_0 \).

  Let \( U \) be a neighborhood of \( x_0 \). By definition of subbase, there exists some family \( V_1, \ldots, V_n \) in \( \mscrS \) such that
  \begin{equation*}
    x_0 \in V_1 \cap \cdots \cap V_n \subseteq U.
  \end{equation*}

  For every \( i = 1, \ldots, n \), there exists some \( k_i \in \mscrK \) such that \( k \geq k_i \) such that \( x_k \in V_i \subseteq U \).

  Since \( \mscrK \) is a directed set, there exists some upper bound \( k_0 \) for \( k_1, \ldots, k_n \). Then, if \( k \geq k_0 \), we have
  \begin{equation*}
    x_0 \in V_1 \cap \cdots \cap V_n \subseteq U.
  \end{equation*}

  Since \( U \) was arbitrary, we conclude that \( x_0 \) is a limit point of the net.
\end{proof}

\begin{proposition}\label{thm:limit_point_iff_in_closure}\mcite[prop. 1.6.3]{Engelking1989}
  Fix a set \( A \subseteq X \). A point \( x_0 \in X \) belongs to \( \cl{A} \) if and only if there exists a net \( \seq{ x_k }_{k \in \mscrK} \subseteq A \) for which \( x_0 \) is a limit point.

  By \fullref{thm:def:net_limit_point/cluster_point_subnet_limit}, we can consider cluster points of nets rather than limit points.
\end{proposition}
\begin{proof}
  The complement of the empty set is the empty set, hence the statement of the proposition holds vacuously. Assume that \( A \) is nonempty.

  \SufficiencySubProof Suppose that \( x_0 \in \cl{A} \). If \( x_0 \in A \), then the one-element net \( (x_0) \) converges to \( x_0 \).

  If \( x_0 \in \fr{A} \), by \fullref{thm:properties_via_bases/boundary}, every neighborhood of \( x_0 \) contains points from \( A \). Therefore, we can build reverse inclusion net in the style of \fullref{ex:def:topological_net/reverse} that converges to \( x_0 \).

  \NecessitySubProof Let \( x_0 \) be a limit point of \( \seq{ x_k }_{k \in \mscrK} \subseteq A \). We will show that \( x_0 \) belongs every closed set that contains \( A \).

  Let \( F \supseteq A \) be a closed set. Denote \( U \coloneqq X \setminus F \). Suppose that \( x_0 \in U \). Then \( U \) is a neighborhood \( x_0 \) and, by \fullref{def:net_cluster_point}, the net \( \seq{ x_k }_{k \in \mscrK} \subseteq A \) is eventually in \( U \). But \( U \) does not contains \( A \).

  The obtained contradiction shows that \( x_0 \) belongs to every closed set containing \( A \) and hence to their intersection, the closure \( \cl A \).
\end{proof}

\begin{proposition}\label{thm:cluster_point_of_set_iff_limit_point_of_net}
  The point \( x_0 \in X \) is a cluster \hyperref[def:set_cluster_point]{point} of the set \( A \) if and only if it is a limit \hyperref[def:net_cluster_point]{point} of some net in \( A \setminus \{ x_0 \} \) (or, equivalently, a cluster point of some net in \( A \setminus \{ x_0 \} \)).
\end{proposition}
\begin{proof}
  \SufficiencySubProof Let \( x_0 \in \drv(A) \). By \fullref{thm:properties_via_bases/cluster}, every neighborhood \( U \) of \( x_0 \) intersects \( A \setminus \{ x_0 \} \). Choose \( x_U \in U \cap (A \setminus \{ x_0 \}) \) for every neighborhood \( U \) of \( x_0 \) and form the reverse inclusion \hyperref[ex:def:topological_net/reverse]{net} \( \{ x_U \}_{U \in T(x)} \). Then \( x_0 \) is a limit point of this net. Furthermore, the net is contained in \( A \setminus \{ x_0 \} \).

  \NecessitySubProof Conversely, if \( \seq{ x_k }_{k \in \mscrK} \subseteq A \setminus \{ x_0 \} \) is a net and if \( x_0 \) is a limit point of this net, then for every neighborhood \( U \) of \( x_0 \) there exists an index \( i_U \) such that for \( k \geq i_U \) we have \(  x_k \in U \). In particular, \( U \cap A \) contains elements other than \( x_0 \). Since this is true for any neighborhood \( U \) of \( x_0 \), by \fullref{thm:properties_via_bases/cluster} we conclude that \( x_0 \) is a cluster point of the set\( A \).
\end{proof}

\begin{corollary}\label{thm:closed_iff_contains_all_net_cluster_points}
  A set is closed if and only if it contains the limit points of all of its nets (or, equivalently, the cluster points of all of its nets).
\end{corollary}
\begin{proof}
  By \fullref{thm:cluster_point_characterization}, the set \( A \) is closed if and only if it contains all of its cluster points. By \fullref{thm:cluster_point_of_set_iff_limit_point_of_net}, this is equivalent to \( A \) containing all limit points of its nets.
\end{proof}

\begin{proposition}\label{thm:net_convergence_via_subbases}
  Fix a topological space \( X \), a point \( x_0 \) and a local \hyperref[def:topological_local_subbase]{subbase} \( P(x_0) \). The point \( x_0 \) is a limit of the net \( \seq{ x_k }_{k \in \mscrK} \subseteq X \) if and only if it is eventually in every element \( U_P \) of the local subbase \( P(x_0) \).
\end{proposition}
\begin{proof}
  \SufficiencySubProof Obvious consequence of the definition of local subbase.
  \NecessitySubProof Fix a neighborhood \( U \) of \( x_0 \). By \fullref{def:topological_local_subbase}, there exists a finite family \( \{ U_k \}_{k=1}^n \subseteq P(x_0) \) such that \( \bigcap_{k=1}^n U_k \subseteq U \). Since the net \( \seq{ x_k }_{k \in \mscrK} \subseteq X \) is eventually in each of \( U_k, k = 1, \ldots, n \), from transitivity of inclusion it follows that the net is eventually in \( U \).
\end{proof}

\begin{definition}\label{def:sequential_topological_closure_operator}
  In analogy to \fullref{def:topological_closure_operator}, we define the \term{sequential closure operator}
  \begin{balign*}
     & \cl^S: \pow(X) \to \pow(X)                                                                                                          \\
     & \cl^S(A) \coloneqq \left\{ x \in X \colon x \text{ is a limit point of some sequence } \seq{ x_k }_{k=1}^\infty \subseteq A \right\}.
  \end{balign*}

  If \( \cl^S(A) = A \), we say that \( A \) is \term{sequentially closed}.
\end{definition}

\begin{definition}\label{def:sequential_space}
  A topological space is called \term{sequential} if every sequentially \hyperref[def:sequential_topological_closure_operator]{closed} set is closed.
\end{definition}

\begin{remark}\label{rem:sequential_spaces}
  By \fullref{thm:limit_point_iff_in_closure}, in a \hyperref[def:sequential_space]{sequential space}, a set is closed if and only if it is sequentially closed.

  By \fullref{thm:closed_iff_contains_all_net_cluster_points}, a set is closed if and only if it contains the limit points of all of its nets.

  Therefore, a set in a sequential space is closed if and only if it contains the limit points of all of its sequences.

  Since we are able to define a topology in terms of closed sets, this means that the topology in a sequential space is completely determined by convergent sequences rather than convergent nets as in general topological spaces.

  This allows us to restrict ourselves only to sequences rather than arbitrary nets in certain spaces like \hyperref[def:metric_space]{metric spaces}.
\end{remark}

\begin{lemma}\label{thm:sequential_space_convergence}
  Let \( X \) be a sequential space and \( x_0 \) limit point of the net \( \seq{ x_k }_{k \in \mscrK} \), then we can define a sequence
  \begin{equation*}
    \seq{ x_k }_{k=1}^\infty \subseteq \{ x_k \colon k \in \mscrK \},
  \end{equation*}
  consisting of members of the net, for which \( x_0 \) is a limit point.
\end{lemma}
\begin{proof}
  Let \( X \) be a first-countable space and let \( x_0 \) be a limit point of the net \( \seq{ x_k }_{k \in \mscrK} \).

  Since \( X \) is a first countable space, we can fix a countable local \hyperref[def:topological_local_base]{base} \( \{ U_k \}_{k=1}^\infty \) at \( x_0 \). For each \( k = 1, 2, \ldots \), define the neighborhood \( V_k \coloneqq \bigcap_{m=1}^k U_m \), so that \( V_k \subseteq V_m \) whenever \( k \geq m \).

  For each neighborhood \( V_k \), since \( \seq{ x_k }_{k \in \mscrK} \) is eventually in \( V_k \), there exists an index \( i_k \) such that \( x_{i_k} \in V_k \).

  Thus, we obtain a sequence \( \{ x_{i_k} \}_{k=1}^\infty \) that is eventually in every neighborhood of the local base \( \{ V_k \}_{k=1}^\infty \) of \( x_0 \), which by \fullref{thm:net_convergence_via_subbases} is sufficient for \( x_0 \) to be a limit point of the sequence.
\end{proof}

\begin{proposition}\label{thm:first_countable_spaces_are_sequential}
  Every first-countable space is sequential.
\end{proposition}
\begin{proof}
  Let \( X \) be a first-countable space and let \( A \subseteq X \) be a sequentially \hyperref[def:sequential_topological_closure_operator]{closed} set. We must show that it is closed.

  Fix a point \( x_0 \in \cl(A) \). We will show that \( x_0 \in A \). By \fullref{thm:limit_point_iff_in_closure}, there is a net \( \seq{ x_k }_{k \in \mscrK} \subseteq A \) for which \( x_0 \) is a limit point.

  By \fullref{thm:sequential_space_convergence}, we can choose a sequence \( \seq{ x_k }_{k=1}^\infty \) that converges to \( x_0 \) out of elements of the net. But since \( X \) is a sequential space, the limit points of any sequence are contained in the sequentially closed set \( A \).

  We showed that \( A = \cl(A) \). Since \( A \) was an arbitrary sequentially closed set, we conclude that the space \( X \) is sequential.
\end{proof}
