\subsection{Totally ordered sets}\label{subsec:totally_ordered_sets}

\begin{definition}\label{def:totally_ordered_set}\cite[def. 2.1.1(ii)]{Hinman2005}
  We say that a partially ordered set is \term{totally ordered} if either the nonstrict order \( \leq \) is \hyperref[def:binary_relation/connected]{connected} or if the strict order \( < \) is \hyperref[def:binary_relation/trichotomic]{trichotomic}.

  The theory, homomorphisms and category \( \cat{Tos} \) are inherited from \fullref{def:partially_ordered_set} with one additional axiom added.
\end{definition}
\begin{comments}
  \item Peter Hinman in \incite[def. 2.1.1(ii)]{Hinman2005} uses our terminology, however some authors give slightly different definitions. \incite[62]{Enderton1977Sets} and \incite[4]{Engelking1989} both define \enquote{linear orders} as what we call \enquote{strict total orders} (Engelking defines partial orders as nonstrict). \incite[14]{Kelley1975} also does this, but mentions the possibility of nonstrict total orders. \incite[def. I.6]{Birkhoff1948} uses the term \enquote{chain} for nonstrict total orders (we also use this in the context of partial orders --- see \fullref{def:partial_order_chain}).
\end{comments}
\begin{defproof}
  Equivalence between nonstrict and strict total orders follows directly from the compatibility condition \eqref{eq:def:preordered_set/compatibility_nonstrict}.
\end{defproof}

\begin{proposition}\label{thm:total_order_embedding_iff_strict}
  An \hyperref[def:order_homomorphism/increasing]{order homomorphism} between \hyperref[def:totally_ordered_set]{totally ordered sets} is an \hyperref[def:order_homomorphism/embedding]{order embedding} if and only if it is \hyperref[def:order_homomorphism/increasing]{strict}.
\end{proposition}
\begin{proof}
  \SufficiencySubProof Follows from \fullref{thm:order_embedding_is_strict}.

  \NecessitySubProof
  \SubProof*{Proof that \( f \) is injective} Let \( f: P \to Q \) be a strict order homomorphism and suppose that \( f(x_1) = f(x_2) \) for some \( x_1 \) and \( x_2 \) in \( P \). We will use the \hyperref[def:binary_relation/trichotomic]{trichotomy} of \( < \).
  \begin{itemize}
    \item If \( x_1 < x_2 \), then \( f(x_1) < f(x_2) \) since \( f \) is strictly monotone, which contradicts our assumption \( f(x_1) = f(x_2) \).

    \item If \( x_1 > x_2 \), similarly \( f(x_1) > f(x_2) \) and we again obtain a contradiction.

    \item It remains for \( x_1 \) to be equal to \( x_2 \).
  \end{itemize}

  Since \( x_1 \) and \( x_2 \) were arbitrary, we conclude that \( f \) is injective.

  \SubProof*{Proof that \( f^{-1} \) is order-preserving} Since \( f \) is injective, its inverse \( f^{-1} \) is a single-valued partial function, and is total and injective on \( f(P) \). It then follows from \fullref{thm:order_embedding_is_strict} that the restriction of \( f^{-1} \) to \( f(P) \) is order-preserving.
\end{proof}

\begin{corollary}\label{thm:totally_ordered_strict_isomorphisms}
  An \hyperref[def:order_homomorphism/increasing]{order homomorphism} between \hyperref[def:totally_ordered_set]{totally ordered sets} is an \hyperref[def:order_homomorphism/embedding]{order isomorphism} if and only if it is \hyperref[def:order_homomorphism/increasing]{strict} and \hyperref[def:function_invertibility/surjective]{surjective}.
\end{corollary}
\begin{proof}
  Follows from \fullref{thm:total_order_embedding_iff_strict}.
\end{proof}

\begin{proposition}\label{thm:totally_ordered_minimal_element_is_minimum}
  In a \hyperref[def:totally_ordered_set]{totally ordered set}, every \hyperref[def:extremal_points/maximal_and_minimal_element]{minimal element} (resp. maximal element) is a \hyperref[def:extremal_points/maximum_and_minimum]{minimum} (resp. maximum).
\end{proposition}
\begin{proof}
  We will only prove the case for minimal elements.

  Let \( x_0 \) be a minimal element of \( (P, \leq) \). By definition of total order, for any \( x \in P \) either \( x \leq x_0 \) or \( x_0 \leq x \). If \( x \leq x_0 \), then since \( x_0 \) is a minimal element, we have \( x = x_0 \).

  Therefore, for any \( x \in P \), either \( x = x_0 \) or \( x > x_0 \). That is, \( x \leq x_0 \) always holds, proving that \( x_0 \) is a minimum.
\end{proof}

\begin{proposition}\label{thm:totally_ordered_segment_isomorphism}
  Let \( (P, \leq) \) be a \hyperref[def:totally_ordered_set]{totally ordered set}. Let \( Q \) be the set containing the \hyperref[def:order_interval/ray]{strict initial segment} \( P_{<x} \) for every member \( x \) of \( P \).

  Then \( (P, \leq) \) is \hyperref[def:order_homomorphism/isomorphism]{order-isomorphic} to \( (Q, \subseteq) \).
\end{proposition}
\begin{proof}
  Explicitly define the isomorphism
  \begin{equation*}
    \begin{aligned}
      &f: P \to Q \\
      &f(x) \coloneqq P_{<x} = \set{ y \in P \given y < x }.
    \end{aligned}
  \end{equation*}

  Note that \( f \) is \hyperref[def:order_homomorphism/increasing]{strictly order-preserving}. Indeed, if \( x < y \), then \( x \in P_{<y} \), but \( x \not\in P_{<x} \) and hence \( P_{<x} \) is a strict subset of \( P_{<y} \).

  Then \fullref{thm:total_order_embedding_iff_strict} implies that \( f \) is an embedding. Since \( f \) is also surjective by definition of \( Q \), it follows that it is a strict isomorphism between \( (P, \leq) \) and \( (Q, \subseteq) \).
\end{proof}

\begin{definition}\label{def:cofinal_set}\mcite[8]{Engelking1989}
  We say that a subset \( A \) of a preordered set \( (P, \leq) \) is \term{cofinal} if, for every \( x \in P \), there exists some \( y \in A \) such that \( x \leq y \).
\end{definition}

\begin{example}\label{ex:def:cofinal_set}
  We list several examples of \hyperref[def:cofinal_set]{cofinal} and non-cofinal sets.

  \begin{itemize}
    \thmitem{ex:def:cofinal_set/finite} In a finite set like \( \set{ 0, 1, 2 } \), the set \( \set{ 2 } \) containing the maximum is cofinal. This is generalized by \fullref{thm:partially_ordered_cofinal_equivalences}.

    \thmitem{ex:def:cofinal_set/integers} Consider the set \( \BbbZ \) of integers. Clearly the set \( 2\BbbZ \) of even integers is cofinal. This is generalized by \fullref{thm:totally_ordered_cofinal_equivalences}.

    \thmitem{ex:def:cofinal_set/net_convergence} Cofinal sets are encountered in topology when discussing convergence of nets --- see \fullref{subsec:net_convergence}.

    \thmitem{ex:def:cofinal_set/regular_cardinals} \hyperref[def:regular_cardinal]{Regular cardinals} are equal to their own \hyperref[def:cofinality]{cofinality}.
  \end{itemize}
\end{example}

\begin{proposition}\label{thm:partially_ordered_cofinal_equivalences}
  Let \( (P, \leq) \) be a \hyperref[def:extremal_points/upper_and_lower_bounds]{bounded from above} partially ordered set and let \( A \subseteq P \). Then \( A \) is \hyperref[def:cofinal_set]{cofinal} if and only if it contains the \hyperref[def:extremal_points/top_and_bottom]{top element} \( \top \).
\end{proposition}
\begin{comments}
  \item There is a somewhat similar result case of totally ordered sets in \fullref{thm:totally_ordered_cofinal_equivalences}.
\end{comments}
\begin{proof}
  \SufficiencySubProof Let \( A \) be a cofinal set. Then \( A \) must contain an element \( x \) such that \( \top \leq x \). But \( \top \) is a maximum and hence \( x = \top \) and thus \( \top \in A \).

  \NecessitySubProof Let \( A \) be a set containing \( \top \). Then for any \( x \in P \) we have \( x \leq \top \) and hence \( A \) is cofinal.
\end{proof}

\begin{lemma}\label{thm:unbounded_totally_ordered_set}
  A totally ordered set is \hyperref[def:extremal_points/upper_and_lower_bounds]{unbounded from above} if and only if, for every element, there exists some strictly larger element.
\end{lemma}
\begin{proof}
  Fix a totally ordered set \( (P, \leq) \).

  Expanding the definition of boundedness, we conclude that
  \begin{displayquote}
    \( P \) is unbounded if and only if there does not exist \( x \in P \) such that, for every \( y \in P \), \( y \leq x \).
  \end{displayquote}

  The consequent of the above statement has a negation in front of its outermost quantifier, hence we can move the negation inwards\footnote{This is justified by \fullref{thm:first_order_quantifiers_are_dual}} to obtain
  \begin{displayquote}
    \( P \) is unbounded if and only if, for every \( x \in P \), there exist \( y \in P \) such that \( y \leq x \) does not hold.
  \end{displayquote}

  Taking into account that \( P \) is totally ordered, \( y \leq x \) does not hold if and only if \( y > x \) holds. Then
  \begin{displayquote}
    \( P \) is unbounded if and only if, for every \( x \in P \), there exist \( y \in P \) such that \( y > x \).
  \end{displayquote}
\end{proof}

\begin{proposition}\label{thm:totally_ordered_cofinal_equivalences}
  Let \( (P, \leq) \) be an \hyperref[def:extremal_points/upper_and_lower_bounds]{unbounded from above} totally ordered set and let \( A \subseteq P \). Then \( A \) is \hyperref[def:cofinal_set]{cofinal} if and only if it is itself unbounded from above.
\end{proposition}
\begin{comments}
  \item There is a somewhat similar result in \fullref{thm:partially_ordered_cofinal_equivalences}.
  \item This equivalence is useful for \hyperref[def:regular_cardinal]{regular cardinals} --- for example \fullref{thm:cardinal_cofinality}.
\end{comments}
\begin{proof}
  We will prove the converse - a set is bounded if and only if it is not cofinal.

  \SufficiencySubProof Let \( A \) be bounded from above by \( x \). \Fullref{thm:unbounded_totally_ordered_set} then implies that there exists some \( y \) such that \( x < y \). But no element of \( A \) is greater than \( y \), implying that \( A \) is not cofinal.

  \NecessitySubProof Suppose that \( A \) is not cofinal. Let \( x \in P \). Then, for every \( y \in A \), \( x < y \) does not hold. That is, \( y \leq x \) for every \( y \in A \). Therefore, \( x \) is bounded.
\end{proof}

\begin{corollary}\label{thm:natural_number_cofinal_subsets}
  For a set of positive integers, the following are equivalent:
  \begin{thmenum}
    \thmitem{thm:natural_number_cofinal_subsets/cofinal} It is \hyperref[def:cofinal_set]{cofinal}
    \thmitem{thm:natural_number_cofinal_subsets/unbounded} It is unbounded.
    \thmitem{thm:natural_number_cofinal_subsets/infinite} It is infinite.
  \end{thmenum}
\end{corollary}
\begin{proof}
  A set of positive integers is infinite if and only if it is unbounded, which \fullref{thm:totally_ordered_cofinal_equivalences} implies holds if and only if it is cofinal.
\end{proof}

\begin{proposition}\label{thm:total_lexicographic_order_is_total_order}
  If \( (P, \leq_P) \) and \( (Q, \leq_Q) \) are \hyperref[def:totally_ordered_set]{totally ordered sets}, then the \hyperref[eq:def:lexicographic_order]{lexicographic} and \hyperref[eq:def:lexicographic_order/reverse]{reverse lexicographic} orders on \( P \times Q \) are \hyperref[def:totally_ordered_set]{strict total order} relations.
\end{proposition}
\begin{comments}
  \item An analogous result holds for partial orders (\fullref{thm:lexicographic_order_is_partial_order}) and well-ordered sets (\fullref{thm:well_ordered_lexicographic_order_is_well_ordered}).
\end{comments}
\begin{proof}
  We have already shown in \fullref{thm:lexicographic_order_is_partial_order} and these are partial orders. It only remains to check trichotomy.

  \SubProofOf[def:binary_relation/trichotomic]{trichotomy} Let \( \prec \) be the lexicographic order on \( P \times Q \). Let \( (a, b) \) and \( (c, d) \) be pairs in \( P \times Q \). Since \( <_P \) and \( <_Q \) are strict total orders, we only have the following possibilities:
  \begin{itemize}
    \item If \( a = c \) and \( b = d \), then \( (a, b) = (c, d) \).
    \item If \( a = c \) and \( b <_Q d \), then \( (a, b) \prec (c, d) \).
    \item If \( a = c \) and \( b >_Q d \), then \( (a, b) \succ (c, d) \).
    \item If \( a <_P c \), then \( (a, b) \prec (c, d) \).
    \item If \( a >_P c \), then \( (a, b) \succ (c, d) \).
  \end{itemize}

  The proof for the reverse lexicographic order is analogous.
\end{proof}

\begin{definition}\label{def:stabilizing_chain}\mimprovised
  We say that the \hyperref[def:sequence]{sequence} \( x_1, x_2, \ldots \) is an \term{ascending} if \( x_k \leq x_{k+1} \) for every \( k = 1, 2, \ldots \) and a \term{descending} if \( x_k \geq x_{k+1} \) instead. They are also referred to as \enquote{ascending/descending chains}.

  If there exists an index \( n \) such that \( x_k = x_n \) for \( k > n \). we say that the chain \term{stabilizes} at \( n \).
\end{definition}

\begin{definition}\label{def:chain_condition}
  We say that a \hyperref[def:partially_ordered_set]{partially ordered set} \( (P, \leq) \) satisfies the \term{ascending chain condition} (resp. \term{descending chain condition}) if any of the following equivalent conditions hold:
  \begin{thmenum}
    \thmitem{def:chain_condition/maximal}\mcite[37]{Birkhoff1948} Every nonempty subset of \( P \) has a \hyperref[def:extremal_points/maximal_and_minimal_element]{maximal} (resp. minimal) element.

    \medskip

    \thmitem{def:chain_condition/stabilization}\mcite[thm. 5]{Birkhoff1948} Every nonstrict ascending (resp. descending) sequence \hyperref[def:stabilizing_chain]{stabilizes}.

    \medskip

    \thmitem{def:chain_condition/infinite}\mcite[thm. 9A]{Enderton1977Sets} There exists no strictly infinitely ascending (resp. descending) sequence in \( P \).

    \medskip
  \end{thmenum}
\end{definition}
\begin{defproof}
  We will restrict ourselves to ascending chains and maximal elements.

  \ImplicationSubProof{def:chain_condition/maximal}{def:chain_condition/stabilization} Suppose that every subset of \( P \) has a maximal element.

  Suppose that there is a nonstrict ascending sequence
  \begin{equation*}
    x_1 \leq x_2 \leq x_3 \leq \cdots.
  \end{equation*}

  Then the set \( \set{ x_k \given k \geq 1 } \) has a maximal element, say \( x_{k_0} \). Then, since \( x_{k_0} \) is maximal, using \hyperref[rem:induction/peano_arithmetic]{natural number induction} we can prove that \( x_{k_0} = x_{k_0 + i} \) for \( i \geq 0 \).

  Thus, the sequence stabilizes.

  \ImplicationSubProof{def:chain_condition/stabilization}{def:chain_condition/infinite} Suppose that every ascending chain in \( P \) stabilizes.

  Every strictly ascending sequence is also nonstrictly ascending, hence it stabilizes, which contradicts our assumption that it is strictly ascending. The obtained contradiction shows that there are not strictly ascending sequences in \( P \).

  \ImplicationSubProof{def:chain_condition/infinite}{def:chain_condition/maximal} Suppose that there exists no infinite strictly ascending sequence in \( P \).

  Let \( A \) be a nonempty subset of \( P \). Aiming at a contradiction, suppose that it has no maximal element.

  Pick \( x_1 \in A \). Since \( x_1 \) is not maximal, there must exist some element \( x_2 \in A \) such that \( x_1 < x_2 \). Using \hyperref[rem:natural_number_recursion]{natural number recursion}, we can build an infinitely strongly ascending sequence
  \begin{equation*}
    x_1 < x_2 < x_3 < \cdots.
  \end{equation*}

  But the existence of such a chain contradicts our assumption. Therefore, every nonempty set in \( P \) must have a maximal element.
\end{defproof}

\begin{definition}\label{def:well_ordered_set}\mcite[37]{Birkhoff1948}
  We say that a \hyperref[def:totally_ordered_set]{totally ordered set} \( (P, \leq) \) is \term{well-ordered} if it satisfies the \hyperref[def:chain_condition]{descending chain condition}.
\end{definition}

\begin{proposition}\label{thm:finite_totally_ordered_set_is_well_ordered}
  Every finite \hyperref[def:totally_ordered_set]{totally ordered set} is \hyperref[def:well_ordered_set]{well-ordered}.
\end{proposition}
\begin{proof}
  Let \( (P, \leq) \) be totally ordered set of finite cardinality \( n \). Let \( A \) be any nonempty subset of \( P \). We will show by induction on \( \card(A) \) that \( A \) has a minimum.

  \begin{itemize}
    \item If \( A = \set{ a } \), then \( a \) is vacuously a minimum.
    \item Suppose that all subsets of \( P \) of cardinality \( k - 1 \) have a minimum. Let \( \card(A) = k \). Pick an arbitrary element \( a \in A \). By the inductive hypothesis, \( A \setminus \set{ a } \) has a minimum, say \( b \). Then \( \min\set{ a, b } \) is a minimum of \( A \).
  \end{itemize}
\end{proof}

\begin{lemma}\label{thm:well_ordered_embedding_extensive}
  Any \hyperref[def:order_homomorphism/embedding]{order embedding} on a well-ordered set is \hyperref[def:extensive_function]{extensive}.

  That is, if \( (P, \leq) \) is a well-ordered set and \( f: P \to P \) is an order embedding, then \( x \leq f(x) \) for any \( x \in P \).
\end{lemma}
\begin{proof}
  We proceed by induction on \( P \). Fix \( x_0 \) and suppose that \( y \leq f(y) \) for all \( y < x_0 \).

  Aiming at a contradiction, suppose that \( f(x_0) < x_0 \). Thus, there exists some \( y_0 < x_0 \) such that \( f(x_0) = y_0 \). By the inductive hypothesis we have \( f(x_0) = y_0 \leq f(y_0) \). Thus, either \( f(x_0) < f(y_0) \), which contradicts that \( f \) is an order homomorphism, or \( f(x_0) = f(y_0) \), which contradicts the injectivity of \( f \).

  The obtained contradiction demonstrates that \( x \leq f(x) \) for all \( x \in P \).
\end{proof}

\begin{proposition}\label{thm:well_ordered_isomorphism_is_unique}
  There is at most one isomorphism between any pair of well-ordered sets.
\end{proposition}
\begin{proof}
  Fix two well-ordered sets \( (P, \leq_P) \) and \( (Q, \leq_Q) \) and let \( f: P \to Q \) and \( g: P \to Q \) be two order isomorphisms.

  For any member \( x \) of \( P \) we have the following possibilities:
  \begin{itemize}
    \item If \( f(x) <_Q g(x) \), then \( g^{-1}(f(x)) <_P x \) and \( f(g^{-1}(f(x))) <_Q f(x) \). Then we can use \hyperref[rem:natural_number_recursion]{natural number recursion} to build an infinitely descending sequence of members of \( Q \). But this is a contradiction because \( Q \) satisfies the descending chain condition.

    \item We can build a similar sequence if \( g(x) <_Q f(x) \).

    \item It remains for \( f(x) = g(x) \) to hold.
  \end{itemize}

  Since \( x \in P \) was arbitrary, we conclude that \( f = g \).
\end{proof}

\begin{proposition}\label{thm:well_ordered_lexicographic_order_is_well_ordered}
  If \( (P, \leq_P) \) and \( (Q, \leq_Q) \) are \hyperref[def:well_ordered_set]{well-ordered sets}, then the \hyperref[eq:def:lexicographic_order]{lexicographic} and \hyperref[eq:def:lexicographic_order/reverse]{reverse lexicographic} orders on \( P \times Q \) are well-ordering relations.
\end{proposition}
\begin{comments}
  \item An analogous result holds for partial orders (\fullref{thm:lexicographic_order_is_partial_order}) and total orders (\fullref{thm:total_lexicographic_order_is_total_order}).
\end{comments}
\begin{proof}
  We have already shown in \fullref{thm:total_lexicographic_order_is_total_order} and these total orders. It only remains to check the \hyperref[def:chain_condition]{descending chain condition}.

  Let \( \prec \) be the lexicographic order on \( P \times Q \).

  Suppose that there exists an infinitely descending sequence
  \begin{equation*}
    \cdots \prec (a_3, b_3) \prec (a_2, b_2) \prec (a_1, b_1).
  \end{equation*}

  Since \( P \) satisfies the descending chain condition, the corresponding sequence \( \seq{ a_k }_{k=1}^\infty \). Then there exists an index \( k_0 \) such that \( a_{k_0} = a_k \) for \( k \geq k_0 \). Therefore, the sequence \( \set{ b_k }_{k=k_0}^\infty \) must be infinitely descending. But this contradicts the chain condition imposed on \( Q \).

  Therefore, no infinitely descending sequence exists in the totally ordered set \( (P \times Q, \prec) \), and thus it is well-ordered.
\end{proof}

\begin{definition}\label{def:order_topology}\mcite[exer. 1.7.4(a)]{Engelking1989}
  Let \( P \) be a \hyperref[def:partially_ordered_set]{totally ordered set} with more than one element. The \term{order topology} induced by \( \leq \) is the topology generated by the \hyperref[def:topological_subbase]{subbase} of open \hyperref[def:order_interval/ray]{rays}
  \begin{equation}\label{eq:def:order_topology/subbase}
    \mscrS \coloneqq \set[\Big]{ (a, \infty) \given a \in P } \cup \set[\Big]{ (-\infty, b) \given b \in P }.
  \end{equation}

  The \hyperref[def:topological_base]{base} corresponding to this subbase is
  \begin{equation}\label{eq:def:order_topology/base}
    \mscrB = \mscrS \cup \set[\Big]{ \varnothing } \cup \set[\Big]{ (a, b) \given a, b \in P \T{and} a < b }.
  \end{equation}

  See our proof of \hyperref[thm:topology_from_base/B1]{B1} for why \( P \) must have more than one element.
\end{definition}
\begin{defproof}
  \SubProof{Proof of compatibility of \( \mscrS \) and \( \mscrB \)} Define
  \begin{equation*}
    \mscrC = \set*{ \bigcap \mscrS \given* \mscrS \T{is a nonempty finite subset of} \mscrS }.
  \end{equation*}

  We will show that \( \mscrB = \mscrC \).

  Let \( B \in \mscrB \). The cases \( B \in \mscrS \) and \( B = \varnothing \) are trivial. Suppose that \( B \not\in \mscrS \). Then there exist points \( a < b \) such that
  \begin{equation*}
    B = (a, b) = (-\infty, b) \cap (a, \infty).
  \end{equation*}

  This is an intersection of members of \( \mscrS \), hence \( B \in \mscrC \). Therefore, \( \mscrB \subseteq \mscrC \).

  Now let \( C = S_1 \cap \cdots \cap S_n \), where \( S_1, \ldots, S_n \) are members of \( \mscrS \). We will show by induction on \( n > 0 \) that \( C \in \mscrB \). The case \( n = 1 \) is trivial. Suppose that all \( n \)-ary intersections belong to \( \mscrB \) and let
  \begin{equation*}
    C = S_1 \cap \cdots \cap S_n \cap S_{n+1}.
  \end{equation*}

  By the inductive hypothesis, we have that \( D \coloneqq S_1 \cap \cdots \cap S_n \) belongs to \( \mscrB \) and thus we have three cases:
  \begin{itemize}
    \item If either \( D = (a, \infty) \) or \( D = (-\infty, b) \), then \( D \in \mscrS \).
    \item If \( D = \varnothing \), then \( D = (-\infty, a) \cap (a, \infty) \) for some \( a \in P \).
    \item If \( D = (a, b) \), then \( D = (-\infty, b) \cap (a, \infty) \).
  \end{itemize}

  In all cases both \( D \) and \( C = D \cap S_{n+1} \) are finite intersection of members of \( \mscrS \). Therefore, \( \mscrC \subseteq \mscrB \). Since we already have the inclusion in the other direction, we conclude that \( \mscrC = \mscrB \).

  \SubProof{Proof that \( \mscrB \) is a base} We will show that the axioms in \fullref{thm:topology_from_base} hold.

  \SubProofOf*[thm:topology_from_base/B1]{B1} Let \( x \in P \).

  If \( x \) is a \hyperref[def:extremal_points/maximum_and_minimum]{maximum}, then take any other value \( y < x \) and the set \( (y, \infty) \) will contain \( x \). We use here that there is more than one element in \( P \).

  If \( x \) is not a maximum, then \( x \) belongs to any interval \( (-\infty, y) \) whenever \( y > x \).

  In both cases there exists an interval in \( S \) containing \( x \). Thus, \( \bigcup S = P \).

  \SubProofOf*[thm:topology_from_base/B2]{B2} Let \( U \) and \( V \) be members of \( \mscrB \). We consider \( 14 \) cases:
  \begin{itemize}
    \item If either \( U = \varnothing \) or \( V = \varnothing \), then \( U \cap V = \varnothing \).
    \item If \( U = (-\infty, u) \) and \( V = (v, \infty) \), then
    \begin{itemize}
      \item If \( u \leq v \), then \( U \cap V = \varnothing \).
      \item If \( v < u \), then \( U \cap V = (v, u) \).
    \end{itemize}

    \item If \( U = (-\infty, u) \) and \( V = (v_1, v_2) \), then
    \begin{itemize}
      \item If \( u \leq v_1 \), then \( U \cap V = \varnothing \).
      \item If \( v_1 < v_2 \leq u \), then \( U \cap V = V =  (v_1, v_2) \).
      \item If \( v_1 \leq u < v_2 \), then \( U \cap V = (v_1, u) \).
    \end{itemize}

    \item If \( U = (u_1, u_2) \) and \( V = (v, \infty) \), then
    \begin{itemize}
      \item If \( u_2 \leq v \), then \( U \cap V = \varnothing \).
      \item If \( u_1 \leq v < u_2 \), then \( U \cap V = (v, u_2) \).
      \item If \( v \leq u_1 < u_2 \), then \( U \cap V = U = (u_1, v_1) \).
    \end{itemize}

    \item If \( U = (u_1, u_2) \) and \( V = (v_1, v_2) \), then
    \begin{itemize}
      \item If \( u_2 < v_1 \), then \( U \cap V = \varnothing \).
      \item If \( u_1 < v_1 < u_2 < v_2 \) then \( U \cap V = (v_1, u_2) \).
      \item If \( u_1 < v_1 < v_2 < u_2 \) then \( U \cap V = V = (v_1, v_2) \).
      \item If \( v_1 < u_1 < u_2 < v_2 \) then \( U \cap V = U = (u_1, u_2) \).
      \item If \( v_1 < u_1 < v_2 < u_2 \) then \( U \cap V = (u_1, v_2) \).
    \end{itemize}
  \end{itemize}

  In all cases, the intersection \( U \cap V \) belongs to \( \mscrB \).
\end{defproof}

\begin{proposition}\label{thm:order_topology_intervals}
  Under the \hyperref[def:order_topology]{order topology} \( \mscrT \) on a \hyperref[def:totally_ordered_set]{totally ordered set} \( (P, \leq) \):
  \begin{thmenum}
    \thmitem{thm:order_topology_intervals/open} \hyperref[def:order_interval/ray]{Open rays} and \hyperref[def:order_interval/open]{open intervals} are open sets.

    \thmitem{thm:order_topology_intervals/closed} \hyperref[def:order_interval/ray]{Closed rays} and \hyperref[def:order_interval/closed]{closed intervals} are closed sets.
  \end{thmenum}
\end{proposition}
\begin{proof}
  \SubProofOf{thm:order_topology_intervals/open} Open rays are members of the subbase \eqref{eq:def:order_topology/subbase}, which makes them open.

  The open interval \( (a, b) \) is a members of the base \eqref{eq:def:order_topology/base}. Hence, it is also open.

  \SubProofOf{thm:order_topology_intervals/closed} The closed ray \( (-\infty, b] \) is the complement of the open ray \( (a, \infty) \), which makes it a closed set. Similarly, \( [a, \infty) \) is the complement of \( (-\infty, b) \), making it a closed set.

  For the closed interval \( [a, b] \), we have
  \begin{equation*}
    [a, b] = (-\infty, b] \cap [a, \infty).
  \end{equation*}

  Both \( (-\infty, b] \) and \( [a, \infty) \) are closed rays, and hence their union is also a closed set. Hence, \( [a, b] \) is closed.
\end{proof}

\begin{example}\label{ex:def:order_topology}
  Examples of \hyperref[def:order_topology]{order topologies} include:
  \begin{itemize}
    \item The order topology on \( \BbbR \), which is equivalent to the \hyperref[def:metric_topology]{metric topology} as shown in \fullref{thm:real_metric_and_order_topologies_coincide}.

    \item All \hyperref[def:ordinal]{ordinals} greater than \( 1 \) induce topological spaces called the \hyperref[def:ordinal_space]{ordinal spaces}.
  \end{itemize}
\end{example}

\begin{definition}\label{def:ordinal_space}\mcite[40]{SteenSeebach1995}
  Let \( \alpha \) be an \hyperref[def:ordinal]{ordinal}. When regarded as the set of smaller ordinals, as shown valid in \fullref{thm:ordinal_is_set_of_smaller_ordinals}, \( \alpha \) is a \hyperref[def:totally_ordered_set]{totally order set} and hence we can endow it with the \hyperref[def:order_topology]{order topology} \( \mscrT \) to obtain a \hyperref[def:topological_space]{topological space}. We call the space \( (\alpha, \mscrT) \) an \term{ordinal space}.
\end{definition}

\begin{proposition}\label{thm:limit_ordinal_order_topology}
  In an \hyperref[def:ordinal_space]{ordinal space} \( (\alpha, \mscrT) \), a nonzero ordinal \( \beta \in \alpha \) is a \hyperref[def:successor_and_limit_ordinal]{limit ordinal} if and only if it is the \hyperref[def:set_cluster_point]{cluster point} of \( \alpha \).
\end{proposition}
\begin{proof}
  \SufficiencySubProof Let \( \beta \) be a limit ordinal. We need to show that every neighborhood of \( \beta \) contains points distinct from \( \beta \).

  \Fullref{thm:properties_via_bases/cluster} allows us to only consider neighborhood in a local base at \( \beta \). We will use the base of open intervals containing \( \beta \).
  \begin{itemize}
    \item For \( \gamma < \beta < \delta \), the interval \( (\gamma, \delta) \) must contain \( \op{succ}(\gamma) \) because, as a limit ordinal, \( \beta \) contains the successors of all smaller ordinals.

    \item The same proof applies to the interval \( (\gamma, \infty) \).

    \item For \( \beta < \delta \), the interval \( (-\infty, \delta) \) contains all predecessors of \( \beta \). It is important here that \( \beta \) is nonzero, which we implicitly assume for limit ordinals.
  \end{itemize}

  Therefore, every neighborhood of the local base at \( \beta \) contains points distinct from \( \beta \), making \( \beta \) a cluster point.

  \NecessitySubProof Let \( \beta \) be a cluster point and let \( \gamma < \beta \). Then the set
  \begin{equation*}
    (\gamma, \op{succ}(\beta)) \setminus \set{ \beta }
    =
    (\gamma, \beta)
  \end{equation*}
  is nonempty, implying that \( \beta \) is not the successor of \( \gamma \). Since this holds for all smaller ordinals, we conclude that \( \beta \) is a limit ordinal.
\end{proof}
