\subsection{Zermelo-Fraenkel set theory}\label{subsec:zermelo_fraenkel_set_theory}

\begin{definition}\label{def:choice_function}\mcite[def. 6.4.1]{Hinman2005}
  Let \( \mscrA \) be a (potentially empty) family of nonempty sets. A \term{choice function} on \( \mscrA \) is a \hyperref[def:function]{function} \( c: \mscrA \to \bigcup \mscrA \) such that \( c(A) \in A \) for all \( A \in \mscrA \).
\end{definition}
\begin{comments}
  \item This is formally the same as a tuple defined in \fullref{def:cartesian_product/tuple}.

  \item A choice function on \( \mscrA \) \enquote{chooses} an element out of each member of \( \mscrA \). We sometimes have a canonical choice function, for example in \fullref{thm:representatives_in_univariate_polynomial_quotient_set}, however for general \hyperref[def:equivalence_relation/quotient]{quotient sets} the existence of a choice function is not by any means obvious.

  \item Furthermore, even the existence of a choice function for a concrete family may be difficult to prove. We guarantee its existence via an important axiom of \logic{ZFC} --- see \fullref{def:zfc/choice}.
\end{comments}

\begin{definition}\label{def:zfc}\mcite[def. 6.1.6]{Hinman2005}
  The \hyperref[def:first_order_theory]{first-order theory} commonly abbreviated as \term{\logic{ZFC}} is based on the same language as \hyperref[def:naive_set_theory]{na\"ive set theory}, but with different axioms. The three letters refer to:
  \begin{itemize}
    \item \textbf{Z}ermelo, who formulated the entire theory except for the \hyperref[def:zfc/replacement]{axiom schema of replacement} and the \hyperref[def:zfc/foundation]{axiom of foundation}.
    \item \textbf{F}raenkel, who simultaneously with Skolem reformulated the theory within first-order logic while also introducing the axiom schema of replacement.
    \item The \hyperref[def:zfc/choice]{axiom of \textbf{c}hoice}, which is part of Zermelo's original theory, but is controversial enough to attract special attention --- see \fullref{thm:axiom_of_choice_equivalences}.
  \end{itemize}

  We are usually only interested in the entire theory. If we wish to avoid the axiom of choice --- for example when proving the equivalences in \fullref{thm:axiom_of_choice_equivalences} --- we can instead use \logic{ZF}, which excludes the axiom of choice. The abbreviation of the latter theory is inaccurate historically, but is nevertheless established.

  If we wish to instead exclude the axiom schema of replacement, we obtain the theory \logic{Z}, however without context it is unclear whether the axiom of choice is included in \logic{Z} or not.

  The full list of axioms is:
  \begin{thmenum}
    \thmitem{def:zfc/extensionality} The \term{axiom of extensionality}, as defined in \fullref{def:naive_set_theory/extensionality}. This is also the only axiom of the theory that does not deal with existence.

    \thmitem{def:zfc/specification} The \term{axiom schema of specification}, also known as the axiom schema of \term{separation} or of \term{restricted comprehension}, states that, given a set \( A \), any formula defines a subset of \( A \).

    More precisely, for each formula \( \varphi \) containing neither \( \sigma \) nor \( \rho \) as free variables, but possibly containing \( \xi \), the following is an axiom:
    \begin{equation}\label{eq:def:zfc/specification}
      \qforall \sigma \qexists \rho \qforall \xi \parens[\Big]{ \xi \in \rho \syniff (\varphi \synwedge \xi \in \sigma) }.
    \end{equation}

    As explained in \fullref{def:naive_set_theory/unrestricted_comprehension} and \fullref{def:set_builder_notation}, this set may depend on parameters, which are other sets. We must formally take the \hyperref[def:universal_closure]{universal closure} of this formula to quantify over all possible values for the parameters.

    Compare this axiom to \hyperref[def:naive_set_theory/unrestricted_comprehension]{unrestricted comprehension}. Informally, restricted comprehension axiom can be obtained by taking the result of unrestricted comprehension and intersecting it with some set \( A \). As mentioned in \fullref{def:set_builder_notation}, in set-builder notation such a set can be denoted via
    \begin{equation*}
      \set[\Big]{ x \in B \given* \Bracks{\varphi}(x, u_1, \ldots, u_n) },
    \end{equation*}
    where \( \Bracks{\varphi}(\cdots) \) is the formula valuation function from \fullref{def:propositional_valuation/formula_valuation_function}.

    Unlike for unrestricted comprehension, some sets definable in the metatheory no longer have a corresponding set within the object logic.

    \thmitem{def:zfc/power_set} The \term{axiom of power sets} states that every set has a corresponding \hyperref[def:basic_set_operations/power_set]{power set}. In symbols:
    \begin{equation}\label{eq:def:zfc/power_set}
      \qforall \sigma \qexists \rho \ref{eq:def:basic_set_operations/power_set/predicate}[\rho, \sigma].
    \end{equation}

    \thmitem{def:zfc/union} The \term{axiom of unions} states that for every set there exists another set that is its \hyperref[def:basic_set_operations/union]{union}. In symbols:
    \begin{equation}\label{eq:def:zfc/union}
      \qforall \sigma \qexists \rho \ref{eq:def:basic_set_operations/union/predicate}[\rho, \sigma].
    \end{equation}

    \thmitem{def:zfc/pairing} The \term{axiom of pairing} states that for any sets \( A \) and \( B \) there exists an unordered pair containing them, i.e. another set that contains exactly \( A \) and \( B \). This is \( \set{ A, B } \) in set-builder notation. In symbols:
    \begin{equation}\label{eq:def:zfc/pairing}
      \qforall \sigma \qforall \tau \qexists \rho \qforall \xi \parens[\Big]{ \xi \in \rho \syniff ( \xi \syneq \sigma \synvee \xi \syneq \tau) }.
    \end{equation}

    \thmitem{def:zfc/infinity} The \term{axiom of infinity} states that an \hyperref[def:inductive_set]{inductive set} exists. In symbols:
    \begin{equation}\label{eq:def:zfc/infinity}
      \qexists \sigma \ref{eq:def:inductive_set/predicate}[\sigma].
    \end{equation}

    As will be shown in \fullref{thm:cumulative_hierarchy_model_of_zfc_without_infinity}, no other axioms of \logic{ZFC} ensure the existence of at least one infinite set.

    \thmitem{def:zfc/choice}\mcite[def. 6.4.1]{Hinman2005} The \term{axiom of choice} states that a \hyperref[def:choice_function]{choice function} exists for any family of nonempty sets. To state the axiom via a formula, we will avoid functions and only state it in terms of the image of the choice function. That is, we will formulate that for each family \( \mscrA \) of nonempty sets, there exists a set \( B \) such that \( A \cap B \) is a singleton set for each \( A \in \mscrA \). In symbols:
    \begin{equation}\label{eq:def:zfc/choice}
      \qforall \sigma \parens[\Bigg]
        {
          \parens[\Big]{ \qforall {\xi \in \sigma} \synneg \ref{eq:def:empty_set/predicate}[\xi] }
          \synimplies
          \parens[\Big]{ \qexists \rho \qforall {\xi \in \sigma} \qExists {\eta \in \rho} \eta \in \xi }.
        }
    \end{equation}

    See \fullref{thm:axiom_of_choice_equivalences} for more statements equivalent to this axiom.

    \thmitem{def:zfc/replacement} The \term{axiom schema of replacement} roughly states that every \hyperref[rem:function_definition]{mapping} that is definable via a formula of \logic{ZFC} is a function. As we have done for the \hyperref[def:zfc/choice]{axiom of choice}, we only formulate the axiom via the image of the function. More concretely, given a formula \( \varphi \) not containing \( \sigma \) nor \( \rho \) as free variables, but possibly containing \( \xi \) and \( \eta \), the following is an axiom:
    \begin{equation}\label{eq:def:zfc/replacement}
      \qforall \sigma \parens[\Bigg]
        {
          \parens[\Big]{ \qforall {\xi \in \sigma} \qExists \eta \varphi }
          \synimplies
          \parens[\Big]{ \qexists \rho \qforall \eta \parens[\Big]{ \eta \in \rho \syniff \qexists {\xi \in \sigma} \varphi } }
        }.
    \end{equation}

    As is the case with the \hyperref[def:zfc/specification]{axiom schema of specification}, the formula \( \varphi \) may depend on parameters, in which case we use its \hyperref[def:universal_closure]{universal closure}.

    This axiom is useful in cases where it is impossible or at least difficult to construct a function, for example in \fullref{thm:zfc_existence_theorems/indexed_family} or \fullref{thm:hartogs_lemma}. This is the case, in general, when dealing with \hyperref[thm:zfc_existence_theorems/indexed_family]{indexed families} rather than \hyperref[def:function]{functions}.

    This is the axiom that makes \logic{ZFC} require large models --- see \fullref{thm:cumulative_hierarchy_model_of_zfc}.

    \thmitem{def:zfc/foundation} The \term{axiom of foundation} states that every nonempty set contains a member disjoint from the set itself. In symbols:
    \begin{equation}\label{eq:def:zfc/foundation}
      \qforall \sigma \parens[\Big]
        {
          \synneg \ref{eq:def:empty_set/predicate}[\sigma]
          \synimplies
          \qexists {\rho \in \sigma} \synneg \qexists \xi \parens{ \xi \in \sigma \synwedge \xi \in \rho }
        }.
    \end{equation}

    This is a powerful axiom because it shows that set membership in \logic{ZFC} is well-founded --- see \fullref{thm:set_membership_is_well_founded}. It is equivalent to \fullref{thm:axiom_of_regularity} and is often itself called the \term{axiom of regularity}.
  \end{thmenum}
\end{definition}

\begin{proposition}\label{thm:zfc_existence_theorems}
  We will now prove that all sets we have considered up until now in \fullref{sec:set_theory} are sets in \hyperref[def:zfc]{\logic{ZFC}}. The uniqueness in all cases follows from the \hyperref[def:zfc/extensionality]{axiom of extensionality}.

  A very fundamental existence result is provided by the fact that we are assuming \hyperref[rem:standard_model_of_set_theory]{standard} and \hyperref[rem:transitive_model_of_set_theory]{transitive} models of \logic{ZFC}. Let \( \mscrV = (V, I) \) be such a model. Then, if \( v \in V \) and \( u \in v \), transitivity implies that \( u \in V \). Since the model is also standard, this shows that both \( u \) and \( v \) are sets within the object theory. Thus, if \( A \) is a set within the object theory and if \( B \in A \) within the metatheory, then necessarily \( B \) itself is a set within the object theory.

  For example, \fullref{thm:zfc_existence_theorems/set_of_functions} shows that the set \( \fun(A, B) \) of functions exists within the object theory for any two sets \( A \) and \( B \) in the object theory. Therefore, every single function between \( A \) and \( B \) is a set within the object theory because it is a member of \( \fun(A, B) \).

  With that in mind, we will show the following:

  \begin{thmenum}
    \thmitem{thm:zfc_existence_theorems/set_containing_all} There is no set that contains all sets.

    \thmitem{thm:zfc_existence_theorems/subset} If \( A \) is a set, then for any formula \( \varphi \) the set
    \begin{equation*}
      \set{ x \in A \given \Bracks{\varphi}(x) }.
    \end{equation*}
    exists and, furthermore, it is a \hyperref[def:subset]{subset} of \( A \).

    Only \hyperref[def:first_order_definability]{definable subsets} of \( A \) can be described in this way, however. See \fullref{thm:zfc_existence_theorems/power_set}.

    \thmitem{thm:zfc_existence_theorems/empty_set} There exists a unique \hyperref[def:empty_set]{empty set}.

    \thmitem{thm:zfc_existence_theorems/singleton} For every set \( A \), there exists a \hyperref[rem:singleton_sets]{singleton set} \( \set{ A } \) that contains only \( A \).

    \thmitem{thm:zfc_existence_theorems/arbitrary_intersection} For any \hi{nonempty} family \( \mscrA \), the \hyperref[def:basic_set_operations/intersection]{intersection} \( \bigcap \mscrA \) exists.

    \thmitem{thm:zfc_existence_theorems/binary_intersection} For any two sets \( A \) and \( B \), their \hyperref[def:basic_set_operations/intersection]{intersection} \( A \cap B \) exists.

    \thmitem{thm:zfc_existence_theorems/arbitrary_union} For any family \( \mscrA \), the \hyperref[def:basic_set_operations/union]{union} \( \bigcup \mscrA \) exists.

    \thmitem{thm:zfc_existence_theorems/binary_union} For any two sets \( A \) and \( B \), their \hyperref[def:basic_set_operations/union]{union} \( A \cup B \) exists.

    \thmitem{thm:zfc_existence_theorems/difference} For any two sets \( A \) and \( B \), their \hyperref[def:basic_set_operations/difference]{difference} \( A \setminus B \) exists.

    \thmitem{thm:zfc_existence_theorems/power_set} For any set \( A \), its \hyperref[def:basic_set_operations/power_set]{power set} \( \pow(A) \) exists.

    As a consequence, even subsets of \( A \) which are not \hyperref[def:first_order_definability]{definable} exist.

    \thmitem{thm:zfc_existence_theorems/successor} For any set \( A \), its \hyperref[def:ordinal_successor]{successor} \( \op{succ}(A) \) exists.

    \thmitem{thm:zfc_existence_theorems/kuratowski_pair} For any two sets \( A \) and \( B \), their \hyperref[def:cartesian_product/kuratowski_pair]{Kuratowski pair} \( \braket{ A, B } \) exists.

    \thmitem{thm:zfc_existence_theorems/indexed_family} For any sets \( \mscrK \) and \( \mscrA \), any \hyperref[def:cartesian_product/indexed_family]{indexed family} \( \seq{ A_k }_{k \in \mscrK} \subseteq \mscrA \) exists.

    \thmitem{thm:zfc_existence_theorems/cartesian_product} For any indexed family \( \set{ A_k }_{k \in \mscrK} \), its \hyperref[def:cartesian_product]{Cartesian product} \( \prod_{k \in \mscrK} A_k \) exists.

    \thmitem{thm:zfc_existence_theorems/set_of_relations} For any two sets \( A \) and \( B \), the set of all relations between \( A \) and \( B \) exists.

    \thmitem{thm:zfc_existence_theorems/set_of_functions} For any two sets \( A \) and \( B \), the set \hyperref[def:function/set_of_functions]{\( \fun(A, B) \)} of functions between them exists.

    \thmitem{thm:zfc_existence_theorems/quotient_set} For any set \( A \) and any \hyperref[def:equivalence_relation]{equivalence relation} \( \cong \), the \hyperref[def:equivalence_relation/quotient]{quotient set} \( A / {\cong} \) exists.

    \thmitem{thm:zfc_existence_theorems/function_evaluation} Fix some sets \( A \) and \( B \) and some indexed family of functions \( \seq{ f_k }_{k \in \mscrK} \) where \( f_k: A \to B_k \) for \( k \in \mscrK \).

    For any \( x \in A \) the corresponding tuple \( \seq{ f_k(x) }_{k \in \mscrK} \) exists.
  \end{thmenum}
\end{proposition}
\begin{proof}
  \SubProofOf{thm:zfc_existence_theorems/set_containing_all} Suppose that \( V \) is a set containing all sets. Take the subset of \( V \) of all sets that do not belong to themselves. We restrict via the \hyperref[def:zfc/specification]{axiom schema of specification} our comprehension to a set \( V \) we know exists, but this nonetheless this leads to \fullref{thm:russels_paradox}. We can thus conclude that \( V \) does not exist.

  \SubProofOf{thm:zfc_existence_theorems/subset} This is a trivial consequence of the \hyperref[def:zfc/specification]{axiom schema of specification}.

  \SubProofOf{thm:zfc_existence_theorems/empty_set} As a consequence of the \hyperref[def:zfc/infinity]{axiom of infinity}, there exists at least one inductive set. Let \( A \) be an inductive set. Then from the \hyperref[def:zfc/specification]{axiom schema of specification} it follows that
  \begin{equation*}
    \set{ x \in A \given \synbot }
  \end{equation*}
  is a set. Furthermore, \( x \) belongs to this set if and only if \( x \) satisfies \( \synbot \), which is impossible, hence the set is empty.

  As a consequence of the \term{axiom of extensionality}, this empty set is unique. As discussed in \fullref{def:empty_set}, we denote this unique empty set by \( \varnothing \).

  \SubProofOf{thm:zfc_existence_theorems/singleton} Fix a set \( A \). The set \( \set{A} \), if it exists, is equal to \( \set{A} = \set{A, A} \), by the \hyperref[def:zfc/extensionality]{axiom of extensionality}.

  Thus, by the \hyperref[def:zfc/pairing]{axiom of pairing}, the singleton set \( \set{A} = \set{A, A} \) actually exists.

  \SubProofOf{thm:zfc_existence_theorems/arbitrary_intersection} Let \( \mscrA \) be a nonempty family of sets. Their intersection \( \bigcap \mscrA \), if it exists, is a subset of every set \( A \in \mscrA \).

  Therefore, since the family \( \mscrA \) is nonempty, the \hyperref[def:zfc/specification]{axiom schema of specification} applied to any set in \( A \in \mscrA \) guarantees the existence of the intersection \( \bigcap \mscrA \). More precisely, for any \( A_0 \in \mscrA \), we can define the intersection of \( A \) as
  \begin{equation*}
    \bigcap \mscrA = \set{ x \in A_0 \given \qexists {A \in \mscrA} x \in A }.
  \end{equation*}

  \SubProofOf{thm:zfc_existence_theorems/binary_intersection} For sets \( A \) and \( B \), by the \hyperref[def:zfc/pairing]{axiom of pairing} the set \( \set{ A, B } \) exists. Then by \fullref{thm:zfc_existence_theorems/arbitrary_intersection}, the binary intersection
  \begin{equation*}
    A \cap B = \bigcap \set{ A, B }
  \end{equation*}
  also exists.

  \SubProofOf{thm:zfc_existence_theorems/arbitrary_union} The existence of arbitrary unions is merely a restatement of the \hyperref[def:zfc/union]{axiom of unions}.

  \SubProofOf{thm:zfc_existence_theorems/binary_union} Similarly to \fullref{thm:zfc_existence_theorems/binary_intersection}, for sets \( A \) and \( B \), by the \hyperref[def:zfc/pairing]{axiom of pairing} the set \( \set{ A, B } \) exists and by the \hyperref[def:zfc/union]{axiom of unions}, the binary union
  \begin{equation*}
    A \cup B = \bigcup \set{ A, B }
  \end{equation*}
  exists.

  \SubProofOf{thm:zfc_existence_theorems/difference} The difference \( A \setminus B \) is guaranteed to exist by \hyperref[def:zfc/specification]{restricted comprehension}:
  \begin{equation*}
    A \setminus B = \set{ x \in A \given x \not\in B }.
  \end{equation*}

  \SubProofOf{thm:zfc_existence_theorems/power_set} The existence of power sets is a restatement of the \hyperref[def:zfc/power_set]{axiom of power sets}.

  \SubProofOf{thm:zfc_existence_theorems/successor} The successor of \( A \) is
  \begin{equation*}
    \op{succ}(A) = \set{ A } \cup A.
  \end{equation*}

  Its existence follows from \fullref{thm:zfc_existence_theorems/singleton} and \fullref{thm:zfc_existence_theorems/binary_union}.

  \SubProofOf{thm:zfc_existence_theorems/kuratowski_pair} The existence of the Kuratowski pair
  \begin{equation*}
    \braket{ A, B } = \set{ \set{ A }, \set{ A, B } }
  \end{equation*}
  can be proven by the \hyperref[def:zfc/pairing]{axiom of pairing} applied first to \( \set{ A, B } \) and then to the pair itself.

  \SubProofOf{thm:zfc_existence_theorems/indexed_family} Given an indexed family \( \seq{ A_k }_{k \in \mscrK} \), by the \hyperref[def:zfc/replacement]{axiom schema of replacement}, there exists a set
  \begin{equation*}
    \mscrA \coloneqq \set{ A_k \given k \in \mscrK }.
  \end{equation*}

  The family \( \seq{ A_k }_{k \in \mscrK} \) is, formally, a set of Kuratowski pairs. Every pair \( \braket{ k, A_k } \) is itself a subset of \( \pow(\mscrK \bigcup \mscrA) \). The family is then a subset of \( \pow(\pow(\mscrK \bigcup \mscrA)) \). Applying the \hyperref[def:zfc/power_set]{axiom of power sets} again, we obtain that the family exits as a set.

  \SubProofOf{thm:zfc_existence_theorems/cartesian_product} By definition, the Cartesian product \( \prod_{k \in \mscrK} A_k \) is a set of indexed by \( \mscrK \) families of members of the union \( \bigcup\set{ A_k \given k \in \mscrK } \).

  We can apply the \hyperref[def:zfc/power_set]{axiom of power sets} one more time in addition to those in \fullref{thm:zfc_existence_theorems/indexed_family} to obtain the set of all indexed by \( \mscrK \) families of members of this union. Then we can apply the \hyperref[def:zfc/specification]{axiom schema of specification} to restrict only to those families that satisfy the condition of \fullref{def:cartesian_product/product}.

  \SubProofOf{thm:zfc_existence_theorems/set_of_relations} All relations between \( A \) and \( B \) are subsets of \( A \times B \), hence elements of \( \pow(A \times B) \). The latter exists by \fullref{thm:zfc_existence_theorems/cartesian_product} and \fullref{thm:zfc_existence_theorems/power_set}.

  \SubProofOf{thm:zfc_existence_theorems/set_of_functions} The set of single-valued functions from \( A \) to \( B \) is a subset of \( \pow(A \times B) \), hence it exists by \fullref{thm:zfc_existence_theorems/set_of_relations} and \fullref{thm:zfc_existence_theorems/power_set}.

  \SubProofOf{thm:zfc_existence_theorems/quotient_set} Let \( A \) be an arbitrary set and \( \cong \) be a binary relation over \( A \). Then \( A / {\cong} \) is a subset of \( \pow(A) \) and hence it exists as a consequence of the \hyperref[def:zfc/power_set]{axiom of power sets} and the \hyperref[def:zfc/specification]{axiom schema of specification} --- see \fullref{thm:equivalence_partition/partition}.

  \SubProofOf{thm:zfc_existence_theorems/function_evaluation} The set \( \seq{ f_k }_{k \in \mscrK} \) exists because it is a member of \( \fun(\mscrK, \fun(A, B)) \), which set exists by \fullref{thm:zfc_existence_theorems/set_of_functions}.

   Then \( \seq{ f_k(x) }_{k \in \mscrK} \) is the function
   \begin{equation*}
     \begin{aligned}
       &g_x: \mscrK \to B, \\
       &g_x(k) \coloneqq f_k(x).
     \end{aligned}
   \end{equation*}
\end{proof}

\begin{theorem}[Existence of single-valued selections]\label{thm:existence_of_single_valued_selections}
  Every \hyperref[def:set_valued_map/partial]{total set-valued map} has a \hyperref[def:function/selection]{single-valued selection}.
\end{theorem}
\begin{comments}
  \item Within \hyperref[def:zfc]{\logic{ZF}}, this theorem is equivalent to the \hyperref[def:zfc/choice]{axiom of choice} --- see \fullref{thm:axiom_of_choice_equivalences/selection}.
\end{comments}
\begin{proof}
  \ImplicationSubProof[def:zfc/choice]{the axiom of choice}[thm:existence_of_single_valued_selections]{selection existence} Let \( F: A \multto B \) be a total set-valued map. Consider the indexed family \( \set{ F(a) }_{a \in A} \). It is a (potentially empty) family of nonempty sets. Thus, we can apply the axiom of choice to obtain some \hyperref[def:choice_function]{choice function}
  \begin{equation*}
    c: \set{ F(a) \given a \in A } \to B.
  \end{equation*}

  Define
  \begin{equation*}
    \begin{aligned}
      &f: A \to B, \\
      &f(a) \coloneqq c(F(a)).
    \end{aligned}
  \end{equation*}

  Then \( f \) is a selection of \( F \).

  \ImplicationSubProof[thm:existence_of_single_valued_selections]{selection existence}[def:zfc/choice]{axiom of choice} Suppose that every total set-valued map has a single-valued selection.

  Fix a family \( \mscrA \) of nonempty sets and define the function
  \begin{equation*}
    \begin{aligned}
      &F: \mscrA \to \bigcup \mscrA, \\
      &F(A) \coloneqq A
    \end{aligned}
  \end{equation*}
  that sends each set in \( \mscrA \) to the corresponding subset of \( \bigcup \mscrA \). In terms of relations, we have \( (A, x) \in \gph(F) \) if and only if \( x \in A \). This is a total set-valued map because every set in \( \mscrA \) is nonempty.

  Then every selection of \( F \) is a choice function for \( \mscrA \).
\end{proof}

\begin{definition}\label{def:disjoint_union}\mcite[198]{Knapp2016BasicAlgebra}
  We define the \term{disjoint union} of the \hyperref[def:cartesian_product/indexed_family]{indexed family} \( \seq{ A_k }_{k \in \mscrK} \) of nonempty sets as
  \begin{equation*}
    \coprod_{k \in \mscrK} A_k \coloneqq \set[\Big]{ (k, x) \given k \in \mscrK \T{and} x \in A_k }.
  \end{equation*}
\end{definition}

\begin{theorem}[Surjective functions are right-invertible]\label{thm:surjective_functions_are_right_invertible}
  Every \hyperref[def:function_invertibility/surjective]{surjective function} is \hyperref[def:morphism_invertibility/right_invertible]{right-invertible}.
\end{theorem}
\begin{comments}
  \item Within \hyperref[def:zfc]{\logic{ZF}}, this theorem is equivalent to the \hyperref[def:zfc/choice]{axiom of choice} --- see \fullref{thm:axiom_of_choice_equivalences/surjective}.
\end{comments}
\begin{proof}
  Let \( f: A \to B \) be any function. Its \hyperref[def:set_valued_map/inverse]{inverse} \( f^{-1}: B \multto A \) is, by definition, a partial set-valued map.

  \ImplicationSubProof[def:zfc/choice]{the axiom of choice}[thm:epimorphisms_split_in_set]{right-invertibility} Suppose that \( f \) is surjective and thus satisfies \fullref{def:function_invertibility/surjective/inverse}, stating that its inverse is total.

  Then the axiom of choice via \fullref{thm:existence_of_single_valued_selections} gives us a \hyperref[def:function/selection]{single-valued selection} \( g \) of \( f^{-1} \).

  Since the value of \( f \) is \( y \) for all members of \( f^{-1}(y) \), and since \( g(y) \in f^{-1}(y) \), we have
  \begin{equation*}
    [f \bincirc g](y) = f(g(y)) = y.
  \end{equation*}

  Therefore, \( g \) is a right inverse of \( f \).

  \ImplicationSubProof[thm:epimorphisms_split_in_set]{right-invertibility}[def:zfc/choice]{the axiom of choice} Suppose that every surjective function is invertible.

  Let \( \mscrA \) be an arbitrary family of nonempty sets. Define the function
  \begin{equation*}
    \begin{aligned}
      &f: \coprod_{A \in \mscrA} A \to \mscrA, \\
      &f(A, x) \coloneqq A,
    \end{aligned}
  \end{equation*}
  where \( \coprod \) denotes the \hyperref[def:disjoint_union]{disjoint union}.

  This function is surjective, thus then there exists some right-inverse
  \begin{equation*}
    g: \mscrA \to \coprod_{A \in \mscrA} A.
  \end{equation*}

  Given a set \( A \) from the family \( \mscrA \), \( g(A) \) is a pair \( (A, x) \) with \( x \in A \). This allows us to define the choice function
  \begin{equation*}
    \begin{aligned}
      &c: \mscrA \to \bigcup \mscrA, \\
      &c(A) \coloneqq x \T{where} (A, x) = g(A).
    \end{aligned}
  \end{equation*}

  Since the family \( \mscrA \) was arbitrary, we can conclude that the axiom of choice holds.
\end{proof}

\begin{proposition}\label{thm:function_invertibility_categorical}
  In relation to \hyperref[def:morphism_invertibility]{morphism invertibility} in the category \hyperref[def:category_of_small_sets]{\( \cat{Set} \)}, we have the following:
  \begin{thmenum}
    \thmitem{thm:function_invertibility_categorical/empty} An \hyperref[def:set_valued_map/empty]{empty function} is always \hyperref[def:function_invertibility/injective]{injective} and \hyperref[def:morphism_invertibility/left_invertible]{left-invertible}, however only if its codomain is empty it is \hyperref[def:morphism_invertibility/left_cancellative]{left-cancellative}, \hyperref[def:morphism_invertibility/right_cancellative]{right-cancellative}, \hyperref[def:morphism_invertibility/right_invertible]{right-invertible} and \hyperref[def:function_invertibility/surjective]{surjective}.

    \thmitem{thm:function_invertibility_categorical/nonempty_left_invertible} A nonempty function is \hyperref[def:morphism_invertibility/left_invertible]{left-invertible} if and only if it is \hyperref[def:function_invertibility/injective]{injective}.

    \thmitem{thm:function_invertibility_categorical/left_cancellative} A function is \hyperref[def:morphism_invertibility/left_cancellative]{left-cancellative} if and only if it is \hyperref[def:function_invertibility/injective]{injective}.

    \thmitem{thm:function_invertibility_categorical/right_invertible} A function is \hyperref[def:morphism_invertibility/right_invertible]{right-invertible} if and only if it is \hyperref[def:function_invertibility/surjective]{surjective}.

    \thmitem{thm:function_invertibility_categorical/right_cancellative} A function is \hyperref[def:morphism_invertibility/right_cancellative]{right-cancellative} if and only if it is \hyperref[def:function_invertibility/surjective]{surjective}.

    \thmitem{thm:function_invertibility_categorical/fully_invertible} A function is \hyperref[def:function_invertibility/bijective]{bijective} if and only if it is \hyperref[def:morphism_invertibility/isomorphism]{fully invertible}.
  \end{thmenum}
\end{proposition}
\begin{comments}
  \item We prove this result here rather than in \fullref{subsec:functions} because it requires \fullref{thm:surjective_functions_are_right_invertible}, which requires the axiom of choice that we have introduced in this section.
\end{comments}
\begin{proof}
  \SubProofOf{thm:function_invertibility_categorical/empty} Let \( g: \varnothing \to C \) be the empty function to \( C \).

  \begin{itemize}
    \item It is vacuously injective.
    \item It is also left-invertible because the only function that can be composed on the right with is the unique function from \( \varnothing \) to \( \varnothing \).
    \item Clearly \( g: \varnothing \to C \) is surjective if and only if \( C = \varnothing \).
    \item For left-invertibility, note that \( g: \varnothing \to C \) composed with the function \( h: C \to D \) on the left is another empty function \( h \bincirc g: \varnothing \to D \). The latter is the identity \( \id_\varnothing \) if and only if \( C = D = \varnothing \).
    \item For right-invertibility, note that \( g: \varnothing \to C \) composed with the function \( f: A \to \varnothing \) on the right is the function \( g \bincirc f: A \to C \). But \( A = \varnothing \) since otherwise \( f \) would be nonempty, hence \( g \bincirc f: \varnothing \to C \). The latter is the identity \( \id_\varnothing \) if and only if \( C = \varnothing \).
    \item For right-cancellation, note that \( g \bincirc f_1 = g \bincirc f_2 \) implies \( f_1 = f_2 \) if and only if \( B \) is empty.
  \end{itemize}

  \SubProofOf{thm:function_invertibility_categorical/nonempty_left_invertible}

  \SufficiencySubProof Let \( f: A \to B \) be a nonempty injective function. It thus satisfies \fullref{def:function_invertibility/injective/inverse}, hence the \hyperref[def:set_valued_map/inverse]{inverse} \( f^{-1}: B \to A \) is a partial single-valued function.

  Fix some value \( a \in A \) and define
  \begin{equation*}
    \begin{aligned}
      &g: B \to A \\
      &g(y) \coloneqq \begin{cases}
        f^{-1}(y), &y \in f(A) \\
        a,         &\T{otherwise.}
      \end{cases}
    \end{aligned}
  \end{equation*}

  This function \( g \) is a left inverse of \( f \) because, for any \( x \in A \),
  \begin{equation*}
    [g \bincirc f](x)
    =
    g(f(x))
    =
    f^{-1}(f(x))
    =
    x.
  \end{equation*}

  We can see that \( g \) is not unique because of our choice of \( a \). We may even define \( g \) to take different values in \( A \) outside \( f(A) \). Thus, \( g \) is non-unique in general. But at least one such function exists.

  \NecessitySubProof Conversely, suppose that \( f: A \to B \) is not necessarily injective and let \( g: B \to A \) be a left inverse of \( f \). Let \( x_1 \) and \( x_2 \) be two different points in \( A \). Since \( g \bincirc f = \id_A \), clearly \( g(f(x_1)) \neq g(f(x_2)) \). If we suppose that \( f(x_1) = f(x_2) \), we would obtain a contradiction since then \( g(f(x_1)) \) would equal \( g(f(x_2)) \). Hence, \( f(x_1) \neq f(x_2) \). This shows that \( f \) is injective.

  \SubProofOf{thm:function_invertibility_categorical/left_cancellative}

  \SufficiencySubProof* The case with an empty function is handled in \fullref{thm:function_invertibility_categorical/empty}, and we assume that it is nonempty.

  Suppose that \( g: B \to C \) is a nonempty left-cancellative function. Let \( y_1 \) and \( y_2 \) be some members of \( B \) such that \( g(y_1) = g(y_2) \).

  Suppose that \( y_1 \neq y_2 \) and define the function
  \begin{equation*}
    \begin{aligned}
      &f: B \to B \\
      &f(y) \coloneqq \begin{cases}
        y_2, &y = y_1 \\
        y_1, &y = y_2 \\
        y,   &y \neq y_1 \T{and} y \neq y_2
      \end{cases}
    \end{aligned}
  \end{equation*}

  Then
  \begin{equation*}
    g(f(y_2)) = g(y_1) = g(y_2) = g(f(y_1))
  \end{equation*}

  For all \( y \in B \) different from \( y_1 \) and \( y_2 \), we have \( y = f(y) \).

  Since \( g \) is left-cancellative, from \( g \bincirc \id_B = g \bincirc f \) it follows that \( \id_B = f \), which is a contradiction.

  It remains for \( y_1 \) to be equal to \( y_2 \). Since these were arbitrary points in \( B \) satisfying \( g(y_1) = g(y_2) \), we conclude that \( g \) is injective.

  \NecessitySubProof* Conversely, if \( f \) is injective, it is left-invertible by \fullref{thm:function_invertibility_categorical/nonempty_left_invertible} and left-cancellative by \fullref{thm:def:morphism_invertibility/split_monomorphism}.

  \SubProofOf{thm:function_invertibility_categorical/right_invertible}

  \SufficiencySubProof* In one direction, we have \fullref{thm:surjective_functions_are_right_invertible}.

  \NecessitySubProof* Conversely, suppose that \( g: B \to A \) is a right inverse of \( f: A \to B \). Let \( y \in B \). We have that \( g(y) \) is in the preimage of \( y \) under \( f \) because \( f(g(y)) = y \). Thus, the preimage is not empty for an arbitrary point in \( B \). We conclude that \( f \) is surjective.

  \SubProofOf{thm:function_invertibility_categorical/right_cancellative} The case with an empty function is handled in \fullref{thm:function_invertibility_categorical/empty}, and we assume that it is nonempty.

  \SufficiencySubProof* Let \( f: A \to B \) be a nonempty right-cancellative function. Suppose that it is not surjective and let \( y_0 \in B \setminus \img f \). Let \( z \) be some set not belonging to \( B \). Define the function
  \begin{equation*}
    \begin{aligned}
      &g: B \to B \cup \set{ z } \\
      &g(y) \coloneqq \begin{cases}
        z, &y = y_0 \\
        y, &y \neq y_0
      \end{cases}
    \end{aligned}
  \end{equation*}

  Since \( f \) is right-cancellative, from \( \id_B \bincirc f = g \bincirc f \) it follows that \( \id_B = g \), which is a contradiction. Therefore, \( f \) is surjective.

  \NecessitySubProof* We can prove the converse using \fullref{thm:function_invertibility_categorical/right_invertible} like we did for \fullref{thm:function_invertibility_categorical/left_cancellative}, however we prefer a direct proof that does not rely on the axiom of choice. As a bonus, this would allow us to prove \fullref{thm:epimorphisms_split_in_set}.

  Conversely, suppose that \( f: A \to B \) is surjective and that for some functions \( g_1, g_2: B \to C \) we have
  \begin{equation*}
    g_1 \bincirc f = g_2 \bincirc f.
  \end{equation*}

  Fix some \( y \in B \). Because \( f \) is surjective, there exists some \( x \in A \) such that \( f(x) = y \). Then
  \begin{equation*}
    g_1(y) = [g_1 \bincirc f](x) = [g_2 \bincirc f](x) = g_2(y).
  \end{equation*}

  Since \( y \in B \) was arbitrary, we conclude that \( f \) is right-cancellative.

  it is right-invertible by \fullref{thm:surjective_functions_are_right_invertible} and right-cancellative by \fullref{thm:def:morphism_invertibility/split_epimorphism}.

  \SubProofOf{thm:function_invertibility_categorical/fully_invertible}

  \SufficiencySubProof* If \( f: \varnothing \to B \) is a bijective empty function, then it is surjective and, by \fullref{thm:function_invertibility_categorical/right_invertible}, it is right-invertible. By \fullref{thm:function_invertibility_categorical/empty}, it is also left-invertible. Thus, it is fully invertible.

  \NecessitySubProof* Conversely, if an empty function \( f: \varnothing \to B \) is fully invertible, by \fullref{thm:function_invertibility_categorical/empty} we have \( A = B = \varnothing \) and hence it is bijective.

  Finally, if \( f: A \to B \) is \hi{nonempty} function, then by \fullref{thm:function_invertibility_categorical/nonempty_left_invertible} and \fullref{thm:function_invertibility_categorical/right_invertible} it is bijective if and only if it is fully invertible.

  In the bijective case, we can avoid the axiom of choice via \fullref{thm:surjective_functions_are_right_invertible} by noting that if \( f \) is bijective, its inverse is single-valued, and thus it is not necessary to do a selection of \( f^{-1} \).
\end{proof}

\begin{theorem}[Axiom of choice equivalences]\label{thm:axiom_of_choice_equivalences}
  The following statements are commonly referred to as \enquote{the} \hyperref[def:zfc/choice]{axiom of choice}:
  \begin{thmenum}[series=thm:axiom_of_choice_equivalences]
    \thmitem{thm:axiom_of_choice_equivalences/choice_sets} For every family of nonempty sets, \( \mscrA \) there exists a set \( B \) such that \( A \cap B \) is a singleton set for every \( A \in \mscrA \).

    \thmitem{thm:axiom_of_choice_equivalences/choice_function} Every family of nonempty sets has a corresponding \hyperref[def:choice_function]{choice function}.

    \thmitem{thm:axiom_of_choice_equivalences/choice_product} The \hyperref[def:cartesian_product]{Cartesian product} of a family of nonempty sets is nonempty.
  \end{thmenum}

  The following statements are equivalent to the axiom of choice, but are not conflated with it:
  \begin{thmenum}[resume=thm:axiom_of_choice_equivalences]
    \thmitem{thm:axiom_of_choice_equivalences/selection} \Fullref{thm:existence_of_single_valued_selections}: Every \hyperref[def:set_valued_map/partial]{total set-valued map} has a \hyperref[def:function/selection]{single-valued selection}.

    \thmitem{thm:axiom_of_choice_equivalences/surjective} \Fullref{thm:surjective_functions_are_right_invertible}: Every \hyperref[def:function_invertibility/surjective]{surjective function} is \hyperref[def:morphism_invertibility/right_invertible]{right-invertible}.

    \thmitem{thm:axiom_of_choice_equivalences/epimorphisms} \Fullref{thm:epimorphisms_split_in_set}: Every \hyperref[def:morphism_invertibility/right_cancellative]{epimorphism} in \hyperref[def:category_of_small_sets]{\( \cat{Set} \)} splits.

    \thmitem{thm:axiom_of_choice_equivalences/fully_faithful_essentially_surjective} \Fullref{thm:fully_faithful_and_essentially_surjective_functor_induces_equivalence}: Every \hyperref[def:functor_invertibility/fully_faithful]{fully faithful} and \hyperref[def:functor_invertibility/surjective_on_objects]{essentially surjective on objects} functor induces a \hyperref[def:category_equivalence]{category equivalence}.

    \thmitem{thm:axiom_of_choice_equivalences/skeletons} \Fullref{thm:category_skeleton_existence}: Every \hyperref[def:category]{category} has a \hyperref[def:skeletal_category]{skeleton}.

    \thmitem{thm:axiom_of_choice_equivalences/well_ordering} \Fullref{thm:well_ordering_theorem}: Every \hyperref[def:set]{set} can be \hyperref[def:well_ordered_set]{well-ordered}.

    \thmitem{thm:axiom_of_choice_equivalences/zorns_lemma} \Fullref{thm:zorns_lemma}: If every \hyperref[def:partial_order_chain]{chain} in a \hyperref[def:partially_ordered_set]{partially ordered set} has an \hyperref[def:extremal_points/bounds]{upper bound}, then the entire set has a \hyperref[def:extremal_points/maximal_and_minimal_element]{maximal element}.

    \thmitem{thm:axiom_of_choice_equivalences/maximal_ideal} \Fullref{thm:maximal_ideal_theorem}: Every proper \hyperref[def:semiring_ideal]{semiring ideal} is contained in a \hyperref[def:semiring_ideal/maximal]{maximal ideal}.

    \thmitem{thm:axiom_of_choice_equivalences/vector_space_bases} \Fullref{thm:vector_space_basis_existence}: Every \hyperref[def:vector_space]{vector space} has a \hyperref[def:hamel_basis]{basis}.

    \thmitem{thm:axiom_of_choice_equivalences/tychonoff} \Fullref{thm:tychonoffs_product_theorem}: The \hyperref[def:topological_product]{topological product} of \hyperref[def:compact_space]{compact spaces} is compact.
  \end{thmenum}
\end{theorem}
\begin{proof}
  The equivalence proofs can be found in the linked theorems since that is usually the most appropriate place to put them.
\end{proof}

\begin{theorem}[Diaconescu-Goodman-Myhill theorem]\label{thm:diaconescu_goodman_myhill_theorem}\mcite[corr. 2]{Diaconescu1975}
  In \hyperref[def:zfc]{\logic{ZF}}, the \hyperref[def:zfc/choice]{axiom of choice} entails the law of the excluded middle \eqref{eq:thm:minimal_propositional_negation_laws/lem}.
\end{theorem}

\begin{remark}\label{rem:well_founded_relation}
  If \( \prec \) is simply a binary relation without any additional conditions imposed, we can still make sense of the chain conditions from \fullref{def:chain_condition}. We will restrict ourselves to the descending chain condition --- we call the relation \( \prec \) \term{well-founded} if any of the following equivalent conditions hold:

  \begin{thmenum}
    \thmitem{rem:well_founded_relation/maximal} \Fullref{def:chain_condition/maximal} can be restated as follows: every nonempty subset of \( P \) has a minimal element \( m \) in the sense that there does not exist any other element \( x \) from \( P \) such that \( x \prec m \).

    \thmitem{rem:well_founded_relation/stabilization} \Fullref{def:chain_condition/stabilization} can be restated as follows: every nonstrict descending sequence
    \begin{equation}\label{eq:rem:well_founded_relation/stabilization}
      \cdots \prec x_3 \prec x_2 \prec x_1
    \end{equation}
    stabilizes in the sense that there exists an index \( n \) such that \( x_k = x_n \) for \( k > n \).

    \thmitem{rem:well_founded_relation/infinite} \Fullref{def:chain_condition/infinite} can be restarted as follows: there exists no strictly infinitely ascending descending sequence in \( P \), i.e. a sequence of the form \eqref{eq:rem:well_founded_relation/stabilization} such that no two members are equal.
  \end{thmenum}
\end{remark}

\begin{proposition}\label{thm:set_membership_is_well_founded}
  Set membership is \hyperref[rem:well_founded_relation]{well-founded} in \logic{ZFC}. More precisely, given a set \( A \), if we regard \( \in \) as a binary relation between members of \( A \), then we would obtain that \( \in \) is a well-founded relation.
\end{proposition}
\begin{proof}
  The empty set is vacuously well-founded, so suppose that \( A \) is nonempty.

  Aiming at a contradiction, suppose also that \( \in \) is not well-founded on \( A \). Then there exists an strictly descending sequence \( \seq{ x_k }_{k=1}^\infty \subseteq A \) such that
  \begin{equation*}
    \cdots \in x_3 \in x_2 \in x_1.
  \end{equation*}

  Denote by \( B \) the set \( \set{ x_k \given k = 1, 2, \ldots } \). By the \hyperref[def:zfc/foundation]{axiom of foundation}, \( B \) contains a set \( C \) which is disjoint from \( B \).

  Clearly \( C \) coincides with at least one of \( x_1, x_2, \ldots \). Let \( C = x_{k_0} \). Since \( x_{k_0} \cap B = \varnothing \), then \( x_{k_0 + 1} \) is either not a member of \( x_{k_0} \) or of \( B \). But it is a member of both by our assumption that the sequence is infinitely descending.

  The obtained contradiction proves that \( \in \) is well-founded on \( A \).
\end{proof}

\begin{corollary}\label{thm:simple_foundation_theorems}
  We list several statements, which are consequences of the \hyperref[def:zfc/foundation]{axiom of foundation} via \fullref{thm:set_membership_is_well_founded}:

  \begin{thmenum}
    \thmitem{thm:simple_foundation_theorems/member_of_itself} No set is a member of itself.

    \thmitem{thm:simple_foundation_theorems/mutual_membership} For any two sets \( A \) and \( B \), it is not possible for both \( A \in B \) and \( B \in A \) to hold.

    \thmitem{thm:simple_foundation_theorems/ordinal_successor_is_injective} The \hyperref[def:ordinal_successor]{ordinal successor operation} is injective. That is, for every two set \( A \) and \( B \) from \( \op{succ}(A) = \op{succ}(B) \) it follows that \( A = B \).
  \end{thmenum}
\end{corollary}
\begin{proof}
  \SubProofOf{thm:simple_foundation_theorems/member_of_itself} The sequence
  \begin{equation*}
    \cdots A \in A \in A
  \end{equation*}
  is infinitely descending, hence \( A \) cannot be a member of itself.

  \SubProofOf{thm:simple_foundation_theorems/mutual_membership} The sequence
  \begin{equation*}
    \cdots A \in B \in A \in B \in A
  \end{equation*}
  is infinitely descending, hence \( A \in B \) and \( B \in A \) cannot simultaneously hold.

  \SubProofOf{thm:simple_foundation_theorems/ordinal_successor_is_injective} Note that \( A \in \op{succ}(A) = \op{succ}(B) \), hence either \( A = B \) or \( A \in B \). We can analogously conclude that either \( A = B \) or \( B \in A \).

  But \( A \in B \) and \( B \in A \) cannot simultaneously hold due to \fullref{thm:simple_foundation_theorems/mutual_membership}. Hence, it remains for \( A = B \) to hold.
\end{proof}

\begin{theorem}[Well-founded induction]\label{thm:well_founded_induction}
  Let \( \mscrL \) be the \hyperref[def:first_order_language]{first-order language} with no functional symbols and a single predicate symbol \( \prec \). We have already mentioned that we cannot formalize the concept of well-foundedness in first-order logic alone, so we will work with \hyperref[def:first_order_structure]{structures} directly.

  Every \hyperref[rem:well_founded_relation]{well-founded} structure \( \mscrX = (X, I) \) over \( \mscrL \) satisfies an inductive axiom schema. For every formula \( \varphi \) over \( \mscrL \) not containing \( \eta \) nor \( \zeta \) as free variables, \( X \) satisfies
  \begin{equation}\label{eq:thm:well_founded_induction}
    \qforall \eta
    \parens[\Big]
      {
        \overbrace
          {
            \underbrace{ \parens[\Big]{ \qforall {\zeta \prec \eta} \varphi[\xi \mapsto \zeta] } }_{\mathclap{\substack{\T{inductive} \\ \T{hypothesis}}}}
            \synimplies
            \underbrace{ \varphi[\xi \mapsto \eta] }_{\mathclap{\substack{\T{inductive step} \\ \T{conclusion}}}}
          }^{\T{inductive step}}
      }
    \synimplies
    \underbrace{ \qforall \eta \varphi[\xi \mapsto \eta] }_{\T{conclusion}}.
  \end{equation}
\end{theorem}
\begin{comments}
  \item See the comments in \fullref{def:peano_arithmetic/PA3} regarding variables and quantification in axiom schemas and \fullref{rem:induction} for a general discussion of induction.
  \item In the special case where \( X = \BbbN \), this is called \term{strong induction} compared to the usual natural number induction \eqref{eq:def:peano_arithmetic/PA3}. This is discussed in \fullref{rem:induction/well_founded}.
\end{comments}
\begin{proof}
  Fix a formula \( \varphi \) in \( \mscrL \). Fix a variable assignment \( v \) in \( X \). We will show that the \hyperref[def:material_implication/contrapositive]{contraposition} of \eqref{eq:thm:well_founded_induction} holds in this model.

  Suppose that there exists a value \( x \in X \) such that \( \Bracks{\varphi}_{v_{\xi \mapsto x}} = F \), that is,
  \begin{equation*}
    \qexists \eta \synneg \varphi[\xi \mapsto \eta].
  \end{equation*}
  holds.

  Let \( x_0 \) be the smallest such value (which is guaranteed to exist because \( X \) is well-founded by \( \prec \)). Thus, the inductive hypothesis \( \qforall {\zeta < x_0} \varphi[\xi \mapsto \zeta] \) holds, but the inductive step conclusion \( \varphi[\xi \mapsto x_0] \) does not.

  Since \( v \) was chosen arbitrarily, this is true for all variable assignments in \( X \). Then the following formula is \hyperref[def:first_order_model]{valid} in \( \mscrX \):
  \begin{equation}\label{eq:thm:well_founded_induction/contraposition}
    \qexists \eta \synneg \varphi[\xi \mapsto \eta]
    \synimplies
    \qexists \eta
    \parens[\Big]
      {
        \qforall {\zeta \prec \eta} \varphi[\xi \mapsto \zeta] \synwedge \synneg \varphi[\xi \mapsto \eta]
      }
  \end{equation}

  This is precisely the contrapositive of \eqref{eq:thm:well_founded_induction}. Since we are working in classical logic, the contrapositive is semantically equivalent to its original implication, hence \( \eqref{eq:thm:well_founded_induction} \) is also valid in \( \mscrX \).
\end{proof}

\begin{theorem}[Epsilon induction]\label{thm:epsilon_induction}
  For every formula \( \varphi \) in the language of set theory not containing \( \eta \) nor \( \zeta \) as free variables, the following is a theorem of \logic{ZFC}:
  \begin{equation*}
    \qforall \eta
    \parens[\Big]
      {
        \overbrace
          {
            \underbrace{ \parens[\Big]{ \qforall {\zeta \in \eta} \varphi[\xi \mapsto \zeta] } }_{\mathclap{\substack{\T{inductive} \\ \T{hypothesis}}}}
            \synimplies
            \underbrace{ \varphi[\xi \mapsto \eta] }_{\mathclap{\substack{\T{inductive step} \\ \T{conclusion}}}}
          }^{\T{inductive step}}
      }
    \synimplies
    \underbrace{ \qforall \eta \varphi[\xi \mapsto \eta] }_{\T{conclusion}}.
  \end{equation*}
\end{theorem}
\begin{comments}
  \item This induction schema is called \enquote{\( \varepsilon \)-induction} because the set membership symbol \( \in \) is derived from \( \varepsilon \) as explained in \fullref{rem:epsilon_and_set_membership}.

  \item See the comments in \fullref{def:peano_arithmetic/PA3} regarding variables and quantification in axiom schemas and \fullref{rem:induction} for a general discussion of induction.
\end{comments}
\begin{proof}
  Every model of \logic{ZFC} is well-founded by \( \in \) as a consequence of \fullref{thm:set_membership_is_well_founded}. The corollary then follows from \fullref{thm:well_founded_induction}.
\end{proof}
